### 2021-06-30 tables/accuracy_table.Rmd
Rename folder

```{r setup, include = FALSE, cache = TRUE}
knitr::opts_chunk$set(echo = FALSE)
library(here)
source(here("code/03_build_models.R"))
```
```{r make_table, include = FALSE} 
# Load packages and functions ---------------------------------------------

library(here)
library(glue)
library(kableExtra)
source(here("code/functions/table_functions.R"))

# Load models data --------------------------------------------------------

source(here("code/03_build_models.R"))

# Build table -------------------------------------------------------------

# Create a list with equation coefficients and accuracy indices
placements <- c("Ankle", "Lower back", "Hip")
coeffs <- list(
  ankle = list(
    resultant = GRF_coefficients_jumping$ankle_resultant,
    vertical = GRF_coefficients_jumping$ankle_vertical
  ),
  back = list(
    resultant = GRF_coefficients_jumping$lower_back_resultant,
    vertical = GRF_coefficients_jumping$lower_back_vertical
  ),
  hip = list(
    resultant = GRF_coefficients_jumping$hip_resultant,
    vertical = GRF_coefficients_jumping$hip_vertical
  )
)
accuracy <- list(
  ankle = list(
    resultant = GRF_accuracy_jumping$ankle_resultant,
    vertical = GRF_accuracy_jumping$ankle_vertical
  ),
  back = list(
    resultant = GRF_accuracy_jumping$lower_back_resultant,
    vertical = GRF_accuracy_jumping$lower_back_vertical
  ),
  hip = list(
    resultant = GRF_accuracy_jumping$hip_resultant,
    vertical = GRF_accuracy_jumping$hip_vertical
  )
)

a <- c("Ankle", "Lower back", "Hip")
b <- c("ankle", "back", "hip")
table <- map2_dfr(a, b, ~ accuracy_table(.x, coeffs[[.y]], accuracy[[.y]]))
```
```{r table, echo = FALSE}
kbl(table, booktabs = TRUE) %>%
  kable_styling(latex_options = "scale_down")
```

### 2021-06-30 code/01_tidy_data.R
Keep only jumping data

@@ -2,197 +2,26 @@
 
 library(here)
 library(tidyverse)
-library(readxl)
-library(janitor)
 library(lvmisc)
 
 # Read anthropometric data ------------------------------------------------
 
-anthropometric <- read_csv(here("data/raw/anthropometric_data.csv")) %>%
-  select(id = ID, everything()) %>%
+anthropometric <- read_csv(here("data", "anthropometric_data.csv")) |>
+  select(id = ID, everything()) |>
   mutate(
     BMI = bmi(body_mass, height / 100),
     BMI_cat = bmi_cat(BMI)
   )
 
-# Obtain running data -----------------------------------------------------
+# Read jumping data -------------------------------------------------------
 
-# Data with force and acceleration
-running_df_wide <- read_xlsx(
-  here("data/raw/GRF_ACC_data_all.xlsx"), sheet = "all"
-) %>%
-  clean_names() %>%
-  select(
-    raw_imu, id, trial = visita, filename = nome,
-    acc_placement = local_acelerometro,
-    speed = velocidade_km_h,
-    res_n_peaks = ftot_numero_picos_ftot,
-    ver_n_peaks = fvt_numero_picos_fvt,
-    pRACC_g = atot_med_picos_atot_g,
-    pRGRF_N = ftot_med_picos_ftot_n,
-    pRGRF_BW = ftot_med_picos_ftot_bw,
-    pVACC_g = avt_med_picos_avt_g,
-    pVGRF_N = fvt_med_picos_fvt_n,
-    pVGRF_BW = fvt_med_picos_fvt_bw
-  ) %>%
-  filter(
-    raw_imu == "imu" &
-    speed > 6 &
-    acc_placement %in% c("ankle", "back", "waist")
-  ) %>%
-  select(id, everything())
-
-running_df_res <- running_df_wide %>%
-  select(
-    id, trial, filename, acc_placement, speed,
-    n_peaks = res_n_peaks, pACC_g = pRACC_g,
-    pGRF_N = pRGRF_N, pGRF_BW = pRGRF_BW
-  ) %>%
+mechanical_load_data <- read_csv(here("data", "jumping_data.csv")) |>
+  select(-c(body_mass, BMI, BMI_cat)) |>
+  left_join(anthropometric, by = c("id", "trial")) |>
   mutate(
-    vector = "resultant",
-    .after = acc_placement
[…]

### 2021-07-01 code/01_tidy_data.R
Create a variable to identify the jump type and height

@@ -19,7 +19,12 @@ mechanical_load_data <- read_csv(here("data", "jumping_data.csv")) |>
   select(-c(body_mass, BMI, BMI_cat)) |>
   left_join(anthropometric, by = c("id", "trial")) |>
   mutate(
-    across(c(acc_placement, vector, jump_type), as_factor)
+    across(c(acc_placement, vector, jump_type), as_factor),
+    jump = as_factor(
+      str_to_sentence(paste0(jump_type, " ", jump_height, "cm"))
+    ),
+    jump = fct_relevel(jump, "Drop jumps 40cm", after = 7),
+    .after = jump_height
   )
 
 # Export data as rda ------------------------------------------------------

### 2021-07-01 code/02_explore.R
Update to include only jump data

@@ -1,171 +1,100 @@
-# Load packages and functions ---------------------------------------------
+# Load packages -----------------------------------------------------------
 
 library(here)
 library(tidyverse)
-source(here("code/functions/plot_functions.R"))
-source(here("code/functions/utils.R"))
+library(corrr)
+library(lvmisc)
 
 # Load data ---------------------------------------------------------------
 
-running_df <- read_csv(here("data/processed/running_data.csv")) %>%
-  mutate(
-    acc_placement = as_factor(acc_placement),
-    vector = as_factor(vector),
-    speed = as_factor(speed),
-    run = as_factor(paste0("running ", speed, "km/h")),
-    BMI_cat = fct_relevel(
-      as_factor(BMI_cat),
-      levels = c(
-        "Normal weight", "Overweight", "Obesity class I"
-      )
-    )
-  )
-
-jumping_df <- read_csv(here("data/processed/jumping_data.csv")) %>%
-  mutate(
-    acc_placement = as_factor(acc_placement),
-    vector = as_factor(vector),
-    jump_type = as_factor(jump_type),
-    jump_height = as_factor(jump_height),
-    jump = as_factor(paste0(jump_type, " ", jump_height, "cm")),
-    jump = fct_relevel(jump, "drop jumps 40cm", after = 7),
-    BMI_cat = fct_relevel(
-      as_factor(BMI_cat),
-      levels = c(
-        "Normal weight", "Overweight", "Obesity class I",
-        "Obesity class II", "Obesity class III"
-      )
-    )
-  )
+load(here("data", "mechanical_load_data.rda"))
 
 # Sample size per activity ------------------------------------------------
 
-running_df %>%
-  group_by(acc_placement, vector, speed) %>%
-  select(acc_placement, vector, speed, pGRF_N, pLR_Ns) %>%
-  summarise_all(~ sum(!is.na(.))) %>%
-  knitr::kable()
+sample_size <- mechanical_load_data |>
+  group_by(acc_placement, vector, jump_type, jump_height) |>
+  select(acc_placement, vector, jump_type, jump_height, pGRF_N, pLR_Ns) |>
+  summarise_all(~ sum(!is.na(.)))
 
[…]

### 2021-07-02 code/03_build_models.R
Update models

@@ -1,191 +1,132 @@
-# Load packages and functions ---------------------------------------------
+# Load packages -----------------------------------------------------------
 
 library(here)
 library(tidyverse)
 library(lme4)
-source(here("code/functions/utils.R"))
-source(here("code/functions/model_functions.R"))
-source(here("code/functions/plot_functions.R"))
+library(lvmisc)
+library(emmeans)
 
-# Load data ---------------------------------------------------------------
+# Load and prepare data ---------------------------------------------------
 
-running_df <- read_csv(here("data/processed/running_data.csv"))
-jumping_df <- read_csv(here("data/processed/jumping_data.csv"))
+load(here("data", "mechanical_load_data.rda"))
 
-# Build models ------------------------------------------------------------
+# Create a list of 3 data frames: one for each accelerometer placement
+mechanical_load_by_placements <- map(
+  c("ankle", "lower_back", "hip"),
+  ~ filter(mechanical_load_data, acc_placement == .x)
+) %>% set_names(c("ankle", "lower_back", "hip"))
 
-model_df <- data.frame(
-  vector = c(rep("resultant", 3), rep("vertical", 3)),
-  placement = rep(c("ankle", "lower_back", "hip"), 2)
+# Filter the data frames in the list, separating the vertical and resultant
+# vectors
+res_mechanical_load_data <- map(
+  mechanical_load_by_placements, filter, vector == "resultant"
+)
+ver_mechanical_load_data <- map(
+  mechanical_load_by_placements, filter, vector == "vertical"
 )
 
-names <- paste0(model_df$placement, "_", model_df$vector)
-
-running_data <- map2(
-  model_df$vector, model_df$placement,
-  ~ filter_data(running_df, .x, .y)
-)
-
-jumping_data <- map2(
-  model_df$vector, model_df$placement,
-  ~ filter_data(jumping_df, .x, .y)
-)
+# Build GRF models --------------------------------------------------------
 
 GRF_formula <- as.formula(
[…]

### 2021-07-02 code/functions/model_functions.R
Delete files

@@ -1,69 +1,0 @@
-get_coefficients <- function(model) {
-  # get_coefficients extract coefficients from a model.
-  #
-  # Args:
-  #   model: A lmerMod object.
-  #
-  # Returns:
-  #   A named numeric vector with the model coefficients.
-  coefficients <- stats::coef(summary(model))
-  coefficients[, 1]
-}
-
-# Define functions to extract data sets
-get_training_data <- function(x) rsample::analysis(x)
-get_testing_data <- function(x) rsample::assessment(x)
-
-loocv <- function(data, formula) {
-  # loocv cross-validates the model using the leave-one-out cross-validation
-  #  (LOOCV) approach.
-  #
-  # Args:
-  #   data: A data frame.
-  #   formula: The model formula.
-  #
-  # Returns:
-  #   A data frame with the actual and predicted values.
-  outcome <- as.character(formula.tools::lhs(formula))
-  loocv_split <- rsample::group_vfold_cv(data, group = "subj")
-  training_data <- purrr::map(loocv_split$splits, get_training_data)
-  testing_data <- purrr::map(loocv_split$splits, get_testing_data)
-
-  trained_models <- purrr::map(training_data, ~ lme4::lmer(formula, data = .x))
-  predicted <- purrr::map2(
-    trained_models, testing_data,
-    ~ stats::predict(.x, newdata = .y, allow.new.levels = TRUE)
-  ) %>%
-    purrr::as_vector() %>%
-    unname()
-  actual <- purrr::map(testing_data, outcome) %>%
-    purrr::as_vector()
-
-  if ("jump_type" %in% colnames(data)) {
-    activity_type <- data$jump_type
-  } else {
-    activity_type <- data$speed
-  }
-
-    tibble::tibble(
-    subj = data$subj, acc_placement = data$acc_placement, vector = data$vector,
-    activity_type, actual, predicted
-  )
-}
-
-accuracy <- function(data, formula, actual, predicted) {
-  # accuracy computes some accuracy indices.
-  #
-  # Args:
-  #   data: A data frame.
-  #   formula: The model formula.
-  #   actual, predicted: A numerical vector.
-  #
-  # Returns:
-  #   A data frame with the accuracy indices.
-  R2 <- piecewiseSEM::rsquared(lme4::lmer(formula, data = data))[[6]]
[…]

### 2021-07-02 code/functions/plot_functions.R

@@ -1,147 +1,0 @@
-histogram <- function(data, vector_name, var, grid_y, grid_x) {
-  # histogram plots a histogram of var with a facet_grid by the specified
-  # variables
-  #
-  # Args:
-  #   data: A data frame.
-  #   vector_name: A character string with the vector name. Either "resultant"
-  #   or "vertical".
-  #   var: A character string with the variable name to be plot in the
-  #   histogram.
-  #   grid_y, grid_x: A character string with the name of the variables to be
-  #   used in the facet_grid().
-  #
-  # Returns:
-  #   A ggplot object.
-  ggplot2::ggplot(dplyr::filter(data, vector == vector_name)) +
-    ggplot2::geom_histogram(ggplot2::aes(x = .data[[var]])) +
-    ggplot2::facet_grid(
-      ggplot2::vars(.data[[grid_y]]), ggplot2::vars(.data[[grid_x]])
-    ) +
-    ggplot2::labs(title = vector_name)
-}
-
-box_plot <- function(data, x, y, placement) {
-  # box_plot plots a boxplot of y by x, grouping by vector (resultant and
-  # vertical) and labeling outliers.
-  #
-  # Args:
-  #   data: A data frame.
-  #   x, y: The name of the variable (unquoted) to be plot in the x and y axis.
-  #   placement: A character string with the name of the accelerometer
-  #   placement (either "ankle", "low back" or "hip").
-  #
-  # Returns:
-  #   A ggplot object.
-  data <- dplyr::mutate(
-    dplyr::group_by(data, {{ x }}, vector),
-    outlier = ifelse(lvmisc::is_outlier({{ y }}, na.rm = TRUE), subj, NA)
-  )
-  ggplot2::ggplot(
-    data = data, ggplot2::aes(x = {{ x }}, y = {{ y }}, fill = vector)
-  ) +
-    ggplot2::geom_boxplot() +
-    ggrepel::geom_text_repel(
-      ggplot2::aes(label = outlier), na.rm = TRUE, position = position_dodge(1)
-    ) +
-    ggplot2::labs(title = placement)
-}
-
-scatterplot <- function(data, x, y, placement) {
-  # scatterplot plots a scatterplot of y by x, with different colours by
-  # BMI_cat and facets by vector.
-  #
-  # Args:
-  #   data: A data frame.
-  #   x, y: The name of the variable (unquoted) to be plot in the x and y axis.
[…]

### 2021-07-02 code/functions/table_functions.R

@@ -1,51 +1,0 @@
-get_equation <- function(outcome, coeffs) {
-  acc <- NULL
-  if (stringr::str_detect(outcome, "pRGRF")) {
-    outcome <- "pRGRF (N)"
-    acc <- "pRACC"
-  } else if (stringr::str_detect(outcome, "pVGRF")) {
-    outcome <- "pVGRF (N)"
-    acc <- "pVACC"
-  }
-
-  coeffs <- round(coeffs, 4)
-
-  coeffs[1] <- ifelse(
-    coeffs[1] > 0,
-    coeffs[1],
-    paste0(" - ", as.character(abs(coeffs[1])))
-  )
-  coeffs[-1] <- purrr::map(
-    as.numeric(coeffs[-1]),
-    ~ ifelse(
-      .x > 0,
-      paste0(" + ", as.character(abs(.x))),
-      paste0(" - ", as.character(abs(.x)))
-    )
-  )
-
-  glue::glue(
-    "{outcome} = {coeffs[1]}{coeffs[2]}({acc})\\
-    {coeffs[3]}(body mass)\\
-    {coeffs[4]}({acc} x body mass)"
-  )
-}
-
-accuracy_table <- function(placement, coeffs, accuracy) {
-  tibble::tibble(
-   "Accelerometer placement" = c(placement, ""),
-   "Regression equations" = c(
-     get_equation("pRGRF", coeffs$resultant),
-     get_equation("pVGRF", coeffs$vertical)
-   ),
-   "R2" = c(accuracy$resultant$R2, accuracy$vertical$R2),
-   "MAE" = c(accuracy$resultant$MAE, accuracy$vertical$MAE),
-   "MAPE" = c(accuracy$resultant$MAPE, accuracy$vertical$MAPE),
-   "RMSE" = c(accuracy$resultant$RMSE, accuracy$vertical$RMSE)
-  ) %>%
-  dplyr::mutate(
-    R2 = round(R2, 2),
-    MAE = round(MAE, 1),
-    RMSE = round(RMSE, 1)
-  )
-}

### 2021-07-02 code/functions/utils.R

@@ -1,82 +1,0 @@
-my_correlate <- function(data, vector_name, placement) {
-  # my_correlate performs a correlation analysis of the selected variables.
-  #
-  # Args:
-  #   data: A data frame.
-  #   vector_name: A character string with the vector name. Either "resultant"
-  #   or "vertical".
-  #   placement: A character string with the name of the accelerometer
-  #   placement (either "ankle", "low back" or "hip").
-  #
-  # Returns:
-  #   A cor_df object.
-  msg <- paste("Accelerometer placement:", placement, "-- Vector:", vector_name)
-  corr <- data %>%
-    dplyr::filter(vector == vector_name & acc_placement == placement) %>%
-    dplyr::select(pGRF_N, pACC_g, body_mass, pLR_Ns, pATR_gs) %>%
-    corrr::correlate()
-  return(list(msg, corr))
-}
-
-filter_data <- function(data, vector_name, placement) {
-  # filter_data filters data by accelerometer placement and vector.
-  #
-  # Args:
-  #   data: A data frame.
-  #   vector_name: A character string with the vector name. Either "resultant"
-  #   or "vertical".
-  #   placement: A character string with the name of the accelerometer
-  #   placement (either "ankle", "low_back" or "hip").
-  #
-  # Returns:
-  #   A data frame.
-  dplyr::filter(data, vector == vector_name, acc_placement == placement)
-}
-
-prepare_data <- function(data) {
-  # prepare_data prepares data to test differences between actual and predicted
-  # values among accelerometer placements and activity types.
-  #
-  # Args:
-  #   data: A data frame.
-  #
-  # Return:
-  #   A tibble.
-  data %>%
-    tidyr::pivot_longer(
-      cols = c(actual, predicted),
-      names_to = "value_type",
-      values_to = "value"
-    ) %>%
-    dplyr::mutate(
-      value_type = dplyr::recode(
-        value_type,
-        "predicted" = paste0(acc_placement, "_", vector),
-        "actual" = paste0("actual_", vector)
-      )
-    )
-}
-
-bind_data <- function(list) {
-  # bind_data combines a list of data frames by row.
-  #
-  # Args:
-  #   list: A list of data frames.
-  #
-  # Returns:
-  #   A data frame.
-  purrr::map_dfr(list, rbind)
-}
-
-write_data <- function(data, name, path) {
-  # write_data writes a data frame into a csv file with the specified filename
-  # and path.
-  #
[…]

### 2021-07-13 code/01_tidy_data.R
Reorder jump factor levels

@@ -24,6 +24,9 @@ mechanical_load_data <- read_csv(here("data", "jumping_data.csv")) |>
       str_to_sentence(paste0(jump_type, " ", jump_height, "cm"))
     ),
     jump = fct_relevel(jump, "Drop jumps 40cm", after = 7),
+    jump = fct_relevel(
+      jump, "Continuous jumps 5cm", "Continuous jumps 15cm", after = Inf
+    ),
     .after = jump_height
   )
 

### 2021-07-13 figures/boxplot.R
Update figure

@@ -2,269 +2,123 @@
 
 library(here)
 library(tidyverse)
-library(ggpattern)
 library(ragg)
 library(patchwork)
 
 # Load data ---------------------------------------------------------------
 
-running_df <- read_csv(here("data/processed/running_data.csv")) %>%
+load(here("data", "mechanical_load_data.rda"))
+mechanical_load_data <- mechanical_load_data |>
   mutate(
-    run = as_factor(paste0("running ", speed, "km/h")),
-  ) %>%
-  select(
-    subj, acc_placement, vector, run, pGRF_BW, pACC_g, pLR_BWs, pATR_gs
-  )
-jumping_df <- read_csv(here("data/processed/jumping_data.csv")) %>%
-  mutate(
-    jump = as_factor(paste0(jump_type, " ", jump_height, " cm"))
-  ) %>%
-  select(
-    subj, acc_placement, vector, jump, pGRF_BW, pACC_g, pLR_BWs, pATR_gs
-  )
-
-data <- running_df %>%
-  full_join(jumping_df) %>%
-  mutate(
-    acc_placement = recode(
-      as_factor(acc_placement),
-      "ankle" = "Ankle",
-      "lower_back" = "Lower Back",
-      "hip" = "Hip"
-    ),
-    activity = ifelse(
-      is.na(as.character(jump)), as.character(run), as.character(jump)
-    ),
-    activity = str_to_sentence(activity),
-    activity = as_factor(activity),
-    activity = fct_relevel(activity, "Drop jumps 40cm", after = 15)
-  )
-
-# Set x axis labels
-x_labs <- scale_x_discrete(
-    labels = c(
-      "Running 7km/h" = quote("Running 7" ~ km %.% h^-1),
-      "Running 8km/h" = quote("Running 8" ~ km %.% h^-1),
-      "Running 9km/h" = quote("Running 9" ~ km %.% h^-1),
-      "Running 10km/h" = quote("Running 10" ~ km %.% h^-1),
-      "Running 11km/h" = quote("Running 11" ~ km %.% h^-1),
-      "Running 12km/h" = quote("Running 12" ~ km %.% h^-1),
-      "Running 13km/h" = quote("Running 13" ~ km %.% h^-1),
-      "Running 14km/h" = quote("Running 14" ~ km %.% h^-1)
+    # acc_placement = fct_relevel(
[…]

### 2021-07-13 figures/fig1.R
Delete commented code

@@ -10,9 +10,6 @@ library(patchwork)
 load(here("data", "mechanical_load_data.rda"))
 mechanical_load_data <- mechanical_load_data |>
   mutate(
-    # acc_placement = fct_relevel(
-    #   acc_placement, "hip", "lower_back", "ankle"
-    # )
     acc_placement = recode_factor(
       acc_placement,
       hip = "Hip",

### 2021-07-13 figures/fig2.R
Initial commit

@@ -1,0 +1,100 @@
+# Load packages -----------------------------------------------------------
+
+library(here)
+library(tidyverse)
+library(lvmisc)
+library(ggsci)
+library(patchwork)
+library(ragg)
+
+# Load data ---------------------------------------------------------------
+
+load(here("output", "loocv_data.rda"))
+
+# Hip pGRF Bland-Altman plot ---------------------------------------------
+
+BA_GRF_res_hip <- cv_res_GRF_models$hip |>
+  plot_bland_altman(color = BMI_cat, alpha = 0.5) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    limits = c(-1500, 1500),
+    expand = c(0, 0),
+    breaks = seq(-1500, 1500, 500)
+  ) +
+  scale_x_continuous(
+    limits = c(1000, 6100),
+    expand = c(0, 0),
+    breaks = seq(1500, 6000, 500)
+  ) +
+  theme_light() +
+  theme(
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 12),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
+  ) +
+  guides(alpha = "none") +
+  labs(
+    title = "Resultant vector - Hip placement",
+    x = "Mean of Actual and Predicted pGRF (N)",
+    y = "Actual - Predicted pGRF (N)"
+  )
+
+# Hip pLR Bland-Altman plot ----------------------------------------------
+
+BA_LR_res_hip <- cv_res_LR_models$hip |>
+  plot_bland_altman(color = BMI_cat, alpha = 0.5) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    labels = scales::label_number(),
+    limits = c(-80000, 80000),
+    expand = c(0, 0),
+    breaks = seq(-80000, 80000, 40000)
+  ) +
+  scale_x_continuous(
+    labels = scales::label_number(),
+    limits = c(0, 270000),
+    expand = c(0, 0),
+    breaks = seq(50000, 250000, 50000)
+  ) +
+  theme_light() +
+  theme(
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 12),
[…]

### 2021-07-13 figures/figS1.R

@@ -1,0 +1,121 @@
+# Load packages and functions ---------------------------------------------
+
+library(here)
+library(tidyverse)
+library(ragg)
+library(patchwork)
+
+# Load data ---------------------------------------------------------------
+
+load(here("data", "mechanical_load_data.rda"))
+mechanical_load_data <- mechanical_load_data |>
+  mutate(
+    acc_placement = recode_factor(
+      acc_placement,
+      hip = "Hip",
+      lower_back = "Lower back",
+      ankle = "Ankle"
+    )
+  )
+
+# GRF boxplot -------------------------------------------------------------
+
+boxplot_GRF_ver <- mechanical_load_data |>
+  filter(acc_placement == "Lower back" & vector == "vertical") |>
+  ggplot(aes(x = jump, y = pGRF_BW)) +
+  geom_boxplot() +
+  scale_y_continuous(
+    limits = c(0, 7),
+    expand = c(0, 0),
+    breaks = seq(0, 7, 1)
+  ) +
+  theme_light() +
+  theme(
+    axis.text.y = element_text(size = 13),
+    axis.title.y = element_text(size = 16)
+  ) +
+  labs(x = "", y = "pGRF (BW)")
+
+# ACC boxplots ------------------------------------------------------------
+
+boxplot_ACC_ver <- mechanical_load_data |>
+  filter(vector == "vertical") |>
+  ggplot(aes(x = jump, y = pACC_g)) +
+  geom_boxplot(aes(fill = acc_placement), outlier.size = 0.8) +
+  scale_fill_manual(values = c("gray90", "gray70", "gray50")) +
+  scale_y_continuous(
+    limits = c(0, 14),
+    expand = c(0, 0),
+    breaks = seq(0, 14, 2)
+  ) +
+  theme_light() +
+  theme(
+    axis.text.y = element_text(size = 13),
+    axis.title.y = element_text(size = 16),
+    legend.title = element_blank()
+  ) +
+  labs(x = "", y = quote("pACC" ~ (italic(g))))
+
+# LR boxplot --------------------------------------------------------------
+
+boxplot_LR_ver <- mechanical_load_data |>
+  filter(acc_placement == "Lower back" & vector == "vertical") |>
+  ggplot(aes(x = jump, y = pLR_BWs)) +
+  geom_boxplot() +
+  scale_y_continuous(
+    limits = c(0, 350),
[…]

### 2021-07-13 figures/figS2.R

@@ -1,0 +1,230 @@
+# Load packages -----------------------------------------------------------
+
+library(here)
+library(tidyverse)
+library(lvmisc)
+library(ggsci)
+library(patchwork)
+library(ragg)
+
+# Load data ---------------------------------------------------------------
+
+load(here("output", "loocv_data.rda"))
+
+# Ankle resultant pGRF Bland-Altman plot ----------------------------------
+
+BA_GRF_res_ankle <- cv_res_GRF_models$ankle |>
+  plot_bland_altman(color = BMI_cat, alpha = 0.5) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    limits = c(-2000, 2000),
+    expand = c(0, 0),
+    breaks = seq(-2000, 2000, 500)
+  ) +
+  scale_x_continuous(
+    limits = c(1000, 5600),
+    expand = c(0, 0),
+    breaks = seq(1500, 5500, 500)
+  ) +
+  theme_light() +
+  theme(
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 12),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
+  ) +
+  guides(alpha = "none") +
+  labs(
+    title = "Resultant vector - Ankle placement",
+    x = "Mean of Actual and Predicted pGRF (N)",
+    y = "Actual - Predicted pGRF (N)"
+  )
+
+# Lower back resultant pGRF Bland-Altman plot -----------------------------
+
+BA_GRF_res_back <- cv_res_GRF_models$lower_back |>
+  plot_bland_altman(color = BMI_cat, alpha = 0.5) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    limits = c(-1600, 1600),
+    expand = c(0, 0),
+    breaks = seq(-1600, 1600, 400)
+  ) +
+  scale_x_continuous(
+    limits = c(1000, 5600),
+    expand = c(0, 0),
+    breaks = seq(1500, 5500, 500)
+  ) +
+  theme_light() +
+  theme(
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 12),
[…]

### 2021-07-13 figures/bland-altman.R
Delete files

@@ -1,478 +1,0 @@
-# Load packages and functions ---------------------------------------------
-
-library(here)
-library(tidyverse)
-library(ragg)
-library(patchwork)
-library(ggsci)
-library(scales)
-source(here("code/functions/plot_functions.R"))
-
-# Load data ---------------------------------------------------------------
-
-running_df <- read_csv(here("data/processed/running_data.csv")) %>%
-  mutate(
-    BMI_cat = fct_relevel(
-      as_factor(BMI_cat),
-      levels = c(
-        "Normal weight", "Overweight", "Obesity class I"
-      )
-    )
-  ) %>%
-  select(subj, BMI_cat)
-
-jumping_df <- read_csv(here("data/processed/jumping_data.csv")) %>%
-  mutate(
-    BMI_cat = fct_relevel(
-      as_factor(BMI_cat),
-      levels = c(
-        "Normal weight", "Overweight", "Obesity class I",
-        "Obesity class II", "Obesity class III"
-      )
-    )
-  ) %>%
-  select(subj, BMI_cat)
-
-# LOOCV data
-GRF_loocv_running <- read_csv(here("data/processed/GRF_loocv_running.csv")) %>%
-  full_join(running_df)
-GRF_loocv_jumping <- read_csv(here("data/processed/GRF_loocv_jumping.csv")) %>%
-  full_join(jumping_df)
-LR_loocv_jumping <- read_csv(here("data/processed/LR_loocv_jumping.csv")) %>%
-  full_join(jumping_df)
-
-# GRF plots ---------------------------------------------------------------
-
-# Running
-BA_GRF_running_ankle_res <- GRF_loocv_running %>%
-  filter(acc_placement == "ankle" & vector == "resultant") %>%
-  publication_bland_altman() +
-  scale_color_nejm() +
-  scale_y_continuous(
-    limits = c(-1500, 600),
-    expand = c(0, 0),
-    breaks = seq(-1500, 600, 300)
-  ) +
-  theme_light() +
-  theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
-  ) +
-  labs(
-    title = "Ankle",
-    x = "Mean of Actual and Predicted pRGRF (N)",
-    y = "Actual - Predicted pRGRF (N)"
-  )
-
[…]

### 2021-07-13 figures/scatterplot.R

@@ -1,605 +1,0 @@
-# Load packages and functions ---------------------------------------------
-
-library(here)
-library(tidyverse)
-library(ragg)
-library(patchwork)
-library(ggsci)
-
-# Load data ---------------------------------------------------------------
-
-running_df <- read_csv(here("data/processed/running_data.csv")) %>%
-  mutate(
-    BMI_cat = fct_relevel(
-      as_factor(BMI_cat),
-      levels = c(
-        "Normal weight", "Overweight", "Obesity class I"
-      )
-    )
-  )
-
-jumping_df <- read_csv(here("data/processed/jumping_data.csv")) %>%
-  mutate(
-    BMI_cat = fct_relevel(
-      as_factor(BMI_cat),
-      levels = c(
-        "Normal weight", "Overweight", "Obesity class I",
-        "Obesity class II", "Obesity class III"
-      )
-    )
-  )
-
-# GRF x ACC plots ---------------------------------------------------------
-
-# Running
-scatter_ACC_running_ankle_res <- running_df %>%
-  filter(acc_placement == "ankle" & vector == "resultant") %>%
-  ggplot(aes(x = pACC_g, y = pGRF_N, color = BMI_cat)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  scale_color_nejm() +
-  scale_y_continuous(
-    limits = c(0, 3500),
-    expand = c(0, 0),
-    breaks = seq(0, 3500, 500)
-  ) +
-  theme_light() +
-  theme(
-    plot.title = element_text(hjust = 0.5),
-    legend.title = element_blank()
-  ) +
-  labs(title = "Ankle", x = quote("pRACC" ~ (italic(g))), y = "pRGRF (N)")
-
-scatter_ACC_running_back_res <- running_df %>%
-  filter(acc_placement == "lower_back" & vector == "resultant") %>%
-  ggplot(aes(x = pACC_g, y = pGRF_N, color = BMI_cat)) +
-  geom_point() +
-  geom_smooth(method = "lm", se = FALSE) +
-  scale_color_nejm() +
-  scale_y_continuous(
-    limits = c(0, 3500),
-    expand = c(0, 0),
[…]

### 2021-07-13 code/funs.R
Initial commit

@@ -1,0 +1,142 @@
+# Build regression formula table
+#
+# Params:
+#    model: A list of length 3 with the lmerMod model objects.
+#    cv: A list of length 3 with the lvmisc_cv objects.
+#    outcome: A character vector with the model outcome. Either "pGRF"
+# or "pLR".
+#    vector: A character vector with the vector used in the model. Either
+# "resultant" or "vertical".
+#
+# Returns:
+#    A tibble with the regression formula table.
+build_formula_table <- function(model, cv, outcome, vector) {
+  model_accuracy <- purrr::map(
+    cv, lvmisc::accuracy, na.rm = TRUE
+  )
+  R2 <- unname(purrr::map_dbl(model_accuracy, "R2_cond"))
+  MAE <- broman::myround(unname(purrr::map_dbl(model_accuracy, "MAE")), 1)
+  MAPE <- paste0(
+    broman::myround(
+      unname(purrr::map_dbl(model_accuracy, "MAPE")) * 100, 1
+    ),
+    "\\%"
+  )
+  RMSE <- broman::myround(unname(purrr::map_dbl(model_accuracy, "RMSE")), 1)
+
+  tibble::tibble(
+    "Vector" = stringr::str_to_sentence(vector),
+    "Accelerometer placement" = c("Ankle", "Lower back", "Hip"),
+    "Regression equations" = unname(
+      purrr::map_chr(model, get_equation, outcome)
+    ),
+    "R2" = R2, "MAE" = MAE, "MAPE" = MAPE, "RMSE" = RMSE
+  )
+}
+
+# Get regression equation formula with coefficients and variables
+#
+# Params:
+#    model: The lmerMod model object.
+#    outcome: A character vector with the model outcome. Either "pGRF"
+# or "pLR".
+#    vector: A character vector with the vector used in the model. Either
+# "resultant" or "vertical".
+#
+# Returns:
+#    A character vector with the model equation.
+get_equation <- function(model, outcome) {
+  model_coefs <- coef(summary(model))[, 1]
+  model_coefs[1] <- ifelse(
+    model_coefs[1] > 0,
+    broman::myround(model_coefs[1], 3),
+    paste0("- ", broman::myround(abs(as.numeric(model_coefs[1])), 3))
+  )
+  model_coefs[-1] <- ifelse(
+    model_coefs[-1] > 0,
+    paste0(" + ", broman::myround(abs(as.numeric(model_coefs[-1])), 3)),
[…]

### 2021-07-13 tables/tab1.R
Initial commit

@@ -1,0 +1,75 @@
+# Load packages and functions ---------------------------------------------
+
+library(here)
+library(tidyverse)
+library(broman)
+library(kableExtra)
+source(here("code", "funs.R"))
+
+# Load and prepare data ---------------------------------------------------
+
+load(here("output", "prediction_models.rda"))
+load(here("output", "loocv_data.rda"))
+
+# Build table -------------------------------------------------------------
+
+res_GRF_tb <- build_formula_table(
+  res_GRF_models, cv_res_GRF_models, "pGRF", "resultant"
+)
+ver_GRF_tb <- build_formula_table(
+  ver_GRF_models, cv_ver_GRF_models, "pGRF", "vertical"
+)
+res_LR_tb <- build_formula_table(
+  res_LR_models, cv_res_LR_models, "pLR", "resultant"
+) %>%
+  mutate(MAE = round(as.numeric(MAE), 0), RMSE = round(as.numeric(RMSE), 0))
+ver_LR_tb <- build_formula_table(
+  ver_LR_models, cv_ver_LR_models, "pLR", "vertical"
+) %>%
+  mutate(MAE = round(as.numeric(MAE), 0), RMSE = round(as.numeric(RMSE), 0))
+
+tab1_df <- rbind(res_GRF_tb, ver_GRF_tb, res_LR_tb, ver_LR_tb)
+# Remove vector name duplicates
+tab1_df[c(2, 3, 5, 6, 8, 9, 11, 12), 1] <- ""
+
+tab1 <- knitr::kable(
+  tab1_df, booktabs = TRUE, escape = FALSE, linesep = "",
+  label = "none",
+  caption = "Regression equations, $R^2$ and accuracy indices",
+  col.names = c(
+    "Vector",
+    linebreak("Accelerometer\nplacement"),
+    "Regression equations",
+    "$R^2$",
+    "MAE",
+    "MAPE",
+    "RMSE"
+  ),
+  align = c(rep("l", 3), rep("r", 4))
+) %>%
+  footnote(
+    general = "Abbreviations: MAE, mean absolute error; MAPE, mean absolute percent error; pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate; RMSE, root mean square error",
+    general_title = "",
+    threeparttable = TRUE
+  ) %>%
+  kable_styling(latex_options = "scale_down") %>%
+  pack_rows(
[…]

### 2021-07-13 code/04_analyse_results.R
Initial commit

@@ -1,0 +1,49 @@
+# Load packages and functions ---------------------------------------------
+
+library(here)
+library(tidyverse)
+library(lvmisc)
+source(here("code", "funs.R"))
+
+# Load and prepare data ---------------------------------------------------
+
+load(here("output", "loocv_data.rda"))
+load(here("data", "mechanical_load_data.rda"))
+
+# Bland-Altman plots tests ------------------------------------------------
+
+# Linear regression diff ~ mean
+ba_regression_res_GRF <- map(cv_res_GRF_models, bland_altman_regression)
+ba_regression_ver_GRF <- map(cv_ver_GRF_models, bland_altman_regression)
+ba_regression_res_LR <- map(cv_res_LR_models, bland_altman_regression)
+ba_regression_ver_LR <- map(cv_ver_LR_models, bland_altman_regression)
+
+# Check if bias differs from 0
+ba_bias_test_res_GRF <- map(cv_res_GRF_models, bland_altman_t_test)
+ba_bias_test_ver_GRF <- map(cv_ver_GRF_models, bland_altman_t_test)
+ba_bias_test_res_LR <- map(cv_res_LR_models, bland_altman_t_test)
+ba_bias_test_ver_LR <- map(cv_ver_LR_models, bland_altman_t_test)
+
+# Accuracy per activity ---------------------------------------------------
+
+# Drop jumps
+drop_jumps_res_accuracy <- cv_res_GRF_models %>%
+  map(~ filter(.x, jump_type == "drop jumps")) %>%
+  map(accuracy, na.rm = TRUE)
+drop_jumps_ver_accuracy <- cv_ver_GRF_models %>%
+  map(~ filter(.x, jump_type == "drop jumps")) %>%
+  map(accuracy, na.rm = TRUE)
+# Box jumps
+box_jumps_res_accuracy <- cv_res_GRF_models %>%
+  map(~ filter(.x, jump_type == "box jumps")) %>%
+  map(accuracy, na.rm = TRUE)
+box_jumps_ver_accuracy <- cv_ver_GRF_models %>%
+  map(~ filter(.x, jump_type == "box jumps")) %>%
+  map(accuracy, na.rm = TRUE)
+# Continuous jumps
+continuous_jumps_res_accuracy <- cv_res_GRF_models %>%
+  map(~ filter(.x, jump_type == "continuous jumps")) %>%
+  map(accuracy, na.rm = TRUE)
[…]

### 2021-07-13 code/04_analyse_results.R
Add LR accuracy per jump type and save the objects

@@ -8,7 +8,6 @@ source(here("code", "funs.R"))
 # Load and prepare data ---------------------------------------------------
 
 load(here("output", "loocv_data.rda"))
-load(here("data", "mechanical_load_data.rda"))
 
 # Bland-Altman plots tests ------------------------------------------------
 
@@ -27,23 +26,56 @@ ba_bias_test_ver_LR <- map(cv_ver_LR_models, bland_altman_t_test)
 # Accuracy per activity ---------------------------------------------------
 
 # Drop jumps
-drop_jumps_res_accuracy <- cv_res_GRF_models %>%
+drop_jumps_res_GRF_accuracy <- cv_res_GRF_models %>%
   map(~ filter(.x, jump_type == "drop jumps")) %>%
   map(accuracy, na.rm = TRUE)
-drop_jumps_ver_accuracy <- cv_ver_GRF_models %>%
+drop_jumps_ver_GRF_accuracy <- cv_ver_GRF_models %>%
+  map(~ filter(.x, jump_type == "drop jumps")) %>%
+  map(accuracy, na.rm = TRUE)
+drop_jumps_res_LR_accuracy <- cv_res_LR_models %>%
+  map(~ filter(.x, jump_type == "drop jumps")) %>%
+  map(accuracy, na.rm = TRUE)
+drop_jumps_ver_LR_accuracy <- cv_ver_LR_models %>%
   map(~ filter(.x, jump_type == "drop jumps")) %>%
   map(accuracy, na.rm = TRUE)
 # Box jumps
-box_jumps_res_accuracy <- cv_res_GRF_models %>%
+box_jumps_res_GRF_accuracy <- cv_res_GRF_models %>%
   map(~ filter(.x, jump_type == "box jumps")) %>%
   map(accuracy, na.rm = TRUE)
-box_jumps_ver_accuracy <- cv_ver_GRF_models %>%
+box_jumps_ver_GRF_accuracy <- cv_ver_GRF_models %>%
+  map(~ filter(.x, jump_type == "box jumps")) %>%
+  map(accuracy, na.rm = TRUE)
+box_jumps_res_LR_accuracy <- cv_res_LR_models %>%
+  map(~ filter(.x, jump_type == "box jumps")) %>%
+  map(accuracy, na.rm = TRUE)
+box_jumps_ver_LR_accuracy <- cv_ver_LR_models %>%
   map(~ filter(.x, jump_type == "box jumps")) %>%
[…]

### 2021-07-13 code/04_analyse_results.R
Correct object names to save

@@ -71,11 +71,11 @@ if (!dir.exists(here("output"))) {
   dir.create(here("output"))
 }
 save(
-  drop_jumps_ver_GRF_accuracy, drop_jumps_ver_GRF_accuracy,
+  drop_jumps_res_GRF_accuracy, drop_jumps_ver_GRF_accuracy,
   drop_jumps_res_LR_accuracy, drop_jumps_ver_LR_accuracy,
-  box_jumps_ver_GRF_accuracy, box_jumps_ver_GRF_accuracy,
+  box_jumps_res_GRF_accuracy, box_jumps_ver_GRF_accuracy,
   box_jumps_res_LR_accuracy, box_jumps_ver_LR_accuracy,
-  continuous_jumps_ver_GRF_accuracy, continuous_jumps_ver_GRF_accuracy,
+  continuous_jumps_res_GRF_accuracy, continuous_jumps_ver_GRF_accuracy,
   continuous_jumps_res_LR_accuracy, continuous_jumps_ver_LR_accuracy,
   file = here("output", "jump_type_accuracy.rda")
 )

### 2021-07-13 code/funs.R
Add build_accuracy_table()

@@ -34,6 +34,30 @@ build_formula_table <- function(model, cv, outcome, vector) {
   )
 }
 
+# Build table with accuracy indices per jump type
+#
+# Params:
+#    accuracy: A list of length 3 with the lvmisc_accuracy objects.
+#    vector: A character vector with the vector used in the model. Either
+# "resultant" or "vertical".
+#    jump_type: A character vector with the jump type. Either "drop", "box" or
+# "continuous".
+#
+# Returns:
+#    A tibble with the table with accuracy indices per jump type.
+build_accuracy_table <- function(accuracy, vector, jump_type) {
+  data <- tibble::tibble(
+    Vector = rep(stringr::str_to_sentence(vector), 3),
+    `Accelerometer placement` = c("Ankle", "Lower back", "Hip"),
+    R2 = broman::myround(unname(purrr::map_dbl(accuracy, "R2_cond")), 2),
+    MAE = broman::myround(unname(purrr::map_dbl(accuracy, "MAE"))),
+    MAPE = broman::myround(unname(purrr::map_dbl(accuracy, "MAPE")) * 100),
+    RMSE = broman::myround(unname(purrr::map_dbl(accuracy, "RMSE"))),
+  )
+  names(data)[3:6] <- paste0(names(data)[3:6], "_", jump_type)
+  data
+}
+
 # Get regression equation formula with coefficients and variables
 #
 # Params:

### 2021-07-13 tables/tab2.R
Initial commit

@@ -1,0 +1,73 @@
+# Load packages and functions ---------------------------------------------
+
+library(here)
+library(tidyverse)
+library(kableExtra)
+source(here("code", "funs.R"))
+
+# Load and prepare data ---------------------------------------------------
+
+load(here("output", "jump_type_accuracy.rda"))
+
+# Build table -------------------------------------------------------------
+
+# GRF table
+res_GRF_tab2_df <- build_accuracy_table(
+  drop_jumps_res_GRF_accuracy, "resultant", "drop"
+) |>
+  left_join(
+    build_accuracy_table(box_jumps_res_GRF_accuracy, "resultant", "box")
+  ) |>
+  left_join(
+    build_accuracy_table(
+      continuous_jumps_res_GRF_accuracy, "resultant", "continuous"
+    )
+  )
+ver_GRF_tab2_df <- build_accuracy_table(
+  drop_jumps_ver_GRF_accuracy, "vertical", "drop"
+) |>
+  left_join(
+    build_accuracy_table(box_jumps_ver_GRF_accuracy, "vertical", "box")
+  ) |>
+  left_join(
+    build_accuracy_table(
+      continuous_jumps_ver_GRF_accuracy, "vertical", "continuous"
+    )
+  )
+GRF_tab2_df <- rbind(res_GRF_tab2_df, ver_GRF_tab2_df)
+# LR table
+res_LR_tab2_df <- build_accuracy_table(
+  drop_jumps_res_LR_accuracy, "resultant", "drop"
+) |>
+  left_join(
+    build_accuracy_table(box_jumps_res_LR_accuracy, "resultant", "box")
+  ) |>
+  left_join(
+    build_accuracy_table(
+      continuous_jumps_res_LR_accuracy, "resultant", "continuous"
+    )
+  )
+ver_LR_tab2_df <- build_accuracy_table(
+  drop_jumps_ver_LR_accuracy, "vertical", "drop"
+) |>
+  left_join(
+    build_accuracy_table(box_jumps_ver_LR_accuracy, "vertical", "box")
+  ) |>
+  left_join(
+    build_accuracy_table(
+      continuous_jumps_ver_LR_accuracy, "vertical", "continuous"
+    )
+  )
+LR_tab2_df <- rbind(res_LR_tab2_df, ver_LR_tab2_df)
+# Join both tables
+tab2_df <- rbind(GRF_tab2_df, LR_tab2_df)
+# Remove vector name duplicates
[…]

### 2021-07-14 manuscript/manuscript.Rmd
Initial commit

---
# title:
output:
  bookdown::pdf_document2:
    toc: false
    includes:
      in_header: "preamble.tex"
    latex_engine: lualatex
  word_document:
    reference_docx: style_reference.docx
always_allow_html: true
# bibliography: bibliography.bib
csl: nlm.csl
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, results = "hide", message = FALSE, warning = FALSE
)
```

```{r source}
library(here)
library(bookdown)
source(here("code", "02_explore.R"))
```

<!--
<center>
Lucas Veras$^{1,2}$, Florêncio Diniz-Sousa$^{1,2}$, Giorjines Boppre$^{1,2}$, Vítor Devezas$^3$, Hugo Santos-Sousa$^3$, John Preto$^3$, João Paulo Vilas-Boas$^{4,5}$, Leandro Machado$^{4,5}$, José Oliveira$^{1,2}$, Hélder Fonseca$^{1,2}$
</center>

\bigskip

1 - Research Center in Physical Activity, Health and Leisure (CIAFEL), Faculty of Sport, University of Porto, Porto, Portugal

2 - Laboratory for Integrative and Translational Research in Population Health (ITR), University of Porto, Porto, Portugal

3 - General Surgery Department, São João Medical Center, Porto, Portugal

4 - Center of Research, Education, Innovation and Intervention in Sport (CIFI2D), Faculty of Sport, University of Porto, Porto, Portugal

5 - Biomechanics Laboratory (LABIOMEP-UP), University of Porto, Porto, Portugal

\leftskip=0.5in \rightskip=0.5in

\bigskip
-->

# Methods {.unnumbered}

The study protocol was approved by the local Ethics Committee (CES 192-14), and all participants were informed about the experiments purpose and protocol before giving written informed consent. A total of `r max(sample_size$pGRF_N)` adults were enrolled (`r unname(sample_size_sex["M"])` males; `r sample_descriptives$age_mean`±`r sample_descriptives$age_sd` years; `r sample_descriptives$height_mean`±`r sample_descriptives$height_sd` cm; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg; `r sample_descriptives$BMI_mean`±`r sample_descriptives$BMI_sd` kg·m^-2^, X̄±SD). No participant reported any neurological or musculoskeletal limitations. Height and body mass were assessed following standard procedures with a stadiometer (model 213; Seca, Hamburg, Germany) and a digital scale (model 899; Seca, Hamburg, Germany), respectively.

### 2021-10-26 manuscript/manuscript.Rmd
Continue Methods section

@@
 Height and body mass were assessed following standard procedures with a stadiometer (model 213; Seca, Hamburg, Germany) and a digital scale (model 899; Seca, Hamburg, Germany), respectively.
+
+After familiarization, participants performed a series of trials envolving jumps of different types and heights.
+First, drop jumps were executed from steps varying from 5cm to 40cm high, with 5cm increments.
+Then, box jumps were performed from the floor to 5cm, 15cm and 30cm high boxes.
+Finally, continuous jumps were carried out in two different heights: 5cm and 15cm.
+Drop and box jumps were realized in two sets of four jumps at each height, with 30s interval, and the continuous jumps in one set of 20s at each height.
+The landing phase of the drop jumps, the take-off phase of the box jumps and both phases of the continuous jumps were recorded in force plates (1000Hz; AMTI Corporation, Watertown, USA).
+Participants were instructed to jump with no restrictions to arms movement, land two-footed and then remain stationary on the force plate for five seconds (except in the continuous jumps).
+
+Throughout the protocol, three activity monitors (GT9X Link; 100Hz; ±16g range; ActiGraph, Pensacola, USA) were worn by each participant, at the following placements: i) at their right hip (along the anterior axillary line, at the level of the iliac crest); ii) at their lower back (at the midpoint between the two posterior superior iliac spines); and iii) at their right ankle (immediately superior to the lateral malleolus).
+Activity monitors at waist level were secured on the same elastic belt with a clip and the activity monitor placed on the ankle was secured by an elastic belt fixed with adhesive tape.
+In all placements, the accelerometer vertical axis was aligned with the standing body longitudinal axis.
+
+The motion sensors used incorporate primary and secondary triaxial accelerometers.
+As the manufacturer proprietary filter is not applied on the secondary accelerometer raw data, to improve the data processing replicability only data from that accelerometer were used in this study.
+Data from FP and accelerometers were operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA) and exported as raw data from the *x*, *y* and *z* vectors.
+GRF was expressed in N and acceleration (ACC) in gravitational ACC units (1*g* = 9.807 m·s^-2^).
+
+FP and accelerometer data were processed using MATLAB (Version 2019a, Mathworks, Natick, USA) according to the following procedures: first, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, to attenuate the noise magnitude.
+This cut-off frequency was selected based on visual inspection of the signals frequency spectrum after a Fast Fourier Transform (FFT).
+Afterwards, the GRF and ACC resultant vector was calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$)

### 2021-10-26 manuscript/manuscript.Rmd
Finish data processing

@@
 
-FP and accelerometer data were processed using MATLAB (Version 2019a, Mathworks, Natick, USA) according to the following procedures: first, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, to attenuate the noise magnitude.
+FP and accelerometer data were processed using MATLAB (Version 2019a, Mathworks, Natick, USA) according to the following procedures: first, the GRF signal was resampled to match the activity monitors sampling frequency (100Hz).
+Then, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, to attenuate the noise magnitude.
 This cut-off frequency was selected based on visual inspection of the signals frequency spectrum after a Fast Fourier Transform (FFT).
-Afterwards, the GRF and ACC resultant vector was calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$)
+Afterwards, the GRF and ACC resultant vector was calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$).
+After that, both signals were adjusted using the time set by the systems clock and then synchronized by the maximum cross-correlation coefficient, with posterior manual adjustments if necessary.
+Next, peak ACC (pACC) were determined by the following criteria: i) a minimum height of three standard deviations above the mean ACC value recorded during a trial; and ii) a separation of at least 0.2 seconds for continuous jumps and 4 seconds for the other jumps.
+Peak GRF (pGRF) were then defined as the highest value between before and after 0.2 or 4 seconds, according to the jump type, after each pACC.
+The rates of change were computed through a centered derivative from the beginning of the foot contact to the curve peak (Equation 1).
+The peak loading rate (pLR) and peak acceleration rate (pAR) were defined as the maximum value in the array (Equation 2).
+
+\[f_i^{'} = \frac{f_{i + 1} - f_{i - 1}}{t_{i + 1} - t_{i - 1}}\]
+
+\[f_{i_{max}}^{'} = max(f_i^{'})\]
+
+Finally, pGRF, pACC, pLR and pAR means of the resultant and its vertical component for each participant at each jump type and height for each accelerometer placement were extracted and used in all remaining analyses.

### 2021-10-26 code/03_build_models.R
Update models formula

@@ -28,7 +28,7 @@ ver_mechanical_load_data <- map(
 
 GRF_formula <- as.formula(
   "pGRF_N ~ pACC_g + body_mass + pACC_g:body_mass +
-    (1 | subj) + (1 | jump_height)"
+    (1 | subj) + (1 | jump_type : jump_height)"
 )
 
 # Resultant vector
@@ -71,7 +71,7 @@ plot_ver_GRF_models <- map(
 
 LR_formula <- as.formula(
   "pLR_Ns ~ pAR_gs + body_mass + pAR_gs:body_mass +
-    (1 | subj) + (1 | jump_height)"
+    (1 | subj) + (1 | jump_type : jump_height)"
 )
 
 # Resultant vector

### 2021-10-26 manuscript/manuscript.Rmd
Begin statistical analyses description

@@
 always_allow_html: true
-# bibliography: bibliography.bib
+bibliography: bibliography.bib
 csl: nlm.csl
@@
 Finally, pGRF, pACC, pLR and pAR means of the resultant and its vertical component for each participant at each jump type and height for each accelerometer placement were extracted and used in all remaining analyses.
+
+Statistical analyses were conducted using R statistical software (version `r paste(R.Version()$major, R.Version()$minor, sep = ".")`, R Foundation for Statistical Computing, Vienna, Austria).
+All statistical analyses were registered in an open platform, where R code utilized can be assessed (LINK).
+The statistically significant value was set as $\alpha$ = 0.05.
+
+Linear mixed models were used to develop pGRF and pLR prediction equations for each vector (resultant and vertical) and accelerometer placement (ankle, lower back and  hip).
+Body mass was tested as predictor in all models, along with the acceleration magnitude (pACC) in the pGRF models and its rate (pAR) in the pLR models.
+All predictors were tested as fixed effects and were shown to be significant.
+Subject and the interaction between jump type and height were tested as random effects, and both factors have shown to improve the models.
+Final models were chosen according to the -2 log-likelihood statistic [@Field2012].
+Traditional coefficient of determination (R^2^) was represented by conditional R^2^, which estimates the variance explained by the whole model [@Nakagawa2013].
+
+Model validation was performed using the leave-one-out cross-validation [@Staudenmayer2012].
+In this approach, each participant’s data is split into a testing dataset (one participant at a time), and the remaining data is in the training dataset.
+New linear mixed models with the same specification as determined for the whole sample are developed using the training dataset and then used to predict the outcome for the participant in the testing dataset.
+This process is, then, repeated for all participants (`r max(sample_size$pGRF_N)` times).
+Data from the testing dataset was used in the remaining statistical analyses.
+
+\newpage
+
+# References {.unnumbered}
+
+<div id="refs"></div>

### 2022-01-25 manuscript/manuscript.Rmd
Finish the methods section

@@
 
+The quality of the models was then evaluated both numerically and graphically.
+The numerical assessment was made by computing the following accuracy indices: the mean absolute error (MAE), mean absolute percent error (MAPE) and root mean square error (RMSE).
+To visualize the agreement between the values predicted by the models and measured by the FP, Bland-Altman plots were drawn.
+This type of plot is made by plotting the mean of the actual and predicted values on the x axis and their difference on the y axis.
+Then, the bias was calculated as the mean of these differences and plotted as a continuous horizontal line, and the limits of agreement were defined by bias ± 1.96 standard deviation and plotted as horizontal dashed lines [@Bland1986].
+To further test if there were any consistent under- or over-estimation of the values by the models, one-sample *t*-tests were done to check whether the bias were different than zero.
+Finally, linear regressions were employed to examine if the difference between actual and predicted values were affected by their magnitude, verifying, therefore, whether the prediction was constant throughout the magnitude range that was tested [@Giavarina2015].
+
 \newpage

### 2022-01-25 manuscript/manuscript.Rmd
Reword the methods section

@@
 
-The study protocol was approved by the local Ethics Committee (CES 192-14), and all participants were informed about the experiments purpose and protocol before giving written informed consent.
-A total of `r max(sample_size$pGRF_N)` adults were enrolled (`r unname(sample_size_sex["M"])` males; `r sample_descriptives$age_mean`±`r sample_descriptives$age_sd` years; `r sample_descriptives$height_mean`±`r sample_descriptives$height_sd` cm; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg; `r sample_descriptives$BMI_mean`±`r sample_descriptives$BMI_sd` kg·m^-2^, X̄±SD).
-No participant reported any neurological or musculoskeletal limitations.
-Height and body mass were assessed following standard procedures with a stadiometer (model 213; Seca, Hamburg, Germany) and a digital scale (model 899; Seca, Hamburg, Germany), respectively.
+A total of `r max(sample_size$pGRF_N)` adults were enrolled in the study (`r unname(sample_size_sex["M"])` males; `r sample_descriptives$age_mean`±`r sample_descriptives$age_sd` years; `r sample_descriptives$height_mean`±`r sample_descriptives$height_sd` cm; `r sample_descriptives$body_mass_mean`±`r sample_descriptives$body_mass_sd` kg; `r sample_descriptives$BMI_mean`±`r sample_descriptives$BMI_sd` kg·m^-2^, X̄±SD).
+The protocol was approved by the local Ethics Committee (CES 192-14), and all participants were informed about the experiments purpose and protocol before giving written informed consent.
+No participant reported any musculoskeletal limitations.
+The measurement of height (stadiometer model 213; Seca, Hamburg, Germany) and body mass (digital scale model 899; Seca, Hamburg, Germany) was accomplished according to standard practice.
 
@@
 Finally, continuous jumps were carried out in two different heights: 5cm and 15cm.
-Drop and box jumps were realized in two sets of four jumps at each height, with 30s interval, and the continuous jumps in one set of 20s at each height.
+Drop and box jumps were executed in two sets of four jumps at each height, with 30s interval, and the continuous jumps in one set of 20s at each height.
 The landing phase of the drop jumps, the take-off phase of the box jumps and both phases of the continuous jumps were recorded in force plates (1000Hz; AMTI Corporation, Watertown, USA).
@@
 Throughout the protocol, three activity monitors (GT9X Link; 100Hz; ±16g range; ActiGraph, Pensacola, USA) were worn by each participant, at the following placements: i) at their right hip (along the anterior axillary line, at the level of the iliac crest); ii) at their lower back (at the midpoint between the two posterior superior iliac spines); and iii) at their right ankle (immediately superior to the lateral malleolus).
-Activity monitors at waist level were secured on the same elastic belt with a clip and the activity monitor placed on the ankle was secured by an elastic belt fixed with adhesive tape.
-In all placements, the accelerometer vertical axis was aligned with the standing body longitudinal axis.
+Monitors worn at hip and lower back were positioned on the same elastic belt with clips, and the ankle-worn monitor was fixed by an elastic belt and adhesive tape.
+Activity monitors were always positioned to ensure the alignment of the accelerometer vertical axis and the standing body craniocaudal axis.
 
-The motion sensors used incorporate primary and secondary triaxial accelerometers.
-As the manufacturer proprietary filter is not applied on the secondary accelerometer raw data, to improve the data processing replicability only data from that accelerometer were used in this study.
+The activity monitors used include a primary and a secondary triaxial accelerometers.
+As the manufacturer proprietary filter is not applied on the secondary accelerometer raw data, only data from this accelerometer were used in this study, to improve data processing replicability.
 Data from FP and accelerometers were operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA) and exported as raw data from the *x*, *y* and *z* vectors.
@@
 
-FP and accelerometer data were processed using MATLAB (Version 2019a, Mathworks, Natick, USA) according to the following procedures: first, the GRF signal was resampled to match the activity monitors sampling frequency (100Hz).
+The MATLAB software (Version 2019a, Mathworks, Natick, USA) was used to process data from both FP and accelerometers.
+Data were processed according to the following procedures: first, the GRF signal was resampled to match the activity monitors sampling frequency (100Hz).
 Then, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, to attenuate the noise magnitude.
@@
 Next, peak ACC (pACC) were determined by the following criteria: i) a minimum height of three standard deviations above the mean ACC value recorded during a trial; and ii) a separation of at least 0.2 seconds for continuous jumps and 4 seconds for the other jumps.
-Peak GRF (pGRF) were then defined as the highest value between before and after 0.2 or 4 seconds, according to the jump type, after each pACC.
+Peak GRF (pGRF) were then defined as the highest value between 0.2 or 4 seconds before and after each pACC, according to the jump type.
 The rates of change were computed through a centered derivative from the beginning of the foot contact to the curve peak (Equation 1).
@@
 
-Statistical analyses were conducted using R statistical software (version `r paste(R.Version()$major, R.Version()$minor, sep = ".")`, R Foundation for Statistical Computing, Vienna, Austria).
-All statistical analyses were registered in an open platform, where R code utilized can be assessed (LINK).
-The statistically significant value was set as $\alpha$ = 0.05.
+Statistical analyses were performed through the R statistical software (version `r paste(R.Version()$major, R.Version()$minor, sep = ".")`, R Foundation for Statistical Computing, Vienna, Austria).
+All code used to generate the results can be found in an open platform (https://github.com/verasls/jumping_GRF_ACC)(ref).
+Statistical significance was set as $\alpha$ = 0.05.
 
@@
 Body mass was tested as predictor in all models, along with the acceleration magnitude (pACC) in the pGRF models and its rate (pAR) in the pLR models.
-All predictors were tested as fixed effects and were shown to be significant.
+These predictors were entered as fixed effects and kept in the model, as all have shown to be significant (*p* < 0.05).
 Subject and the interaction between jump type and height were tested as random effects, and both factors have shown to improve the models.
 Final models were chosen according to the -2 log-likelihood statistic [@Field2012].
-Traditional coefficient of determination (R^2^) was represented by conditional R^2^, which estimates the variance explained by the whole model [@Nakagawa2013].
+The conditional coefficient of determination (R^2^), which estimates the variance explained by both fixed and random factors [@Nakagawa2013], was computed.
 
 Model validation was performed using the leave-one-out cross-validation [@Staudenmayer2012].
-In this approach, each participant’s data is split into a testing dataset (one participant at a time), and the remaining data is in the training dataset.
-New linear mixed models with the same specification as determined for the whole sample are developed using the training dataset and then used to predict the outcome for the participant in the testing dataset.
-This process is, then, repeated for all participants (`r max(sample_size$pGRF_N)` times).
-Data from the testing dataset was used in the remaining statistical analyses.
+This method consists in separating data from one of the subjects into a testing dataset, and keeping the remainder in the training dataset.
+Then, a new model is developed using the training dataset, with the same parameters that were defined using the whole sample.
+This new model is then applied to predict the outcome for the subject in the testing dataset.
+The whole process is repeated once for each participant (`r max(sample_size$pGRF_N)` times).
+Data from the testing dataset were used in all subsequent analyses.
 

### 2022-01-25 figures/fig1.R
Correct the LR unit in the y axis text

@@ -73,7 +73,7 @@ boxplot_LR_res <- mechanical_load_data |>
     axis.text.y = element_text(size = 13),
     axis.title.y = element_text(size = 16)
   ) +
-  labs(x = "", y = quote("pLR" ~ (N %.% s^-1)))
+  labs(x = "", y = quote("pLR" ~ (BW %.% s^-1)))
 
 # AR boxplots -------------------------------------------------------------
 

### 2022-01-25 figures/figS1.R

@@ -73,7 +73,7 @@ boxplot_LR_ver <- mechanical_load_data |>
     axis.text.y = element_text(size = 13),
     axis.title.y = element_text(size = 16)
   ) +
-  labs(x = "", y = quote("pLR" ~ (N %.% s^-1)))
+  labs(x = "", y = quote("pLR" ~ (BW %.% s^-1)))
 
 # AR boxplots -------------------------------------------------------------
 

### 2022-01-25 figures/fig1.R
Reorder plots in the panels

@@ -50,6 +50,7 @@ boxplot_ACC_res <- mechanical_load_data |>
   ) +
   theme_light() +
   theme(
+    axis.text.x = element_text(size = 13, angle = 45, hjust = 1),
     axis.text.y = element_text(size = 13),
     axis.title.y = element_text(size = 16),
     legend.title = element_blank()
@@ -69,7 +70,6 @@ boxplot_LR_res <- mechanical_load_data |>
   ) +
   theme_light() +
   theme(
-    axis.text.x = element_text(size = 13, angle = 45, hjust = 1),
     axis.text.y = element_text(size = 13),
     axis.title.y = element_text(size = 16)
   ) +
@@ -99,8 +99,8 @@ boxplot_AR_res <- mechanical_load_data |>
 # Combine and save plots --------------------------------------------------
 
 fig1 <- boxplot_GRF_res + theme(axis.text.x = element_blank()) +
-  boxplot_ACC_res + theme(axis.text.x = element_blank()) +
-  boxplot_LR_res +
+  boxplot_LR_res + theme(axis.text.x = element_blank()) +
+  boxplot_ACC_res +
   boxplot_AR_res +
   plot_annotation(tag_levels = "A") +
   plot_layout(guides = "collect") &

### 2022-01-25 figures/figS1.R

@@ -50,6 +50,7 @@ boxplot_ACC_ver <- mechanical_load_data |>
   ) +
   theme_light() +
   theme(
+    axis.text.x = element_text(size = 13, angle = 45, hjust = 1),
     axis.text.y = element_text(size = 13),
     axis.title.y = element_text(size = 16),
     legend.title = element_blank()
@@ -69,7 +70,6 @@ boxplot_LR_ver <- mechanical_load_data |>
   ) +
   theme_light() +
   theme(
-    axis.text.x = element_text(size = 13, angle = 45, hjust = 1),
     axis.text.y = element_text(size = 13),
     axis.title.y = element_text(size = 16)
   ) +
@@ -99,8 +99,8 @@ boxplot_AR_ver <- mechanical_load_data |>
 # Combine and save plots --------------------------------------------------
 
 figS1 <- boxplot_GRF_ver + theme(axis.text.x = element_blank()) +
-  boxplot_ACC_ver + theme(axis.text.x = element_blank()) +
-  boxplot_LR_ver +
+  boxplot_LR_ver + theme(axis.text.x = element_blank()) +
+  boxplot_ACC_ver +
   boxplot_AR_ver +
   plot_annotation(tag_levels = "A") +
   plot_layout(guides = "collect") &

### 2022-01-25 manuscript/manuscript.Rmd
Write the results section

@@
 library(bookdown)
+library(broman)
 source(here("code", "02_explore.R"))
+```
+
+```{r GRF_accuracy}
+load(here("output", "prediction_models.rda"))
+res_GRF_accuracy <- purrr::map(
+  accuracy_res_GRF_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("ankle", "lower back", "hip"),
+    vector = "resultant"
+  )
+ver_GRF_accuracy <- purrr::map(
+  accuracy_ver_GRF_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("ankle", "lower back", "hip"),
+    vector = "vertical"
+  )
+GRF_accuracy <- rbind(res_GRF_accuracy, ver_GRF_accuracy)
+hip_GRF_accuracy <- GRF_accuracy %>%
+  filter(placement == "hip" & vector == "resultant")
+```
+
+```{r LR_accuracy}
+load(here("output", "prediction_models.rda"))
+res_LR_accuracy <- purrr::map(
+  accuracy_res_LR_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("ankle", "lower back", "hip"),
+    vector = "resultant"
+  )
+ver_LR_accuracy <- purrr::map(
+  accuracy_ver_LR_models,
+  ~ dplyr::mutate(.x, MAPE = as.numeric(MAPE))
+) %>%
+  purrr::map_dfr(rbind) %>%
+  dplyr::mutate(
+    placement = c("ankle", "lower back", "hip"),
+    vector = "vertical"
+  )
+LR_accuracy <- rbind(res_LR_accuracy, ver_LR_accuracy)
+hip_LR_accuracy <- LR_accuracy %>%
+  filter(placement == "hip" & vector == "resultant")
+```
+
+```{r ba_agreement}
+load(here("output", "loocv_data.rda"))
+percent_inside_loa <- function(data) {
+  lower <- lvmisc::loa(
+    data[[".actual"]], data[[".predicted"]], na.rm = TRUE
+  )$lower
+  upper <- lvmisc::loa(
+    data[[".actual"]], data[[".predicted"]], na.rm = TRUE
+  )$upper
+  diff <- dplyr::transmute(
+    data, diff = .data[[".actual"]] - .data[[".predicted"]]
+  )
+  diff <- dplyr::mutate(
+    diff,
+    inside = as.factor(ifelse(diff > lower & diff < upper, "yes", "no"))
+  )
+  round(prop.table(table(diff$inside))[[2]] * 100, 0)
+}
+
+ba_agreement <- min(
+  map_dbl(cv_res_GRF_models, percent_inside_loa) %>% min(),
+  map_dbl(cv_ver_GRF_models, percent_inside_loa) %>% min(),
+  map_dbl(cv_res_LR_models, percent_inside_loa) %>% min(),
+  map_dbl(cv_ver_LR_models, percent_inside_loa) %>% min()
+)
+```
+
+```{r ba_regression}
+source(here("code", "04_analyse_results.R"))
+ba_regression_ankle_R2 <- c(
+  ba_regression_res_GRF$ankle["adj.r.squared"],
+  ba_regression_ver_GRF$ankle["adj.r.squared"],
+  ba_regression_res_LR$ankle["adj.r.squared"],
+  ba_regression_ver_LR$ankle["adj.r.squared"]
+) %>%
+  as_vector()
 ```
@@
 
+# Results {.unnumbered}
+
+In our protocol, three distinct jump types were performed, in several different heights.
+This variety of jump trials produced a varying response in both the FP- and accelerometer-measured variables.
+These responses in pGRF, pLR, pACC and pAR, regarding their resultant vector, can be observed in Figure 1 (see Supplemental Figure S1 for the vertical vector).
+Panels A and B show the FP-measured variables normalized by the subjects body mass, the pGRF and pLR, respectively.
+It can be seen that in the drop jump trials, there is a tendency to increase the loading magnitude the higher the jump, and in the box jumps this tendency is lower.
+As for the continuous jumps, both assayed heights elicit resonably constant loadings.
+The same pattern can be observed in the accelerometer-measured variables, pACC and pAR, in Panels C and D, respectively.
+Also, regarding the influence of the accelerometer placement, in most of the trials accelerometers worn at hip and lower back produced approximately similar values, with a tendency to higher dispersion in data derived from lower back accelerometers.
+
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{fig1.png}
+		\caption{INSERT CAPTION}
+	\end{measuredfigure}
+\end{figure}
+
+Accelerometer-based mechanical loading prediction equations were developed based on linear mixed models.
+The regression coefficients, R^2^ and accuracy indices of the models derived from accelerometers worn at ankle, lower back and hip, and from the resultant and vertical vectors are presented in Table 1.
+The R^2^ values for the pGRF models ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`.
+These values represent a reasonably good prediction accuracy, that is also reflected on the MAPE values between `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% and `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%.
+Despite similar accuracy results achieved by all equations, data from the resultant vector of accelerometers at the hip have shown the lower accuracy error, represented by the smallest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
+Compared to the pGRF models, the pLR equations yielded poorer results.
+Their R^2^ ranged from `r min(LR_accuracy$R2_cond)` to `r max(LR_accuracy$R2_cond)` and MAPE from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%.
+However, similar to the pGRF models, the best pLR prediction equation in terms of prediction accuracy was the one from the resultant vector of hip-worn accelerometers, with a MAE of `r myround(hip_LR_accuracy$MAE, 1)` N, MAPE of `r myround(hip_LR_accuracy$MAPE * 100, 1)`\% and RMSE of `r myround(hip_LR_accuracy$RMSE, 1)` N.
+
+```{r accuracy_table, results = "show"}
+source(here("tables", "tab1.R"))
+tab1
+```
+
+To visualize the agreement between the pGRF and pLR values measured by the FP and predicted by the models, a series of Bland-Altman plots were built for both the resultant and vetical vectors and for all accelerometer placements.
+Figure 2 shows the Bland-Altman plots for the pGRF (Panel A) and pLR (Panel B) resultant vector of hip-worn accelerometers.
+Both plots have shown good agreement levels, as at least `r ba_agreement`\% of the data points are within the limits of agreement and data tended to cluster around zero.
+Bland-Altman plots for the other accelerometer placements and vectors have shown the same patterns described previously, and can be found in the Supplemental Figure S2.
+Also, as could be observed by the one-sample *t*-tests results, none of the equations presented a bias different than zero (*p* > 0.05).
+Furthermore, all of the equations have shown a proportional bias, i.e., the magnitude of the values influence the prediction error.
+However, as the linear regressions presented a very low R^2^ (maximum value of `r myround(max(ba_regression_ankle_R2), 2)`) this effect is neglectable.
+
+Finally, it is important to mention that an R package was developed [@Veras2021] to make easier to apply all the necessary data processing steps to use the prediction equations.
+The package website (https://lveras.com/impactr) has all the installation instructions and documentation necessary to apply this method on your own data.
+
 \newpage

### 2022-01-26 code/funs.R
Define functions to predict mechanical loading using equations from Veras et al. 2020

@@ -164,3 +164,124 @@ compute_accuracy <- function(actual, predicted) {
     2
   )
 }
+
+# Predict mechanical loading using the equations by Veras et al. (2020)
+# (doi.org/10.1007/s00198-020-05295-2)
+#
+# Params:
+#    data: A data.frame
+#    outcome, acc_placement, vector: A character string
+#
+# Returns:
+#    The input data.frame with the predicted values in the column
+# .predicted_Veras2020
+predict_Veras2020 <- function(data, outcome, acc_placement, vector) {
+  c <- get_coefficients(outcome, acc_placement, vector)
+  dplyr::mutate(
+    data,
+    .predicted_Veras2020 = c$b0 +
+      c$b1 * pACC_g +
+      c$b2 * (pACC_g^2) +
+      c$b3 * body_mass +
+      c$b4 * pACC_g * body_mass
+  )
+}
+
+# Get the mechanical loading prediction equation coefficients
+#
+# Coefficients from the equations by Veras et al. 2020
+# (doi.org/10.1007/s00198-020-05295-2)
+#
+# Params:
+#    outcome, acc_placement, vector: A character string
+#
+# Returns:
+#    A named list with the coefficients as follows:
+#      b0: Intercept
+#      b1: ACC
+#      b2: ACC2
+#      b3: body mass
+#      b4: ACC:body_mass
+get_coefficients <- function(outcome, acc_placement, vector) {
+  if (
+    outcome == "GRF" & acc_placement == "lower_back" & vector == "resultant"
+  ) {
+    list(
+      b0 = - 698.7031,
+      b1 = 1047.5129,
+      b2 = - 345.2605,
+      b3 = 3.8294,
+      b4 = 6.0219
+    )
+  } else if (
+    outcome == "GRF" & acc_placement == "hip" & vector == "resultant"
+  ) {
+    list(
+      b0 = - 300.9909,
+      b1 = 522.6850,
+      b2 = - 171.5606,
+      b3 = 3.9596,
+      b4 = 5.3671
+    )
+  } else if (
+    outcome == "GRF" & acc_placement == "lower_back" & vector == "vertical"
+  ) {
+    list(
+      b0 = - 776.8934,
+      b1 = 1042.9052,
+      b2 = - 336.2115,
+      b3 = 6.2132,
+      b4 = 5.0805
+    )
+  } else if (
[…]

### 2022-01-26 code/05_compare_equations.R
Initial commit

@@ -1,0 +1,92 @@
+# Load packages and functions ---------------------------------------------
+
+library(here)
+library(tidyverse)
+library(lvmisc)
+source(here("code", "funs.R"))
+
+# Load and prepare data ---------------------------------------------------
+
+load(here("data", "mechanical_load_data.rda"))
+
+lower_back_res_data <- mechanical_load_data |>
+  filter(acc_placement == "lower_back" & vector == "resultant")
+hip_res_data <- mechanical_load_data |>
+  filter(acc_placement == "hip" & vector == "resultant")
+
+lower_back_ver_data <- mechanical_load_data |>
+  filter(acc_placement == "lower_back" & vector == "vertical")
+hip_ver_data <- mechanical_load_data |>
+  filter(acc_placement == "hip" & vector == "vertical")
+
+# Predict using Veras et al. 2020 -----------------------------------------
+
+# GRF
+lower_back_res_GRF_df <- predict_Veras2020(
+  lower_back_res_data,
+  outcome = "GRF", acc_placement = "lower_back", vector = "resultant"
+)
+hip_res_GRF_df <- predict_Veras2020(
+  hip_res_data,
+  outcome = "GRF", acc_placement = "hip", vector = "resultant"
+)
+
+lower_back_ver_GRF_df <- predict_Veras2020(
+  lower_back_ver_data,
+  outcome = "GRF", acc_placement = "lower_back", vector = "vertical"
+)
+hip_ver_GRF_df <- predict_Veras2020(
+  hip_ver_data,
+  outcome = "GRF", acc_placement = "hip", vector = "vertical"
+)
+
+# LR
+lower_back_res_LR_df <- predict_Veras2020(
+  lower_back_res_data,
+  outcome = "LR", acc_placement = "lower_back", vector = "resultant"
+)
+hip_res_LR_df <- predict_Veras2020(
+  hip_res_data,
+  outcome = "LR", acc_placement = "hip", vector = "resultant"
+)
+
+lower_back_ver_LR_df <- predict_Veras2020(
+  lower_back_ver_data,
+  outcome = "LR", acc_placement = "lower_back", vector = "vertical"
+)
+hip_ver_LR_df <- predict_Veras2020(
+  hip_ver_data,
[…]

### 2022-01-26 code/05_compare_equations.R
Save predictions

@@ -90,3 +90,16 @@ compute_accuracy(
 compute_accuracy(
   hip_ver_LR_df$pLR_Ns, hip_ver_LR_df$.predicted_Veras2020
 )
+
+# Save predictions --------------------------------------------------------
+
+if (!dir.exists(here("output"))) {
+  dir.create(here("output"))
+}
+save(
+  lower_back_res_GRF_df, hip_res_GRF_df,
+  lower_back_ver_GRF_df, hip_ver_GRF_df,
+  lower_back_res_LR_df, hip_res_LR_df,
+  lower_back_ver_LR_df, hip_ver_LR_df,
+  file = here("output", "veras2020_predictions.rda")
+)

### 2022-01-26 figures/fig3.R
Intial commit

@@ -1,0 +1,151 @@
+# Load packages -----------------------------------------------------------
+
+library(here)
+library(tidyverse)
+library(lvmisc)
+library(ggsci)
+library(patchwork)
+library(ragg)
+
+# Load and prepare data ---------------------------------------------------
+
+load(here("output", "veras2020_predictions.rda"))
+load(here("output", "loocv_data.rda"))
+
+plot_data_GRF <- cv_res_GRF_models$hip |>
+  left_join(
+    select(
+      hip_res_GRF_df,
+      id, trial, filename, subj, acc_placement, vector,
+      jump_type, jump_height, .predicted_Veras2020
+    )
+  ) |>
+  select(pACC_g, .predicted, .predicted_Veras2020) |>
+  pivot_longer(
+    cols = starts_with("."),
+    names_to = "equation",
+    values_to = "pGRF_N"
+  ) |>
+  mutate(
+    equation = recode_factor(
+      as.factor(equation),
+      .predicted = "Jumping equation",
+      .predicted_Veras2020 = "Walking equation"
+    )
+  )
+
+plot_data_LR <- cv_res_LR_models$hip |>
+  left_join(
+    select(
+      hip_res_LR_df,
+      id, trial, filename, subj, acc_placement, vector,
+      jump_type, jump_height, .predicted_Veras2020
+    )
+  ) |>
+  select(pAR_gs, .predicted, .predicted_Veras2020) |>
+  pivot_longer(
+    cols = starts_with("."),
+    names_to = "equation",
+    values_to = "pLR_Ns"
+  ) |>
+  mutate(
+    equation = recode_factor(
+      as.factor(equation),
+      .predicted = "Jumping equation",
+      .predicted_Veras2020 = "Walking equation"
+    )
+  )
+
+# Plot GRF predictions ----------------------------------------------------
+
+predicted_GRF_plot <- ggplot(plot_data_GRF) +
+  geom_point(
+    aes(x = pACC_g, y = pGRF_N, color = equation),
+    alpha = 0.3
+  ) +
+  geom_hline(yintercept = 0) +
+  scale_color_manual(values = c("#0072B5", "#BC3C29")) +
+  scale_y_continuous(
+    expand = c(0, 0),
+    limits = c(-15000, 10000),
+    breaks = seq(-15000, 10000, 5000)
+  ) +
+  scale_x_continuous(
+    expand = c(0, 0),
+    limits = c(0, 14),
[…]

### 2022-01-26 code/05_compare_equations.R
Save the accuracy indices dataframes

@@ -62,32 +62,32 @@ hip_ver_LR_df <- predict_Veras2020(
 # Compute accuracy --------------------------------------------------------
 
 # GRF
-compute_accuracy(
+lower_back_res_GRF_accuracy <- compute_accuracy(
   lower_back_res_GRF_df$pGRF_N, lower_back_res_GRF_df$.predicted_Veras2020
 )
-compute_accuracy(
+hip_res_GRF_accuracy <- compute_accuracy(
   hip_res_GRF_df$pGRF_N, hip_res_GRF_df$.predicted_Veras2020
 )
 
-compute_accuracy(
+lower_back_ver_GRF_accuracy <- compute_accuracy(
   lower_back_ver_GRF_df$pGRF_N, lower_back_ver_GRF_df$.predicted_Veras2020
 )
-compute_accuracy(
+hip_ver_GRF_accuracy <- compute_accuracy(
   hip_ver_GRF_df$pGRF_N, hip_ver_GRF_df$.predicted_Veras2020
 )
 
 # LR
-compute_accuracy(
+lower_back_res_LR_accuracy <- compute_accuracy(
   lower_back_res_LR_df$pLR_Ns, lower_back_res_LR_df$.predicted_Veras2020
 )
-compute_accuracy(
+hip_res_LR_accuracy <- compute_accuracy(
   hip_res_LR_df$pLR_Ns, hip_res_LR_df$.predicted_Veras2020
 )
 
-compute_accuracy(
+lower_back_ver_LR_accuracy <- compute_accuracy(
   lower_back_ver_LR_df$pLR_Ns, lower_back_ver_LR_df$.predicted_Veras2020
 )
-compute_accuracy(
+hip_ver_LR_accuracy <- compute_accuracy(
   hip_ver_LR_df$pLR_Ns, hip_ver_LR_df$.predicted_Veras2020
 )
 
@@ -102,4 +102,11 @@ save(
   lower_back_res_LR_df, hip_res_LR_df,
   lower_back_ver_LR_df, hip_ver_LR_df,
   file = here("output", "veras2020_predictions.rda")
+)
+save(
+  lower_back_res_GRF_accuracy, hip_res_GRF_accuracy,
+  lower_back_ver_GRF_accuracy, hip_ver_GRF_accuracy,
+  lower_back_res_LR_accuracy, hip_res_LR_accuracy,
+  lower_back_ver_LR_accuracy, hip_ver_LR_accuracy,
+  file = here("output", "veras2020_accuracy.rda")
 )

### 2022-01-26 tables/tab2.R
Initial commit

@@ -1,0 +1,64 @@
+# Load packages and functions ---------------------------------------------
+
+library(here)
+library(tidyverse)
+library(lvmisc)
+library(broman)
+library(kableExtra)
+source(here("code", "funs.R"))
+
+# Load and prepare data ---------------------------------------------------
+
+load(here("output", "veras2020_accuracy.rda"))
+
+# Build table -------------------------------------------------------------
+
+tab2_df <- list(
+  lower_back_res_GRF_accuracy,
+  hip_res_GRF_accuracy,
+  lower_back_ver_GRF_accuracy,
+  hip_ver_GRF_accuracy,
+  lower_back_res_LR_accuracy,
+  hip_res_LR_accuracy,
+  lower_back_ver_LR_accuracy,
+  hip_ver_LR_accuracy
+) |>
+  map_dfr(~ mutate(.x, MAPE = as.numeric(MAPE) * 100)) |>
+  mutate(
+    across(where(is.numeric), myround, 1),
+    MAPE = paste0(MAPE, "\\%"),
+    Vector = rep(c("Resultant", "", "Vertical", ""), 2),
+    `Accelerometer placement` = rep(c("Lower back", "Hip"), 4),
+    .before = MAE
+  )
+
+tab2 <- knitr::kable(
+  tab2_df, booktabs = TRUE, escape = FALSE, linesep = "",
+  label = "none",
+  caption = "Regression equations, $R^2$ and accuracy indices",
+  col.names = c(
+    "Vector",
+    linebreak("Accelerometer\nplacement"),
+    "MAE",
+    "MAPE",
+    "RMSE"
+  ),
+  align = c(rep("l", 2), rep("r", 3))
+) %>%
+  footnote(
+    general = "Abbreviations: MAE, mean absolute error; MAPE, mean absolute percent error; RMSE, root mean square error",
+    general_title = "",
+    threeparttable = TRUE
+  ) %>%
+  # kable_styling(latex_options = "scale_down") %>%
+  pack_rows(
+    "Peak ground reaction force prediction", 1, 4,
+    latex_gap_space = "0.5em",
+    bold = FALSE
+  ) %>%
+  pack_rows(
+    "Peak loading rate prediction", 5, 8,
+    latex_gap_space = "0.5em",
+    bold = FALSE
+  ) %>%
+  landscape()

### 2022-01-26 manuscript/manuscript.Rmd
Add content to methods and results sections

@@
 The numerical assessment was made by computing the following accuracy indices: the mean absolute error (MAE), mean absolute percent error (MAPE) and root mean square error (RMSE).
+Despite the absence of a standard criterion for establishing an acceptable error in mechanical loading prediction, we regarded a MAPE <8\% to be an accurate prediction based on past research [@Fortune2014; @Neugebauer2014; @Ancillao2018, @Veras2020].
 To visualize the agreement between the values predicted by the models and measured by the FP, Bland-Altman plots were drawn.
@@
 Finally, linear regressions were employed to examine if the difference between actual and predicted values were affected by their magnitude, verifying, therefore, whether the prediction was constant throughout the magnitude range that was tested [@Giavarina2015].
+
+Lastly, our pGRF and pLR prediction models were compared with a previously published equation by our group [@Veras2020].
+As, for the best of our knowledge, there is no mechanical loading prediction equation developed for jumping activies, we chose to compare it with previous models from our group that were validated in a similar sample.
+The models used in the comparison were the ones predicting pGRF and pLR from both the resultant and vertical vectors of accelerometers worn at lower back and hip.
+The MAE, MAPE and RMSE were calculated using the values predicted using Veras et al.
+[@Veras2020] equation.
 
@@
 The regression coefficients, R^2^ and accuracy indices of the models derived from accelerometers worn at ankle, lower back and hip, and from the resultant and vertical vectors are presented in Table 1.
-The R^2^ values for the pGRF models ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`.
-These values represent a reasonably good prediction accuracy, that is also reflected on the MAPE values between `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% and `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%.
+The R^2^ values for the pGRF models ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`, showing that the models can explain at least `r round(min(GRF_accuracy$R2_cond), 1) * 100`\% of the pGRF variance.
+The observed MAPE values were between `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% and `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which falls outside of the proposed threshold of 8\%.
+Nevertheless, these MAPE values are similar to the results usually found in accelerometer-based energy expenditure prediction [@Diniz-Sousa2020], which have widespread use in research.
 Despite similar accuracy results achieved by all equations, data from the resultant vector of accelerometers at the hip have shown the lower accuracy error, represented by the smallest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
-Compared to the pGRF models, the pLR equations yielded poorer results.
+Compared to the pGRF models, the pLR equations yielded poorer results, which is expected based on past research [@Fortune2014; @Veras2020].
 Their R^2^ ranged from `r min(LR_accuracy$R2_cond)` to `r max(LR_accuracy$R2_cond)` and MAPE from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%.
@@
 
+```{r accuracy_table2, results = "show"}
+source(here("tables", "tab2.R"))
+tab2
+```
+
 Finally, it is important to mention that an R package was developed [@Veras2021] to make easier to apply all the necessary data processing steps to use the prediction equations.

### 2022-01-27 manuscript/manuscript.Rmd
Add new results regarding the regressions comparison

@@
 
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{fig1.png}
+		\caption{INSERT CAPTION}
+	\end{measuredfigure}
+\end{figure}
+
 In our protocol, three distinct jump types were performed, in several different heights.
@@
 
-\begin{figure}[ht]
-	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
-	\centering
-	\begin{measuredfigure}
-		\includegraphics[width=\textwidth]{fig1.png}
-		\caption{INSERT CAPTION}
-	\end{measuredfigure}
-\end{figure}
+```{r accuracy_table, results = "show"}
+source(here("tables", "tab1.R"))
+tab1
+```
 
@@
 
-```{r accuracy_table, results = "show"}
-source(here("tables", "tab1.R"))
-tab1
-```
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{fig2.png}
+		\caption{INSERT CAPTION}
+	\end{measuredfigure}
+\end{figure}
 
@@
 
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{fig3.png}
+		\caption{INSERT CAPTION}
+	\end{measuredfigure}
+\end{figure}
+
+Although the accuracy of our equations showed to be beyond the established 8\% threshold, when applying a model not developed for jumping activies we can se that the accuracy tremendously decreases, reaching values of MAPE around `r floor(mean(as.numeric(substr(tab2_df$MAPE, 1, nchar(tab2_df$MAPE) - 2))[1:4]) / 50) * 50`\% for pGRF prediction and around `r floor(mean(as.numeric(substr(tab2_df$MAPE, 1, nchar(tab2_df$MAPE) - 2))[5:8]) / 10) * 10`\% for pLR prediction as can be observed in Table 2.
+Furthermore, applying to our accelerometer-recorded jumping data a model that is not specific for this activity yield unresonable results, such as the negative values for pGRF that can be observed in Figure 3, Panel A.
+
 Finally, it is important to mention that an R package was developed [@Veras2021] to make easier to apply all the necessary data processing steps to use the prediction equations.

### 2022-02-04 manuscript/manuscript.Rmd
Begin discussion

@@
 
+# Discussion {.unnumbered}
+
+This study aimed to develop accelerometer-based mechanical loading predition equations for jumps of diverse types and heights, testing different accelerometer placements and using both the resultant and vertical vectors.
+The results have shown that jumping pGRF prediction models based on accelerometry data have a lower accuracy compared to models developed for other activies, such as walking and running.
+Despite that, these models are the correct choice for this kind of activity eliciting hight levels of impact, as applying a model developed for walking in a sample similar to the current study yielded far worse results in terms of prediction accuracy.
+Also, as expected, the pLR prediction models performed poorly compared to the pGRF equations.
+
 \newpage

### 2022-02-04 figures/fig3.R
Add the actual values to the plot and change the points shape by the type of measure

@@ -17,20 +17,21 @@ plot_data_GRF <- cv_res_GRF_models$hip |>
     select(
       hip_res_GRF_df,
       id, trial, filename, subj, acc_placement, vector,
-      jump_type, jump_height, .predicted_Veras2020
+      jump_type, jump_height, pGRF_N, .predicted_Veras2020
     )
   ) |>
-  select(pACC_g, .predicted, .predicted_Veras2020) |>
+  select(pACC_g, pGRF_N, .predicted, .predicted_Veras2020) |>
   pivot_longer(
-    cols = starts_with("."),
-    names_to = "equation",
-    values_to = "pGRF_N"
+    cols = c(pGRF_N, .predicted, .predicted_Veras2020),
+    names_to = "measure",
+    values_to = "pGRF"
   ) |>
   mutate(
-    equation = recode_factor(
-      as.factor(equation),
-      .predicted = "Jumping equation",
-      .predicted_Veras2020 = "Walking equation"
+    measure = recode_factor(
+      as.factor(measure),
+      pGRF_N = "Measured by the force plates",
+      .predicted = "Predicted by equation for jumping",
+      .predicted_Veras2020 = "Predicted by equation for walking"
     )
   )
 
@@ -39,20 +40,21 @@ plot_data_LR <- cv_res_LR_models$hip |>
     select(
       hip_res_LR_df,
       id, trial, filename, subj, acc_placement, vector,
-      jump_type, jump_height, .predicted_Veras2020
+      jump_type, jump_height, pLR_Ns, .predicted_Veras2020
     )
   ) |>
-  select(pAR_gs, .predicted, .predicted_Veras2020) |>
+  select(pAR_gs, pLR_Ns, .predicted, .predicted_Veras2020) |>
   pivot_longer(
-    cols = starts_with("."),
-    names_to = "equation",
-    values_to = "pLR_Ns"
+    cols = c(pLR_Ns, .predicted, .predicted_Veras2020),
+    names_to = "measure",
+    values_to = "pLR"
   ) |>
   mutate(
-    equation = recode_factor(
-      as.factor(equation),
-      .predicted = "Jumping equation",
-      .predicted_Veras2020 = "Walking equation"
+    measure = recode_factor(
+      as.factor(measure),
+      pLR_Ns = "Measured by the force plates",
[…]

### 2022-02-06 manuscript/manuscript.Rmd
Write discussion section

@@
 The results have shown that jumping pGRF prediction models based on accelerometry data have a lower accuracy compared to models developed for other activies, such as walking and running.
-Despite that, these models are the correct choice for this kind of activity eliciting hight levels of impact, as applying a model developed for walking in a sample similar to the current study yielded far worse results in terms of prediction accuracy.
+Despite that, these models are the correct choice for this kind of activity eliciting hight levels of impact, as applying a model developed for lower impact activities in a sample similar to the current study yielded far worse results in terms of prediction accuracy.
 Also, as expected, the pLR prediction models performed poorly compared to the pGRF equations.
+
+Bone is known to increase its tissue mass and enhance its micro-architecture arrangement in response to mechanical stimulation, particularly at higher magnitudes.
+This type of high-intensity impact activies have shown to be more associated with bone health improval, even if they are usually performed in low quantities throughout the day.
+In fact, Vainionpaa et al demonstrated that performing a relatively low number of pACC above 4.9*g* daily is positively associated with increases in bone mineral density.
+Usually, this acceleration magnitude can only be achieved by explosive (?) movements such as jumps.
+Nevertheless, accelerometry-based mechanical loading regression models have mostly been validated with a set of activies, such as walking and running, that elict relatively low acceleration magnitudes, usually ranging from X to Y*g*.
+
+The use of a prediction model based on accelerometer data should ideally be done in conditions similar to the study that validated the model, specially in terms of sample characteristics and type and intensity of the activities performed.
+Therefore, if we want to accurately predict the mechanical loading caused by high-impact activities, we should choose a prediction model that has been validated using these activities.
+Even though the cross-validation accuracy of such a model was poorer compared with other models developed for activities such as walking and running, as demonstrated by our results, the use of these models is crucial to achieve an acceptable prediction accuracy.
+It was demonstrated by testing a model previously developed by our group [@Veras2020] in a similar sample, but on walking, on the accelerometer-recorded jumping data from the current study.
+The accuracy indices yielded by applying these models have greatly worsened and these equations also produced some abnormal values, mainly in the pGRF model, in which it predicted negative values of force.
+These negative pGRF values were mainly as a result of a predictor, the second degree polynomial of the acceleration, possessing a negative coefficient.
+While this coefficient captured the plateauing of pGRF as pACC magnitude augments during walking, effectively improving the model [@Veras2020] for walking, this same pattern is not present in the pGRF vs.
+pACC relationship during jumping, and, therefore, only impairs the accuracy of these equations when applyed to this kind of activity.
+
+Loading rate is a highly sensitive parameter, that is usually calculated by deriving the force from the beggining of the foot contact to the floor to the point where the pGRF is achieved.
+This is a small time window, usually between X and Y seconds for the jumps that were performed in our protocol, and requires a fine sampling acquisition that captures small graduations on the force values throughout time, which is difficult to achieve with common accelerometer sampling frequecies, that are usually 100Hz at maximum.
+Based on this and on on results from previous investigations [@Fortune2014; @Veras2020], pLR prediction models are expected to have a lower accuracy compared to the pGRF equations.
+Nevertheless, although our pLR prediction models had a worse accuracy compared with the pGRF models, its accuracy was similar or even better than the results from previous studies that developed models to predict pLR during walking only [@Fortune2014; @Veras2020].
+<!-- TODO: Try and state that this samplling frquency issue is theoretical -->
+
+In our protocol, three accelerometer placements (ankle, lower back and hip) and two vectors (resultant and vertical) have been tested.
+Overall, the prediction models performance was reasonably consistent among the different placement and vector combinations, however, some conclusions can be draw.
+First, the hip accelerometer placement, that has traditionally been widely employed in research for the direct measure of physical activity, yielded a sligthly better accuracy compared the other placements tested.
+This could be useful, as previous data collected from hip-worn accelerometers, even for other purposes, can be adequately used to obtain mechanical loading predictions.
+Also, studies that aim to extract mechanical loading and other variables from accelerometry data, that use the hip as the default placement, can comfortably do so.
+Second, as the magnitude of the resultant vector of both pGRF and pLR comes from (?) the vertical vector, and there were no important (?) differences in performance between models using both vectors, they can be used interchangeably to predict mechanical loading.
+Besides that, while models using data from the vertical vector allow the use uniaxial accelerometers, data derived from the resultant vector have the important advantage of not depending on the correct acceleromter axes orientation, that cannot be guaranteed in free-living data collection.
+
+The development of accelerometry-based mechanical loading prediction models was done with the direct purpose of allowing to assess the loading beared by the skeletal system during daily physical activity.
+This would allow some clinical applications.
+For instance, these models, when used in combination with other prediction equations developed for lower impact activities [@Veras2020], can be used to fully analyse the mechanical loading spectrum that people with different characteristics are submitted to daily and relate to bone health parameters.
+This would allow a more complete understanding on how mechanical loading influence bone healh, and also allow to identify which loading volume and magnitude range, for example, are more efficient in inducing bone health benefits.
+Also, by the development and distribution of an R package that automates the steps necessary to apply the models that were proposed was made to enhance its applicability.
+This can prove to be an important tool, as the application of these models depend on non-trivial data processing skills, such as applying digital filters to reduce noise from the signal, detection of the peaks in this signal and the computation of derivatives to obtain the acceleration rate, necessary for the prediction of pLR.
 

### 2022-02-07 manuscript/manuscript.Rmd
Finish the discussion

@@
 Furthermore, all of the equations have shown a proportional bias, i.e., the magnitude of the values influence the prediction error.
-However, as the linear regressions presented a very low R^2^ (maximum value of `r myround(max(ba_regression_ankle_R2), 2)`) this effect is neglectable.
+However, as the linear regressions presented a very small R^2^ (maximum value of `r myround(max(ba_regression_ankle_R2), 2)`) this effect is neglectable.
 
@@
 This type of high-intensity impact activies have shown to be more associated with bone health improval, even if they are usually performed in low quantities throughout the day.
-In fact, Vainionpaa et al demonstrated that performing a relatively low number of pACC above 4.9*g* daily is positively associated with increases in bone mineral density.
-Usually, this acceleration magnitude can only be achieved by explosive (?) movements such as jumps.
-Nevertheless, accelerometry-based mechanical loading regression models have mostly been validated with a set of activies, such as walking and running, that elict relatively low acceleration magnitudes, usually ranging from X to Y*g*.
+In fact, Vainionpää et al demonstrated that performing a relatively low number (<100) of pACC above 4.9*g* daily is positively associated with increases in bone mineral density.
+Usually, this acceleration magnitude can be achieved mostly by ballistic movements such as jumps.
+Nevertheless, accelerometry-based mechanical loading regression models have mostly been validated with a set of activies, such as walking and running, that elict relatively low acceleration magnitudes, usually ranging from approximately 1.2 to 6*g* [@Neugebauer2014; @Veras2020].
 
@@
 Loading rate is a highly sensitive parameter, that is usually calculated by deriving the force from the beggining of the foot contact to the floor to the point where the pGRF is achieved.
-This is a small time window, usually between X and Y seconds for the jumps that were performed in our protocol, and requires a fine sampling acquisition that captures small graduations on the force values throughout time, which is difficult to achieve with common accelerometer sampling frequecies, that are usually 100Hz at maximum.
+This is a small time window, usually less than 0.2 seconds for the jumps that were performed in our protocol, and requires a fine sampling acquisition that captures small graduations on the force values throughout time, which is difficult to achieve with common accelerometer sampling frequecies, that are usually 100Hz at maximum.
 Based on this and on on results from previous investigations [@Fortune2014; @Veras2020], pLR prediction models are expected to have a lower accuracy compared to the pGRF equations.
 Nevertheless, although our pLR prediction models had a worse accuracy compared with the pGRF models, its accuracy was similar or even better than the results from previous studies that developed models to predict pLR during walking only [@Fortune2014; @Veras2020].
-<!-- TODO: Try and state that this samplling frquency issue is theoretical -->
 
@@
 
+Although the models presented in this study showed to be valid, some limitations ought to be considered.
+First, the choice of activity to elicit high-impact loading was jumping, and several combinations of jump type and height were tested.
+Nevertheless, the results presented here cannot be assured for other high-impact activities, or even for other types of jump that were not tested.
+Second, no external sample for model validation was recruited.
+However, the leave-one-out cross-validation method that was used is the recommended in these situations [@Staudenmayer2012].
+Third, to apply these models in accelerometer-recorded free-living physical activity data, there is a need to identify, in the acceleration signal, the periods in which the activity being performed is jumping.
+Therefore, methods for physical activity pattern detection should be used in combination with these models.
+
+Mechanical loading prediction equations based on accelerometer data have already been proposed for several populations [@Neugebauer2012; @Neugebauer2018; @Veras2020] and using accelerometer in different body placements [@Fortune2014; @Veras2020].
+What these past few studies have in common is that they validated their models using locomotion activities, as walking only or walking and runnning.
+The current study, by setting means to predict mechanical loading in several types and heights of jumps, broadens the application of such models, getting a step closer of the ability to accurately identify daily-living mechanical loading due to physical activity.
+This ability would allow the monitoring of loading in different contexts and to more fully understand the relationship between mechanical stimulation and its consequent effect on bone health.
+
 \newpage

### 2022-02-07 manuscript/manuscript.Rmd
Add references

@@
 
-Bone is known to increase its tissue mass and enhance its micro-architecture arrangement in response to mechanical stimulation, particularly at higher magnitudes.
+Bone is known to increase its tissue mass and enhance its micro-architecture arrangement in response to mechanical stimulation [@Turner2003; @Fonseca2014], particularly at higher magnitudes [@Vainionpaa2006; @Stiles2017].
 This type of high-intensity impact activies have shown to be more associated with bone health improval, even if they are usually performed in low quantities throughout the day.
-In fact, Vainionpää et al demonstrated that performing a relatively low number (<100) of pACC above 4.9*g* daily is positively associated with increases in bone mineral density.
-Usually, this acceleration magnitude can be achieved mostly by ballistic movements such as jumps.
+In fact, Vainionpää et al [@Vainionpaa2006] demonstrated that performing a relatively low number (<100) of pACC above 4.9*g* daily is positively associated with increases in bone mineral density.
+Usually, this acceleration magnitude can be achieved mostly by ballistic movements such as jumps [@Vainionpaa2006; @Rowlands2012].
 Nevertheless, accelerometry-based mechanical loading regression models have mostly been validated with a set of activies, such as walking and running, that elict relatively low acceleration magnitudes, usually ranging from approximately 1.2 to 6*g* [@Neugebauer2014; @Veras2020].
 
-The use of a prediction model based on accelerometer data should ideally be done in conditions similar to the study that validated the model, specially in terms of sample characteristics and type and intensity of the activities performed.
+The use of a prediction model based on accelerometer data should ideally be done in conditions similar to the study that validated the model, specially in terms of sample characteristics and type and intensity of the activities performed [@Welk2005].
 Therefore, if we want to accurately predict the mechanical loading caused by high-impact activities, we should choose a prediction model that has been validated using these activities.
@@
 
-Loading rate is a highly sensitive parameter, that is usually calculated by deriving the force from the beggining of the foot contact to the floor to the point where the pGRF is achieved.
+Loading rate is a highly sensitive parameter, that is usually calculated by deriving the force from the beggining of the foot contact to the floor to the point where the pGRF is achieved [@Liikavainio2007; @Fortune2014; @Veras2020].
 This is a small time window, usually less than 0.2 seconds for the jumps that were performed in our protocol, and requires a fine sampling acquisition that captures small graduations on the force values throughout time, which is difficult to achieve with common accelerometer sampling frequecies, that are usually 100Hz at maximum.
@@
 Overall, the prediction models performance was reasonably consistent among the different placement and vector combinations, however, some conclusions can be draw.
-First, the hip accelerometer placement, that has traditionally been widely employed in research for the direct measure of physical activity, yielded a sligthly better accuracy compared the other placements tested.
+First, the hip accelerometer placement, that has traditionally been widely employed in research for the direct measure of physical activity [@Migueles2017], yielded a sligthly better accuracy compared the other placements tested.
 This could be useful, as previous data collected from hip-worn accelerometers, even for other purposes, can be adequately used to obtain mechanical loading predictions.
 Also, studies that aim to extract mechanical loading and other variables from accelerometry data, that use the hip as the default placement, can comfortably do so.
-Second, as the magnitude of the resultant vector of both pGRF and pLR comes from (?) the vertical vector, and there were no important (?) differences in performance between models using both vectors, they can be used interchangeably to predict mechanical loading.
-Besides that, while models using data from the vertical vector allow the use uniaxial accelerometers, data derived from the resultant vector have the important advantage of not depending on the correct acceleromter axes orientation, that cannot be guaranteed in free-living data collection.
+Second, as the magnitude of the resultant vector of both pGRF and pLR, in the activities that were tested, is derived mostly from the vertical vector, and there were no considerable differences in performance between models using both vectors, they can be used interchangeably to predict mechanical loading.
+Besides that, while models using data from the vertical vector allow the use uniaxial accelerometers, data derived from the resultant vector have the important advantage of not depending on the correct acceleromter axes orientation, that cannot be guaranteed in free-living data collection [@Welk2005].
 
@@
 This would allow a more complete understanding on how mechanical loading influence bone healh, and also allow to identify which loading volume and magnitude range, for example, are more efficient in inducing bone health benefits.
-Also, by the development and distribution of an R package that automates the steps necessary to apply the models that were proposed was made to enhance its applicability.
+Also, by the development and distribution of an R package that automates the steps necessary to apply the models that were proposed was made to enhance its applicability [@Veras2021].
 This can prove to be an important tool, as the application of these models depend on non-trivial data processing skills, such as applying digital filters to reduce noise from the signal, detection of the peaks in this signal and the computation of derivatives to obtain the acceleration rate, necessary for the prediction of pLR.

### 2022-02-07 manuscript/manuscript.Rmd
Add figure captions

@@
 		\includegraphics[width=\textwidth]{fig1.png}
-		\caption{INSERT CAPTION}
+    \caption{Distribution of the resultant vector of mechanical loading (peak ground reaction force and peak loading rate) and acceleration (peak acceleration and peak acceleration rate) variables per jump type and height combination.
+Abbreviations: BW, body weight; pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.}
 	\end{measuredfigure}
@@
 		\includegraphics[width=\textwidth]{fig2.png}
-		\caption{INSERT CAPTION}
+		\caption{Bland-Altman plots showing agreement between actual and predicted peak ground reaction force (Panel A) and loading rate (Panel B) for the resultant vector from hip-worn accelerometers.
+Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
+Abbreviations: pGRF, peak ground reaction force; pLR, peak loading rate.}
 	\end{measuredfigure}
@@
 		\includegraphics[width=\textwidth]{fig3.png}
-		\caption{INSERT CAPTION}
+    \caption{A comparison of the relationship between peak acceleration and peak ground reaction force (Panel A) and peak acceleration rate and peak loading rate (Panel B) in three ways: i) as measured by the force plates (black round points); ii) predicted by the equations developed in the current study (blue triangle points); and iii) predicted by the equations developed for walking by Veras et al (red square points).
+Abbreviations: pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.}
 	\end{measuredfigure}
@@
 <div id="refs"></div>
+
+\newpage
+
+# Figure captions {.unnumbered}
+
+**Figure 1:** Distribution of the resultant vector of mechanical loading (peak ground reaction force and peak loading rate) and acceleration (peak acceleration and peak acceleration rate) variables per jump type and height combination.
+Abbreviations: BW, body weight; pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.
+
+**Figure 2:** Bland-Altman plots showing agreement between actual and predicted peak ground reaction force (Panel A) and loading rate (Panel B) for the resultant vector from hip-worn accelerometers.
+Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
+Abbreviations: pGRF, peak ground reaction force; pLR, peak loading rate.
+
+**Figure 3:** A comparison of the relationship between peak acceleration and peak ground reaction force (Panel A) and peak acceleration rate and peak loading rate (Panel B) in three ways: i) as measured by the force plates (black round points); ii) predicted by the equations developed in the current study (blue triangle points); and iii) predicted by the equations developed for walking by Veras et al (red square points).
+Abbreviations: pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.
+
+**Supplemental Figure S1:** Distribution of the vertical vector of mechanical loading (peak ground reaction force and peak loading rate) and acceleration (peak acceleration and peak acceleration rate) variables per jump type and height combination.
+Abbreviations: BW, body weight; pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.
+
+**Supplemental Figure S2:** Bland-Altman plots showing agreement between actual and predicted peak ground reaction force for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias ±1.96 standard deviation).
+Abbreviations: pGRF, peak ground reaction force.

### 2022-02-27 manuscript/manuscript.Rmd
Add objective
Write the introduction

@@
 ---
-# title:
+title: "Using raw accelerometer data to predict high-impact mechanical loading"
 output:
@@
 
+# Introduction {.unnumbered}
+
+Bone reponds to mechanical loading by altering its properties, such as density, geometry, and strength [@Turner2003; @Fonseca2014; @Santos2017].
+Physical activity is a way of eliciting mechanical stimuli, and it is well established to improve bone health.
+However, there is no consensus in which characteristics of the physical activity-related mechanical stimulus are better associated with bone health benefits.
+Nevertheless, there is evidence that that high-impact exercise, although performed scarcely throughout the day, may play an important role in inducing positive bone adaptations [@Vainionpaa2006; @Stiles2017].
+
+But, even when knowing precisely which is the best loading regimen to improve bone health, the question of how to measure and monitor this load remains.
+Bone mechanical loading is usually measured in laboratory settings, using instruments such as force plates (FP) [@Medved2000].
+Measurements in this instruments, that are confined to a laboratory space, are not applicable to controling the loads a person is routinely subjected in free-living contexts.
+To surpass this issue, in the last few years, methods using wearable devices such as accelerometers have been proposed to measure mechanical loading, more precisely,  ground reaction force (GRF) [@Neugebauer2012; @Neugebauer2014; @Fortune2014; @Neugebauer2018; @Veras2020] and loading rate (LR) [@Fortune2014; @Veras2020].
+These devices are light, small, can be used throughout days, and have a relatively affordable price.
+The aforementioned characteristics make the accelerometers well suited instruments to measure mechanical loading on a daily basis.
+
+However, the validity of the accelerometer-predicted mechanical loading values depends on some factors, as the sample characteristics [@Welk2005].
+In fact, due to this, several calibration studies have been done in several populations, as children [@Neugebauer2012], adults [@Neugebauer2014; @Fortune2014], military personnel [@Neugebauer2018], and subjects with obesity [@Veras2020].
+Another important factor that influence the predictions validity is the type of activities performed in the calibration study [@Bassett2012].
+Until now, studies that calibrate accelerometers to predict mechanical loading variables have employed usually activities as walking [@Fortune2014; @Neugebauer2018; @Veras2020], or walking and running [@Neugebauer2012; @Neugebauer2014].
+Nevertheless, activities eliciting higher impact loads have not yet been tested.
+
+Therefore, the objective of this study was to develop accelerometry-based mechanical loading prediction equations for high-impact activites in subjects from normal weight to severe obesity.
+
 # Methods {.unnumbered}
@@
 
-After familiarization, participants performed a series of trials envolving jumps of different types and heights.
+In this study, jumps were chosen as the high-impact activites to be performed.
+After familiarization, participants completed a series of trials envolving jumps of different types and heights.
 First, drop jumps were executed from steps varying from 5cm to 40cm high, with 5cm increments.
@@
 Drop and box jumps were executed in two sets of four jumps at each height, with 30s interval, and the continuous jumps in one set of 20s at each height.
-The landing phase of the drop jumps, the take-off phase of the box jumps and both phases of the continuous jumps were recorded in force plates (1000Hz; AMTI Corporation, Watertown, USA).
+The landing phase of the drop jumps, the take-off phase of the box jumps and both phases of the continuous jumps were recorded in FC (1000Hz; AMTI Corporation, Watertown, USA).
 Participants were instructed to jump with no restrictions to arms movement, land two-footed and then remain stationary on the force plate for five seconds (except in the continuous jumps).
@@
 The rates of change were computed through a centered derivative from the beginning of the foot contact to the curve peak (Equation 1).
-The peak loading rate (pLR) and peak acceleration rate (pAR) were defined as the maximum value in the array (Equation 2).
+The peak LR (pLR) and peak acceleration rate (pAR) were defined as the maximum value in the array (Equation 2).
 

### 2022-02-27 manuscript/manuscript.Rmd
Revise the manuscript

@@
 
-But, even when knowing precisely which is the best loading regimen to improve bone health, the question of how to measure and monitor this load remains.
+But, even when knowing precisely which the best loading regimen is to improve bone health, the question of how to measure and monitor this load remains.
 Bone mechanical loading is usually measured in laboratory settings, using instruments such as force plates (FP) [@Medved2000].
-Measurements in this instruments, that are confined to a laboratory space, are not applicable to controling the loads a person is routinely subjected in free-living contexts.
+Measurements in these instruments, that are confined to a laboratory space, are not applicable to controling the loads a person is routinely subjected in free-living contexts.
 To surpass this issue, in the last few years, methods using wearable devices such as accelerometers have been proposed to measure mechanical loading, more precisely,  ground reaction force (GRF) [@Neugebauer2012; @Neugebauer2014; @Fortune2014; @Neugebauer2018; @Veras2020] and loading rate (LR) [@Fortune2014; @Veras2020].
@@
 In fact, due to this, several calibration studies have been done in several populations, as children [@Neugebauer2012], adults [@Neugebauer2014; @Fortune2014], military personnel [@Neugebauer2018], and subjects with obesity [@Veras2020].
-Another important factor that influence the predictions validity is the type of activities performed in the calibration study [@Bassett2012].
-Until now, studies that calibrate accelerometers to predict mechanical loading variables have employed usually activities as walking [@Fortune2014; @Neugebauer2018; @Veras2020], or walking and running [@Neugebauer2012; @Neugebauer2014].
+Another important factor that influences the predictions validity is the type of activities performed in the calibration study [@Bassett2012].
+Until now, studies that calibrate accelerometers to predict mechanical loading variables have usually employed activities as walking [@Fortune2014; @Neugebauer2018; @Veras2020], or walking and running [@Neugebauer2012; @Neugebauer2014].
 Nevertheless, activities eliciting higher impact loads have not yet been tested.
@@
 The accuracy indices yielded by applying these models have greatly worsened and these equations also produced some abnormal values, mainly in the pGRF model, in which it predicted negative values of force.
-These negative pGRF values were mainly as a result of a predictor, the second degree polynomial of the acceleration, possessing a negative coefficient.
+These negative pGRF values were mainly as a result of a predictor, the second-degree polynomial of the acceleration, possessing a negative coefficient.
 While this coefficient captured the plateauing of pGRF as pACC magnitude augments during walking, effectively improving the model [@Veras2020] for walking, this same pattern is not present in the pGRF vs.
@@
 First, the choice of activity to elicit high-impact loading was jumping, and several combinations of jump type and height were tested.
-Nevertheless, the results presented here cannot be assured for other high-impact activities, or even for other types of jump that were not tested.
+Nevertheless, the results presented here cannot be assured for other high-impact activities, or even for other types of jumps that were not tested.
 Second, no external sample for model validation was recruited.

### 2022-03-04 manuscript/manuscript.Rmd
Revise the manuscript

@@
 
-Bone reponds to mechanical loading by altering its properties, such as density, geometry, and strength [@Turner2003; @Fonseca2014; @Santos2017].
+Bone responds to mechanical loading by altering its properties, such as density, geometry, and strength [@Turner2003; @Fonseca2014; @Santos2017].
 Physical activity is a way of eliciting mechanical stimuli, and it is well established to improve bone health.
 However, there is no consensus in which characteristics of the physical activity-related mechanical stimulus are better associated with bone health benefits.
-Nevertheless, there is evidence that that high-impact exercise, although performed scarcely throughout the day, may play an important role in inducing positive bone adaptations [@Vainionpaa2006; @Stiles2017].
+Nevertheless, there is evidence that high-impact exercise, although performed scarcely throughout the day, may play an important role in inducing positive bone adaptations [@Vainionpaa2006; @Stiles2017].
 
@@
 Bone mechanical loading is usually measured in laboratory settings, using instruments such as force plates (FP) [@Medved2000].
-Measurements in these instruments, that are confined to a laboratory space, are not applicable to controling the loads a person is routinely subjected in free-living contexts.
+Measurements in these instruments, that are confined to a laboratory space, are not applicable in controlling the loads that a person is routinely subjected to in free-living contexts.
 To surpass this issue, in the last few years, methods using wearable devices such as accelerometers have been proposed to measure mechanical loading, more precisely,  ground reaction force (GRF) [@Neugebauer2012; @Neugebauer2014; @Fortune2014; @Neugebauer2018; @Veras2020] and loading rate (LR) [@Fortune2014; @Veras2020].
 These devices are light, small, can be used throughout days, and have a relatively affordable price.
-The aforementioned characteristics make the accelerometers well suited instruments to measure mechanical loading on a daily basis.
+These characteristics make the accelerometers well-suited instruments to measure mechanical loading on a daily basis.
 
 However, the validity of the accelerometer-predicted mechanical loading values depends on some factors, as the sample characteristics [@Welk2005].
-In fact, due to this, several calibration studies have been done in several populations, as children [@Neugebauer2012], adults [@Neugebauer2014; @Fortune2014], military personnel [@Neugebauer2018], and subjects with obesity [@Veras2020].
-Another important factor that influences the predictions validity is the type of activities performed in the calibration study [@Bassett2012].
-Until now, studies that calibrate accelerometers to predict mechanical loading variables have usually employed activities as walking [@Fortune2014; @Neugebauer2018; @Veras2020], or walking and running [@Neugebauer2012; @Neugebauer2014].
+In fact, due to this, several calibration studies have been done in several populations, such as children [@Neugebauer2012], adults [@Neugebauer2014; @Fortune2014], military personnel [@Neugebauer2018], and subjects with obesity [@Veras2020].
+Another important factor that influences the validity of the predictions is the type of activities performed in the calibration study [@Bassett2012].
+Until now, studies that calibrate accelerometers to predict mechanical loading variables have usually employed activities such as walking [@Fortune2014; @Neugebauer2018; @Veras2020], or walking and running [@Neugebauer2012; @Neugebauer2014].
 Nevertheless, activities eliciting higher impact loads have not yet been tested.
 
-Therefore, the objective of this study was to develop accelerometry-based mechanical loading prediction equations for high-impact activites in subjects from normal weight to severe obesity.
+Therefore, the objective of this study was to develop accelerometry-based mechanical loading prediction equations for high-impact activities in subjects from normal weight to severe obesity.
 
@@
 
-In this study, jumps were chosen as the high-impact activites to be performed.
-After familiarization, participants completed a series of trials envolving jumps of different types and heights.
+In this study, jumps were chosen as the high-impact activities to be performed.
+After familiarization, participants completed a series of trials involving jumps of different types and heights.
 First, drop jumps were executed from steps varying from 5cm to 40cm high, with 5cm increments.
-Then, box jumps were performed from the floor to 5cm, 15cm and 30cm high boxes.
-Finally, continuous jumps were carried out in two different heights: 5cm and 15cm.
+Then, box jumps were performed from the floor to 5cm, 15cm, and 30cm high boxes.
+Finally, continuous jumps were carried out at two different heights: 5cm and 15cm.
 Drop and box jumps were executed in two sets of four jumps at each height, with 30s interval, and the continuous jumps in one set of 20s at each height.
-The landing phase of the drop jumps, the take-off phase of the box jumps and both phases of the continuous jumps were recorded in FC (1000Hz; AMTI Corporation, Watertown, USA).
-Participants were instructed to jump with no restrictions to arms movement, land two-footed and then remain stationary on the force plate for five seconds (except in the continuous jumps).
+The landing phase of the drop jumps, the take-off phase of the box jumps, and both phases of the continuous jumps were recorded in FC (1000Hz; AMTI Corporation, Watertown, USA).
+Participants were instructed to jump with no restrictions to arms movement, land two-footed, and then remain stationary on the force plate for five seconds (except in the continuous jumps).
 
@@
 After that, both signals were adjusted using the time set by the systems clock and then synchronized by the maximum cross-correlation coefficient, with posterior manual adjustments if necessary.
-Next, peak ACC (pACC) were determined by the following criteria: i) a minimum height of three standard deviations above the mean ACC value recorded during a trial; and ii) a separation of at least 0.2 seconds for continuous jumps and 4 seconds for the other jumps.
-Peak GRF (pGRF) were then defined as the highest value between 0.2 or 4 seconds before and after each pACC, according to the jump type.
+Next, peak ACC (pACC) was determined by the following criteria: i) a minimum height of three standard deviations above the mean ACC value recorded during a trial; and ii) a separation of at least 0.2 seconds for continuous jumps and 4 seconds for the other jumps.
+Peak GRF (pGRF) was then defined as the highest value between 0.2 or 4 seconds before and after each pACC, according to the jump type.
 The rates of change were computed through a centered derivative from the beginning of the foot contact to the curve peak (Equation 1).
@@
 
-Linear mixed models were used to develop pGRF and pLR prediction equations for each vector (resultant and vertical) and accelerometer placement (ankle, lower back and  hip).
-Body mass was tested as predictor in all models, along with the acceleration magnitude (pACC) in the pGRF models and its rate (pAR) in the pLR models.
+Linear mixed models were used to develop pGRF and pLR prediction equations for each vector (resultant and vertical) and accelerometer placement (ankle, lower back, and  hip).
+Body mass was tested as a predictor in all models, along with the acceleration magnitude (pACC) in the pGRF models and its rate (pAR) in the pLR models.
 These predictors were entered as fixed effects and kept in the model, as all have shown to be significant (*p* < 0.05).
@@
 Lastly, our pGRF and pLR prediction models were compared with a previously published equation by our group [@Veras2020].
-As, for the best of our knowledge, there is no mechanical loading prediction equation developed for jumping activies, we chose to compare it with previous models from our group that were validated in a similar sample.
+As, for the best of our knowledge, there is no mechanical loading prediction equation developed for jumping activities, we chose to compare it with previous models from our group that were validated in a similar sample.
 The models used in the comparison were the ones predicting pGRF and pLR from both the resultant and vertical vectors of accelerometers worn at lower back and hip.
@@
 This variety of jump trials produced a varying response in both the FP- and accelerometer-measured variables.
-These responses in pGRF, pLR, pACC and pAR, regarding their resultant vector, can be observed in Figure 1 (see Supplemental Figure S1 for the vertical vector).
+These responses in pGRF, pLR, pACC, and pAR, regarding their resultant vector, can be observed in Figure 1 (see Supplemental Figure S1 for the vertical vector).
 Panels A and B show the FP-measured variables normalized by the subjects body mass, the pGRF and pLR, respectively.
 It can be seen that in the drop jump trials, there is a tendency to increase the loading magnitude the higher the jump, and in the box jumps this tendency is lower.
-As for the continuous jumps, both assayed heights elicit resonably constant loadings.
+As for the continuous jumps, both assayed heights elicit reasonably constant loadings.
 The same pattern can be observed in the accelerometer-measured variables, pACC and pAR, in Panels C and D, respectively.
@@
 Accelerometer-based mechanical loading prediction equations were developed based on linear mixed models.
-The regression coefficients, R^2^ and accuracy indices of the models derived from accelerometers worn at ankle, lower back and hip, and from the resultant and vertical vectors are presented in Table 1.
+The regression coefficients, R^2^ and accuracy indices of the models derived from accelerometers worn at ankle, lower back, and hip, and from the resultant and vertical vectors are presented in Table 1.
 The R^2^ values for the pGRF models ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`, showing that the models can explain at least `r round(min(GRF_accuracy$R2_cond), 1) * 100`\% of the pGRF variance.
@@
 Nevertheless, these MAPE values are similar to the results usually found in accelerometer-based energy expenditure prediction [@Diniz-Sousa2020], which have widespread use in research.
-Despite similar accuracy results achieved by all equations, data from the resultant vector of accelerometers at the hip have shown the lower accuracy error, represented by the smallest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%) and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
+Despite similar accuracy results achieved by all equations, data from the resultant vector of accelerometers at the hip have shown the lower accuracy error, represented by the smallest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%), and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
 Compared to the pGRF models, the pLR equations yielded poorer results, which is expected based on past research [@Fortune2014; @Veras2020].
 Their R^2^ ranged from `r min(LR_accuracy$R2_cond)` to `r max(LR_accuracy$R2_cond)` and MAPE from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%.
-However, similar to the pGRF models, the best pLR prediction equation in terms of prediction accuracy was the one from the resultant vector of hip-worn accelerometers, with a MAE of `r myround(hip_LR_accuracy$MAE, 1)` N, MAPE of `r myround(hip_LR_accuracy$MAPE * 100, 1)`\% and RMSE of `r myround(hip_LR_accuracy$RMSE, 1)` N.
+However, similar to the pGRF models, the best pLR prediction equation in terms of prediction accuracy was the one from the resultant vector of hip-worn accelerometers, with a MAE of `r myround(hip_LR_accuracy$MAE, 1)` N, MAPE of `r myround(hip_LR_accuracy$MAPE * 100, 1)`\%, and RMSE of `r myround(hip_LR_accuracy$RMSE, 1)` N.
 
@@
 
-To visualize the agreement between the pGRF and pLR values measured by the FP and predicted by the models, a series of Bland-Altman plots were built for both the resultant and vetical vectors and for all accelerometer placements.
+To visualize the agreement between the pGRF and pLR values measured by the FP and predicted by the models, a series of Bland-Altman plots were built for both the resultant and vertical vectors and for all accelerometer placements.
 Figure 2 shows the Bland-Altman plots for the pGRF (Panel A) and pLR (Panel B) resultant vector of hip-worn accelerometers.
@@
 Also, as could be observed by the one-sample *t*-tests results, none of the equations presented a bias different than zero (*p* > 0.05).
-Furthermore, all of the equations have shown a proportional bias, i.e., the magnitude of the values influence the prediction error.
+Furthermore, all of the equations have shown a proportional bias, i.e., the magnitude of the values influences the prediction error.
 However, as the linear regressions presented a very small R^2^ (maximum value of `r myround(max(ba_regression_ankle_R2), 2)`) this effect is neglectable.
@@
 
-Although the accuracy of our equations showed to be beyond the established 8\% threshold, when applying a model not developed for jumping activies we can se that the accuracy tremendously decreases, reaching values of MAPE around `r floor(mean(as.numeric(substr(tab2_df$MAPE, 1, nchar(tab2_df$MAPE) - 2))[1:4]) / 50) * 50`\% for pGRF prediction and around `r floor(mean(as.numeric(substr(tab2_df$MAPE, 1, nchar(tab2_df$MAPE) - 2))[5:8]) / 10) * 10`\% for pLR prediction as can be observed in Table 2.
-Furthermore, applying to our accelerometer-recorded jumping data a model that is not specific for this activity yield unresonable results, such as the negative values for pGRF that can be observed in Figure 3, Panel A.
+Although the accuracy of our equations showed to be beyond the established 8\% threshold, when applying a model not developed for jumping activities, we can see that the accuracy tremendously decreases, reaching values of MAPE around `r floor(mean(as.numeric(substr(tab2_df$MAPE, 1, nchar(tab2_df$MAPE) - 2))[1:4]) / 50) * 50`\% for pGRF prediction and around `r floor(mean(as.numeric(substr(tab2_df$MAPE, 1, nchar(tab2_df$MAPE) - 2))[5:8]) / 10) * 10`\% for pLR prediction as can be observed in Table 2.
+Furthermore, applying to our accelerometer-recorded jumping data a model that is not specific for this activity yields unreasonable results, such as the negative values for pGRF that can be observed in Figure 3, Panel A.
 
-Finally, it is important to mention that an R package was developed [@Veras2021] to make easier to apply all the necessary data processing steps to use the prediction equations.
+Finally, it is important to mention that an R package was developed [@Veras2021] to make it easier to apply all the necessary data processing steps to use the prediction equations.
 The package website (https://lveras.com/impactr) has all the installation instructions and documentation necessary to apply this method on your own data.
@@
 
-This study aimed to develop accelerometer-based mechanical loading predition equations for jumps of diverse types and heights, testing different accelerometer placements and using both the resultant and vertical vectors.
-The results have shown that jumping pGRF prediction models based on accelerometry data have a lower accuracy compared to models developed for other activies, such as walking and running.
+This study aimed to develop accelerometer-based mechanical loading prediction equations for jumps of diverse types and heights, testing different accelerometer placements and using both the resultant and vertical vectors.
+The results have shown that jumping pGRF prediction models based on accelerometry data have a lower accuracy compared to models developed for other activities, such as walking and running.
 Despite that, these models are the correct choice for this kind of activity eliciting hight levels of impact, as applying a model developed for lower impact activities in a sample similar to the current study yielded far worse results in terms of prediction accuracy.
@@
 Bone is known to increase its tissue mass and enhance its micro-architecture arrangement in response to mechanical stimulation [@Turner2003; @Fonseca2014], particularly at higher magnitudes [@Vainionpaa2006; @Stiles2017].
-This type of high-intensity impact activies have shown to be more associated with bone health improval, even if they are usually performed in low quantities throughout the day.
+This type of high-intensity impact activities have shown to be more associated with bone health improvements, even if they are usually performed in low quantities throughout the day.
 In fact, Vainionpää et al [@Vainionpaa2006] demonstrated that performing a relatively low number (<100) of pACC above 4.9*g* daily is positively associated with increases in bone mineral density.
 Usually, this acceleration magnitude can be achieved mostly by ballistic movements such as jumps [@Vainionpaa2006; @Rowlands2012].
-Nevertheless, accelerometry-based mechanical loading regression models have mostly been validated with a set of activies, such as walking and running, that elict relatively low acceleration magnitudes, usually ranging from approximately 1.2 to 6*g* [@Neugebauer2014; @Veras2020].
+Nevertheless, accelerometry-based mechanical loading regression models have mostly been validated with a set of activities, such as walking and running, that elicit relatively low acceleration magnitudes, usually ranging from approximately 1.2 to 6*g* [@Neugebauer2014; @Veras2020].
 
@@
 
-Loading rate is a highly sensitive parameter, that is usually calculated by deriving the force from the beggining of the foot contact to the floor to the point where the pGRF is achieved [@Liikavainio2007; @Fortune2014; @Veras2020].
-This is a small time window, usually less than 0.2 seconds for the jumps that were performed in our protocol, and requires a fine sampling acquisition that captures small graduations on the force values throughout time, which is difficult to achieve with common accelerometer sampling frequecies, that are usually 100Hz at maximum.
-Based on this and on on results from previous investigations [@Fortune2014; @Veras2020], pLR prediction models are expected to have a lower accuracy compared to the pGRF equations.
+Loading rate is a highly sensitive parameter, that is usually calculated by deriving the force from the beginning of the foot contact to the floor to the point where the pGRF is achieved [@Liikavainio2007; @Fortune2014; @Veras2020].
+This is a small time window, usually less than 0.2 seconds for the jumps that were performed in our protocol, and requires a fine sampling acquisition that captures small graduations on the force values throughout time, which is difficult to achieve with common accelerometer sampling frequencies, that are usually 100Hz at maximum.
+Based on this and on results from previous investigations [@Fortune2014; @Veras2020], pLR prediction models are expected to have a lower accuracy compared to the pGRF equations.
 Nevertheless, although our pLR prediction models had a worse accuracy compared with the pGRF models, its accuracy was similar or even better than the results from previous studies that developed models to predict pLR during walking only [@Fortune2014; @Veras2020].
 
-In our protocol, three accelerometer placements (ankle, lower back and hip) and two vectors (resultant and vertical) have been tested.
-Overall, the prediction models performance was reasonably consistent among the different placement and vector combinations, however, some conclusions can be draw.
-First, the hip accelerometer placement, that has traditionally been widely employed in research for the direct measure of physical activity [@Migueles2017], yielded a sligthly better accuracy compared the other placements tested.
+In our protocol, three accelerometer placements (ankle, lower back, and hip) and two vectors (resultant and vertical) have been tested.
+Overall, the prediction models performance was reasonably consistent among the different placement and vector combinations, however, some conclusions can be drawn.
+First, the hip accelerometer placement, that has traditionally been widely employed in research for the direct measure of physical activity [@Migueles2017], yielded a slightly better accuracy compared the other placements tested.
 This could be useful, as previous data collected from hip-worn accelerometers, even for other purposes, can be adequately used to obtain mechanical loading predictions.
@@
 Second, as the magnitude of the resultant vector of both pGRF and pLR, in the activities that were tested, is derived mostly from the vertical vector, and there were no considerable differences in performance between models using both vectors, they can be used interchangeably to predict mechanical loading.
-Besides that, while models using data from the vertical vector allow the use uniaxial accelerometers, data derived from the resultant vector have the important advantage of not depending on the correct acceleromter axes orientation, that cannot be guaranteed in free-living data collection [@Welk2005].
+Besides that, while models using data from the vertical vector allow the use of uniaxial accelerometers, data derived from the resultant vector have the important advantage of not depending on the correct accelerometer axes orientation, which cannot be guaranteed in free-living data collection [@Welk2005].
 
@@
 For instance, these models, when used in combination with other prediction equations developed for lower impact activities [@Veras2020], can be used to fully analyse the mechanical loading spectrum that people with different characteristics are submitted to daily and relate to bone health parameters.
-This would allow a more complete understanding on how mechanical loading influence bone healh, and also allow to identify which loading volume and magnitude range, for example, are more efficient in inducing bone health benefits.
+This would allow a more complete understanding on how mechanical loading influences bone health, and also allow to identify which loading volume and magnitude range, for example, are more efficient in inducing bone health benefits.
 Also, by the development and distribution of an R package that automates the steps necessary to apply the models that were proposed was made to enhance its applicability [@Veras2021].
@@
 
-Mechanical loading prediction equations based on accelerometer data have already been proposed for several populations [@Neugebauer2012; @Neugebauer2018; @Veras2020] and using accelerometer in different body placements [@Fortune2014; @Veras2020].
-What these past few studies have in common is that they validated their models using locomotion activities, as walking only or walking and runnning.
-The current study, by setting means to predict mechanical loading in several types and heights of jumps, broadens the application of such models, getting a step closer of the ability to accurately identify daily-living mechanical loading due to physical activity.
+Mechanical loading prediction equations based on accelerometer data have already been proposed for several populations [@Neugebauer2012; @Neugebauer2018; @Veras2020] and using an accelerometer in different body placements [@Fortune2014; @Veras2020].
+What these past few studies have in common is that they validated their models using locomotion activities, such as walking only or walking and running.
+The current study, by setting means to predict mechanical loading in several types and heights of jumps, broadens the application of such models, getting a step closer to the ability to accurately identify daily-living mechanical loading due to physical activity.
 This ability would allow the monitoring of loading in different contexts and to more fully understand the relationship between mechanical stimulation and its consequent effect on bone health.

### 2022-10-31 manuscript/manuscript.Rmd
Revise manuscript

@@
 
-Bone responds to mechanical loading by altering its properties, such as density, geometry, and strength [@Turner2003; @Fonseca2014; @Santos2017].
-Physical activity is a way of eliciting mechanical stimuli, and it is well established to improve bone health.
-However, there is no consensus in which characteristics of the physical activity-related mechanical stimulus are better associated with bone health benefits.
-Nevertheless, there is evidence that high-impact exercise, although performed scarcely throughout the day, may play an important role in inducing positive bone adaptations [@Vainionpaa2006; @Stiles2017].
+Bone responds to mechanical loading by altering its properties, such as mass, microarchitecture, and geometry, which ultimately dictate its strength and resistance to fracture [@Turner2003; @Fonseca2014; @Santos2017].
+In fact, physical activity is a paramount element to maintain and improve bone health throughout the life course [@Kohrt2004].
+However, the mechanical loading osteogenic effect is highly dependent of the physical activity characteristics [@Burr2002].
+Several animal studies [@Turner1998; @Turner2003; @Turner2005] helped to form our understanding of the principles that drive the bone tissue adaptative response.
+In general terms, high-impact dynamic loading, applied at high frequencies and including rest periods seems to favor increases in bone strength [@Turner2003].
+This seems to corroborate with the available evidence from humans.
+For instance, findings from different types of sports have shown that athletes practicing modalities involving high- and/or odd-impact (e.g., volleyball and football) have a greater bone strength than compared to repetitive and low-impact sports (e.g., cycling and swimming) [@Nikander2010].
+Nevertheless, the precise way in which the manipulation of mechanical loading variables influences the bone response in humans is not fully understand, particularly due to a difficulty in accurately assessing bone mechanical loading and its components.
 
-But, even when knowing precisely which the best loading regimen is to improve bone health, the question of how to measure and monitor this load remains.
-Bone mechanical loading is usually measured in laboratory settings, using instruments such as force plates (FP) [@Medved2000].
-Measurements in these instruments, that are confined to a laboratory space, are not applicable in controlling the loads that a person is routinely subjected to in free-living contexts.
-To surpass this issue, in the last few years, methods using wearable devices such as accelerometers have been proposed to measure mechanical loading, more precisely,  ground reaction force (GRF) [@Neugebauer2012; @Neugebauer2014; @Fortune2014; @Neugebauer2018; @Veras2020] and loading rate (LR) [@Fortune2014; @Veras2020].
-These devices are light, small, can be used throughout days, and have a relatively affordable price.
-These characteristics make the accelerometers well-suited instruments to measure mechanical loading on a daily basis.
+Some questionnaires have been developed to capture bone-stimulating characteristics of specific activities [@Groothausen1997; @Weeks2008], but, although the scores generated by such subjective methods seem to be related to bone status, this association appears to be limited and inconsistent across studies [@Farr2011; @Kim2016; @Kim2018].
+A more robust approach to assess mechanical loading is the use of equipment such as force plates, the gold standard for ground reaction forces measurement [@Medved2000].
+However, the use of force plates is less applicable in some clinical and research settings, as it is an immobile and expensive device, limited to laboratory conditions.
+In the last few yers, objective methods based on wearable devices have been proposed as a solution to overcome these limitations, using acceleration data to estimate mechanical loading components, typically assessed only through force plates, namely peak ground reaction forces (pGRF) [@Neugebauer2012; @Neugebauer2014; @Fortune2014; @Neugebauer2018; @Veras2020; @Veras2022] and peak loading rate (pLR) [@Fortune2014; @Veras2020; @Veras2022].
+Although accelerometry-based equations have shown a high prediction accuracy [@Veras2020; @Veras2022], this versatile mechanical loading assessment approach is not widely used in either clinical practice or research yet.
+One of the issues that may contribute to this fact is the lack of validation studies covering a broader spectrum of activities, namely those known to have a greater osteogenic effect, such as high-impact activities.
 
-However, the validity of the accelerometer-predicted mechanical loading values depends on some factors, as the sample characteristics [@Welk2005].
-In fact, due to this, several calibration studies have been done in several populations, such as children [@Neugebauer2012], adults [@Neugebauer2014; @Fortune2014], military personnel [@Neugebauer2018], and subjects with obesity [@Veras2020].
-Another important factor that influences the validity of the predictions is the type of activities performed in the calibration study [@Bassett2012].
-Until now, studies that calibrate accelerometers to predict mechanical loading variables have usually employed activities such as walking [@Fortune2014; @Neugebauer2018; @Veras2020], or walking and running [@Neugebauer2012; @Neugebauer2014].
-Nevertheless, activities eliciting higher impact loads have not yet been tested.
-
-Therefore, the objective of this study was to develop accelerometry-based mechanical loading prediction equations for high-impact activities in subjects from normal weight to severe obesity.
+Until now, studies that calibrate accelerometers to predict mechanical loading variables have usually employed only simple activities such as walking [@Fortune2014; @Neugebauer2018; @Veras2020] or walking and running [@Neugebauer2012; @Neugebauer2014; @Veras2022], while activities eliciting higher impact loads have not yet been tested.
+The fact that these accelerometry-based prediction equations are only valid for a limited set of activities hinders their ability to accurately quantify the amount of loading a person is submitted to daily, as the physical activities performed by a subject are often composed of other types of activity than just walking and running.
+Therefore, the validity of these methods should be expanded to other activities that are commonly used to induce positive bone adaptations.
+Jumping exercises, for instance, are frequently used as a strategy to improve bone health in post-menopausal women [@Du2021] and elder men [@Allison2015].
+The direct measurement of the mechanical loading involved in these activities would greatly improve our ability to precisely control the loading regimen elicited by these physical exercises and to establish a more reliable relationship between mechanical stimulation and bone anabolic response and, consequently, to establish a clearer relationship between exercise dose and the extent of the bone adaptive response [@Hansford2022].
+Therefore, the objective of this study was to develop accelerometry-based mechanical loading prediction equations for high-impact activities in adult subjects with a broad range of body masses, from normal weight to severe obesity.
 
@@
 No participant reported any musculoskeletal limitations.
-The measurement of height (stadiometer model 213; Seca, Hamburg, Germany) and body mass (digital scale model 899; Seca, Hamburg, Germany) was accomplished according to standard practice.
+The measurement of height (stadiometer model 213; Seca, Hamburg, Germany) and body mass (digital scale model 899; Seca, Hamburg, Germany) was accomplished according to standard procedures.
 
-In this study, jumps were chosen as the high-impact activities to be performed.
-After familiarization, participants completed a series of trials involving jumps of different types and heights.
+In this study, jumps were chosen as the high-impact activity to be performed.
+After familiarization with the protocol, participants completed a series of trials involving jumps of different types and heights.
 First, drop jumps were executed from steps varying from 5cm to 40cm high, with 5cm increments.
-Then, box jumps were performed from the floor to 5cm, 15cm, and 30cm high boxes.
+Then, box jumps were performed starting from the floor up to 5cm, 15cm, and 30cm high boxes.
 Finally, continuous jumps were carried out at two different heights: 5cm and 15cm.
 Drop and box jumps were executed in two sets of four jumps at each height, with 30s interval, and the continuous jumps in one set of 20s at each height.
-The landing phase of the drop jumps, the take-off phase of the box jumps, and both phases of the continuous jumps were recorded in FC (1000Hz; AMTI Corporation, Watertown, USA).
-Participants were instructed to jump with no restrictions to arms movement, land two-footed, and then remain stationary on the force plate for five seconds (except in the continuous jumps).
+The landing phase of the drop jumps, the take-off phase of the box jumps, and both phases of the continuous jumps were recorded in a force plate (1000Hz; AMTI Corporation, Watertown, USA).
+Participants were instructed to jump with no restrictions regarding arms movements, to land two-footed and to remain stationary on the force plate for five seconds after landing (except for the continuous jumps).
 
 Throughout the protocol, three activity monitors (GT9X Link; 100Hz; ±16g range; ActiGraph, Pensacola, USA) were worn by each participant, at the following placements: i) at their right hip (along the anterior axillary line, at the level of the iliac crest); ii) at their lower back (at the midpoint between the two posterior superior iliac spines); and iii) at their right ankle (immediately superior to the lateral malleolus).
-Monitors worn at hip and lower back were positioned on the same elastic belt with clips, and the ankle-worn monitor was fixed by an elastic belt and adhesive tape.
-Activity monitors were always positioned to ensure the alignment of the accelerometer vertical axis and the standing body craniocaudal axis.
+Monitors worn at the hip and lower back were positioned on the same elastic belt with clips, and the ankle-worn monitor was fixed by an elastic belt and adhesive tape.
+Activity monitors were always positioned to ensure the alignment of the accelerometer vertical axis and the standing body longitudinal axis.
+The hip and lower back placements were chosen due to their proximity to the body center of mass, and the ankle placement was chosen as its location is closer to the ground contact point.
 
@@
 As the manufacturer proprietary filter is not applied on the secondary accelerometer raw data, only data from this accelerometer were used in this study, to improve data processing replicability.
-Data from FP and accelerometers were operated through the manufacturer supplied software (for the FP: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA) and exported as raw data from the *x*, *y* and *z* vectors.
-GRF was expressed in N and acceleration (ACC) in gravitational ACC units (1*g* = 9.807 m·s^-2^).
+Data from force plates and accelerometers were operated through the manufacturer supplied software (for the force plates: Netforce, Version 3.5.1; AMTI Corporation, Watertown, USA; for the accelerometers: ActiLife version 6.13.3; Actigraph, Pensacola, USA) and exported as raw data from the x, y and z vectors.
+GRF was expressed in Newton (N) and acceleration in gravitational acceleration units (1*g* = 9.807 m·s^-2^).
 
 The MATLAB software (Version 2019a, Mathworks, Natick, USA) was used to process data from both FP and accelerometers.
-Data were processed according to the following procedures: first, the GRF signal was resampled to match the activity monitors sampling frequency (100Hz).
-Then, both GRF and ACC signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, to attenuate the noise magnitude.
+Data were processed according to the following procedures: first, the ground reaction force signal was resampled to match the activity monitors sampling frequency (100Hz).
+Then, both ground reaction force and acceleration signals were filtered using a Butterworth fourth-order low-pass filter, with 20 Hz cut-off frequency, to attenuate the noise magnitude.
 This cut-off frequency was selected based on visual inspection of the signals frequency spectrum after a Fast Fourier Transform (FFT).
-Afterwards, the GRF and ACC resultant vector was calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$).
+Afterwards, the ground reaction force and acceleration resultant vector was calculated ($r_i = \sqrt{x_i^2 + y_i^2 + z_i^2}$).
 After that, both signals were adjusted using the time set by the systems clock and then synchronized by the maximum cross-correlation coefficient, with posterior manual adjustments if necessary.
-Next, peak ACC (pACC) was determined by the following criteria: i) a minimum height of three standard deviations above the mean ACC value recorded during a trial; and ii) a separation of at least 0.2 seconds for continuous jumps and 4 seconds for the other jumps.
-Peak GRF (pGRF) was then defined as the highest value between 0.2 or 4 seconds before and after each pACC, according to the jump type.
+Next, peak acceleration (pACC) was determined by the following criteria: i) a minimum height of three standard deviations above the mean ACC value recorded during a trial; and ii) a separation of at least 0.2 seconds for continuous jumps and 4 seconds for the other jumps.
+Then, pGRF was then defined as the highest value between 0.2 or 4 seconds before and after each pACC, according to the jump type.
 The rates of change were computed through a centered derivative from the beginning of the foot contact to the curve peak (Equation 1).
-The peak LR (pLR) and peak acceleration rate (pAR) were defined as the maximum value in the array (Equation 2).
+The pLR and peak acceleration rate (pAR) were defined as the maximum value in the array (Equation 2).
 
@@
 
-Finally, pGRF, pACC, pLR and pAR means of the resultant and its vertical component for each participant at each jump type and height for each accelerometer placement were extracted and used in all remaining analyses.
+Finally, pGRF, pACC, pLR and pAR means of the resultant vector and its vertical component for each participant at each jump type and height for each accelerometer placement were extracted and used in all remaining analyses.
 
 Statistical analyses were performed through the R statistical software (version `r paste(R.Version()$major, R.Version()$minor, sep = ".")`, R Foundation for Statistical Computing, Vienna, Austria).
-All code used to generate the results can be found in an open platform (https://github.com/verasls/jumping_GRF_ACC)(ref).
+All code used to generate the results can be found in an open platform (https://bit.ly/3szNSJS).
 Statistical significance was set as $\alpha$ = 0.05.
@@
 Body mass was tested as a predictor in all models, along with the acceleration magnitude (pACC) in the pGRF models and its rate (pAR) in the pLR models.
-These predictors were entered as fixed effects and kept in the model, as all have shown to be significant (*p* < 0.05).
-Subject and the interaction between jump type and height were tested as random effects, and both factors have shown to improve the models.
+These predictors were entered as fixed effects and kept in the model, as all were shown to be significant (*p* < 0.05).
+Subject and the interaction between jump type and jump height were tested as random effects, and both factors were shown to improve the models.
 Final models were chosen according to the -2 log-likelihood statistic [@Field2012].
@@
 Model validation was performed using the leave-one-out cross-validation [@Staudenmayer2012].
-This method consists in separating data from one of the subjects into a testing dataset, and keeping the remainder in the training dataset.
-Then, a new model is developed using the training dataset, with the same parameters that were defined using the whole sample.
+This method consists in separating data from one of the subjects into a testing dataset and keeping the remainder in the training dataset.
+Then, a new model is developed using the training dataset, with the same parameters that were defined using the entire sample.
 This new model is then applied to predict the outcome for the subject in the testing dataset.
@@
 The numerical assessment was made by computing the following accuracy indices: the mean absolute error (MAE), mean absolute percent error (MAPE) and root mean square error (RMSE).
-Despite the absence of a standard criterion for establishing an acceptable error in mechanical loading prediction, we regarded a MAPE <8\% to be an accurate prediction based on past research [@Fortune2014; @Neugebauer2014; @Ancillao2018, @Veras2020].
-To visualize the agreement between the values predicted by the models and measured by the FP, Bland-Altman plots were drawn.
-This type of plot is made by plotting the mean of the actual and predicted values on the x axis and their difference on the y axis.
+To visually analyze the agreement between the values predicted by the models and measured by the force plates, Bland-Altman plots were drawn.
+These plots are constructed by plotting the mean of the actual and predicted values on the x axis and their difference on the y axis.
 Then, the bias was calculated as the mean of these differences and plotted as a continuous horizontal line, and the limits of agreement were defined by bias ± 1.96 standard deviation and plotted as horizontal dashed lines [@Bland1986].
-To further test if there were any consistent under- or over-estimation of the values by the models, one-sample *t*-tests were done to check whether the bias were different than zero.
-Finally, linear regressions were employed to examine if the difference between actual and predicted values were affected by their magnitude, verifying, therefore, whether the prediction was constant throughout the magnitude range that was tested [@Giavarina2015].
-
-Lastly, our pGRF and pLR prediction models were compared with a previously published equation by our group [@Veras2020].
-As, for the best of our knowledge, there is no mechanical loading prediction equation developed for jumping activities, we chose to compare it with previous models from our group that were validated in a similar sample.
-The models used in the comparison were the ones predicting pGRF and pLR from both the resultant and vertical vectors of accelerometers worn at lower back and hip.
-The MAE, MAPE and RMSE were calculated using the values predicted using Veras et al.
-[@Veras2020] equation.
+To further test if there were any systematic under- or over-estimation of the values by the models, one-sample *t*-tests were done to check whether the bias were significantly different from zero.
+Finally, linear regressions were employed to examine if the difference between actual and predicted values were affected by their magnitude, verifying, therefore, whether the prediction was constant throughout the entire magnitude range [@Giavarina2015].
 
@@
 
-In our protocol, three distinct jump types were performed, in several different heights.
-This variety of jump trials produced a varying response in both the FP- and accelerometer-measured variables.
-These responses in pGRF, pLR, pACC, and pAR, regarding their resultant vector, can be observed in Figure 1 (see Supplemental Figure S1 for the vertical vector).
-Panels A and B show the FP-measured variables normalized by the subjects body mass, the pGRF and pLR, respectively.
-It can be seen that in the drop jump trials, there is a tendency to increase the loading magnitude the higher the jump, and in the box jumps this tendency is lower.
-As for the continuous jumps, both assayed heights elicit reasonably constant loadings.
-The same pattern can be observed in the accelerometer-measured variables, pACC and pAR, in Panels C and D, respectively.
-Also, regarding the influence of the accelerometer placement, in most of the trials accelerometers worn at hip and lower back produced approximately similar values, with a tendency to higher dispersion in data derived from lower back accelerometers.
+In our protocol, three distinct jump types were performed, each one using several different heights.
+This variety of jump trials produced a considerable heterogeneity in the response of both the force plates- and accelerometer-measured variables.
+The results obtained for pGRF, pLR, pACC, and pAR, regarding their resultant vector, can be observed in Figure 1 (see Supplemental Figure S1 for the vertical vector).
+Panels A and B show the force plates-measured variables normalized by the subject’s body mass, the pGRF and pLR, respectively.
+It can be seen that, in the drop jump trials, there is a tendency for the loading magnitude to increase as the jump height increases.
+This tendency is also observed for the box jumps, but with a lower magnitude.
+As for the continuous jumps, both assayed heights elicited reasonably constant loadings.
+Overall, a similar pattern can be observed in the accelerometer-measured variables, pACC and pAR, in panels C and D, respectively.
+Nevertheless, this pattern is less clear, particularly for the continuous jumps.
+Regarding the influence of the accelerometer placement on the variables collected, in most of the trial’s, accelerometers worn at the hip and lower back produced approximately similar values, with a tendency for higher dispersion in data derived from the lower back accelerometers.
 
@@
 ```
+Accelerometer-based mechanical loading prediction equations were developed based on linear mixed models.
+The regression coefficients, R^2^ and accuracy indices of the models derived from accelerometers worn at the ankle, lower back and hip, and from the resultant and vertical vectors are presented in Table 1.
+The R^2^ values for the pGRF models ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`, showing that all the models were able to explain at least `r round(min(GRF_accuracy$R2_cond), 1) * 100`\% of the pGRF variance, and the observed MAPE values were between `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% and `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%.
+Despite similar accuracy results achieved by all equations, the model derived from data from the resultant vector of accelerometers worn at the hip were shown to have the lower accuracy error, represented by the smallest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%), and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
 
-Accelerometer-based mechanical loading prediction equations were developed based on linear mixed models.
-The regression coefficients, R^2^ and accuracy indices of the models derived from accelerometers worn at ankle, lower back, and hip, and from the resultant and vertical vectors are presented in Table 1.
-The R^2^ values for the pGRF models ranged from `r min(GRF_accuracy$R2_cond)` to `r max(GRF_accuracy$R2_cond)`, showing that the models can explain at least `r round(min(GRF_accuracy$R2_cond), 1) * 100`\% of the pGRF variance.
-The observed MAPE values were between `r myround(min(GRF_accuracy$MAPE) * 100, 1)`\% and `r myround(max(GRF_accuracy$MAPE) * 100, 1)`\%, which falls outside of the proposed threshold of 8\%.
-Nevertheless, these MAPE values are similar to the results usually found in accelerometer-based energy expenditure prediction [@Diniz-Sousa2020], which have widespread use in research.
-Despite similar accuracy results achieved by all equations, data from the resultant vector of accelerometers at the hip have shown the lower accuracy error, represented by the smallest MAE (`r myround(hip_GRF_accuracy$MAE, 1)` N), MAPE (`r myround(hip_GRF_accuracy$MAPE * 100, 1)`\%), and RMSE (`r myround(hip_GRF_accuracy$RMSE, 1)` N).
-Compared to the pGRF models, the pLR equations yielded poorer results, which is expected based on past research [@Fortune2014; @Veras2020].
-Their R^2^ ranged from `r min(LR_accuracy$R2_cond)` to `r max(LR_accuracy$R2_cond)` and MAPE from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%.
-However, similar to the pGRF models, the best pLR prediction equation in terms of prediction accuracy was the one from the resultant vector of hip-worn accelerometers, with a MAE of `r myround(hip_LR_accuracy$MAE, 1)` N, MAPE of `r myround(hip_LR_accuracy$MAPE * 100, 1)`\%, and RMSE of `r myround(hip_LR_accuracy$RMSE, 1)` N.
+Compared to the pGRF models, the pLR equations yielded poorer results.
+Although their R^2^ ranged from `r min(LR_accuracy$R2_cond)` to `r max(LR_accuracy$R2_cond)`, error measures were high, with MAPE ranging from `r myround(min(LR_accuracy$MAPE) * 100, 1)`\% to `r myround(max(LR_accuracy$MAPE) * 100, 1)`\%.
+Nevertheless, similar to the pGRF models developed, the best pLR prediction equation in terms of prediction accuracy was the one from the resultant vector derived from hip-worn accelerometers, even thought error estimates were higher, with a MAE of `r myround(hip_LR_accuracy$MAE, 1)` N, MAPE of `r myround(hip_LR_accuracy$MAPE * 100, 1)`\%, and RMSE of `r myround(hip_LR_accuracy$RMSE, 1)` N.
 
@@
 
-To visualize the agreement between the pGRF and pLR values measured by the FP and predicted by the models, a series of Bland-Altman plots were built for both the resultant and vertical vectors and for all accelerometer placements.
-Figure 2 shows the Bland-Altman plots for the pGRF (Panel A) and pLR (Panel B) resultant vector of hip-worn accelerometers.
-Both plots have shown good agreement levels, as at least `r ba_agreement`\% of the data points are within the limits of agreement and data tended to cluster around zero.
-Bland-Altman plots for the other accelerometer placements and vectors have shown the same patterns described previously, and can be found in the Supplemental Figure S2.
-Also, as could be observed by the one-sample *t*-tests results, none of the equations presented a bias different than zero (*p* > 0.05).
+To visualize the agreement between the pGRF and pLR values measured by the force plates and predicted by the models, a series of Bland-Altman plots were built for both the resultant and vertical vectors and for all accelerometer placements.
+Figure 2 shows the Bland-Altman plots for the pGRF (panel A) and pLR (panel B) resultant vector of hip-worn accelerometers.
+Both plots have shown good agreement levels, as at least `r ba_agreement`\% of the data points are within the limits of agreement and the data tended to cluster around zero.
+Bland-Altman plots for the other accelerometer placements and vectors have also shown the same previously described patterns and can be found in Supplemental Figure S2.
+Also, as could be observed by the one-sample *t*-tests results, none of the equations presented a bias significantly different than zero (*p* > 0.05).
 Furthermore, all of the equations have shown a proportional bias, i.e., the magnitude of the values influences the prediction error.
-However, as the linear regressions presented a very small R^2^ (maximum value of `r myround(max(ba_regression_ankle_R2), 2)`) this effect is neglectable.
+However, as the linear regressions used to assess the proportional bias presented a very small R^2^ (maximum value of `r myround(max(ba_regression_ankle_R2), 2)`), this effect is neglectable.
 
-```{r accuracy_table2, results = "show"}
-source(here("tables", "tab2.R"))
-tab2
-```
-
-\begin{figure}[ht]
-	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
-	\centering
-	\begin{measuredfigure}
-		\includegraphics[width=\textwidth]{fig3.png}
-    \caption{A comparison of the relationship between peak acceleration and peak ground reaction force (Panel A) and peak acceleration rate and peak loading rate (Panel B) in three ways: i) as measured by the force plates (black round points); ii) predicted by the equations developed in the current study (blue triangle points); and iii) predicted by the equations developed for walking by Veras et al (red square points).
-Abbreviations: pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.}
-	\end{measuredfigure}
-\end{figure}
-
-Although the accuracy of our equations showed to be beyond the established 8\% threshold, when applying a model not developed for jumping activities, we can see that the accuracy tremendously decreases, reaching values of MAPE around `r floor(mean(as.numeric(substr(tab2_df$MAPE, 1, nchar(tab2_df$MAPE) - 2))[1:4]) / 50) * 50`\% for pGRF prediction and around `r floor(mean(as.numeric(substr(tab2_df$MAPE, 1, nchar(tab2_df$MAPE) - 2))[5:8]) / 10) * 10`\% for pLR prediction as can be observed in Table 2.
-Furthermore, applying to our accelerometer-recorded jumping data a model that is not specific for this activity yields unreasonable results, such as the negative values for pGRF that can be observed in Figure 3, Panel A.
-
-Finally, it is important to mention that an R package was developed [@Veras2021] to make it easier to apply all the necessary data processing steps to use the prediction equations.
-The package website (https://lveras.com/impactr) has all the installation instructions and documentation necessary to apply this method on your own data.
+Finally, an R package was also developed [@Veras2021] to simplify the application of all the necessary data processing steps involved in the implementation of the developed prediction equations.
+The package website (https://verasls.github.io/impactr) has all the installation instructions and documentation necessary to apply this method on the user’s own data.
 
@@
 
-This study aimed to develop accelerometer-based mechanical loading prediction equations for jumps of diverse types and heights, testing different accelerometer placements and using both the resultant and vertical vectors.
-The results have shown that jumping pGRF prediction models based on accelerometry data have a lower accuracy compared to models developed for other activities, such as walking and running.
-Despite that, these models are the correct choice for this kind of activity eliciting hight levels of impact, as applying a model developed for lower impact activities in a sample similar to the current study yielded far worse results in terms of prediction accuracy.
-Also, as expected, the pLR prediction models performed poorly compared to the pGRF equations.
+This study aimed to develop accelerometer-based mechanical loading prediction equations, namely for pGRF and pLR, for jumping exercises of different types and heights, as well as to test different accelerometer placements and data derived from both the resultant and vertical acceleration vectors.
+The results have shown that pGRF prediction models for jumping activities based on accelerometry data have a MAPE around 13%, while pLR prediction equations resulted in a lower accuracy, with a MAPE around 22%.
+Also, these models are the first to be developed specifically for high-impact jumping activities in adults.
 
-Bone is known to increase its tissue mass and enhance its micro-architecture arrangement in response to mechanical stimulation [@Turner2003; @Fonseca2014], particularly at higher magnitudes [@Vainionpaa2006; @Stiles2017].
-This type of high-intensity impact activities have shown to be more associated with bone health improvements, even if they are usually performed in low quantities throughout the day.
-In fact, Vainionpää et al [@Vainionpaa2006] demonstrated that performing a relatively low number (<100) of pACC above 4.9*g* daily is positively associated with increases in bone mineral density.
-Usually, this acceleration magnitude can be achieved mostly by ballistic movements such as jumps [@Vainionpaa2006; @Rowlands2012].
-Nevertheless, accelerometry-based mechanical loading regression models have mostly been validated with a set of activities, such as walking and running, that elicit relatively low acceleration magnitudes, usually ranging from approximately 1.2 to 6*g* [@Neugebauer2014; @Veras2020].
+Bone is known to increase its tissue mass and enhance its micro-architecture arrangement in response to mechanical stimulation [@Turner2003; @Fonseca2014], particularly of higher magnitudes [@Vainionpaa2006; @Stiles2017].
+These types of high-intensity impact activities have shown to be more associated with bone health improvements, even if they are usually performed in low volumes throughout the day.
+In fact, Vainionpää et al [@Vainionpaa2006] demonstrated that performing a relatively low number (<100) of pACC above 4.9g daily is positively associated with increases in bone mineral density.
+Usually, this acceleration magnitude can be achieved mostly by specific movements such as jumps [@Vainionpaa2006; @Rowlands2012] .
+Nevertheless, accelerometry-based mechanical loading regression models have mostly been validated for a set of activities, such as walking and running, that elicit relatively low acceleration magnitudes [@Neugebauer2014; @Veras2020; @Veras2022].
+The results of this study provide new mechanical loading prediction models for jumping, which are especially important as it is one of the most popular and effective activities prescribed for bone health improvement [@Beck2017].
 
-The use of a prediction model based on accelerometer data should ideally be done in conditions similar to the study that validated the model, specially in terms of sample characteristics and type and intensity of the activities performed [@Welk2005].
-Therefore, if we want to accurately predict the mechanical loading caused by high-impact activities, we should choose a prediction model that has been validated using these activities.
-Even though the cross-validation accuracy of such a model was poorer compared with other models developed for activities such as walking and running, as demonstrated by our results, the use of these models is crucial to achieve an acceptable prediction accuracy.
-It was demonstrated by testing a model previously developed by our group [@Veras2020] in a similar sample, but on walking, on the accelerometer-recorded jumping data from the current study.
-The accuracy indices yielded by applying these models have greatly worsened and these equations also produced some abnormal values, mainly in the pGRF model, in which it predicted negative values of force.
-These negative pGRF values were mainly as a result of a predictor, the second-degree polynomial of the acceleration, possessing a negative coefficient.
-While this coefficient captured the plateauing of pGRF as pACC magnitude augments during walking, effectively improving the model [@Veras2020] for walking, this same pattern is not present in the pGRF vs.
-pACC relationship during jumping, and, therefore, only impairs the accuracy of these equations when applyed to this kind of activity.
+The development of such models was done with the direct purpose of allowing to assess the loading sustained by the skeletal system during daily physical activity.
+Nowadays, physical activity monitorization and exercise prescription aiming bone health promotion has been carried out almost blindly or through the use of subjective instruments such as questionaries.
+Moreover, the absence of a trustworthy method to monitor mechanical loading on a daily basis may preclude the prescription and self-involvement of patients in high-impact activities that may expose them to a higher risk of joint injuries.
+These issues may be overcome with the recent popularization of wearable sensors, such as accelerometers, and their ability to be used as a tool for mechanical loading prediction.
+This would allow some important clinical applications, as the models developed in this study, when used in combination with other prediction equations for other types of activities [@Veras2022], can be used to fully analyze the mechanical loading spectrum that people are submitted to daily and relate to bone health parameters.
+Therefore, a more complete understanding can be achieved on how mechanical loading influences bone health, by, for instance, identifying the loading volume and magnitude range that are more efficient in inducing bone health benefits.
+Also, the development and distribution of an R package that automates the steps necessary to apply these models enhances their applicability to potential users [@Veras2021].
+This can prove to be an important tool, as the application of these models depend on some laborious steps, such as the application of digital filters to reduce noise from the signal, detection of the peaks in this signal and the computation of derivatives to obtain the acceleration rate, necessary for the pRL prediction.
 
-Loading rate is a highly sensitive parameter, that is usually calculated by deriving the force from the beginning of the foot contact to the floor to the point where the pGRF is achieved [@Liikavainio2007; @Fortune2014; @Veras2020].
-This is a small time window, usually less than 0.2 seconds for the jumps that were performed in our protocol, and requires a fine sampling acquisition that captures small graduations on the force values throughout time, which is difficult to achieve with common accelerometer sampling frequencies, that are usually 100Hz at maximum.
-Based on this and on results from previous investigations [@Fortune2014; @Veras2020], pLR prediction models are expected to have a lower accuracy compared to the pGRF equations.
+When comparing the accuracy indices of the pGRF prediction equations developed in the current study with other equations from past investigations, developed for walking and running, it can be seen that our models have a lower accuracy [@Neugebauer2012; @Neugebauer2014; @Neugebauer2018; @Veras2020; @Veras2022].
+This can, at least partially, be explained by the high heterogeneity in the jumping movement patterns, as several combinations of jump type and height were tested, and also that subjects with different skill levels were included in the sample.
+Nevertheless, these MAPE values are similar to the results usually found in accelerometer-based energy expenditure prediction [@Diniz-Sousa2020], which have widespread use in research.
+As for the pLR prediction equations, they presented a lower accuracy compared to the pGRF equations, which corroborates with previous findings [@Fortune2014; @Veras2020; @Veras2022].
+This is possibly due to the fact that loading rate is typically calculated through a derivative of the force signal from the beginning of the foot contact to the floor to the point where the pGRF is achieved [@Liikavainio2007; @Fortune2014; @Veras2020; @Veras2022].
+As it is a small time window, usually less than 0.2 seconds for the jumps that were performed in our protocol, it requires a fine sampling acquisition that captures small graduations on the force values throughout time, which is difficult to achieve with common accelerometer sampling frequencies, that are usually 100Hz at maximum.
 Nevertheless, although our pLR prediction models had a worse accuracy compared with the pGRF models, its accuracy was similar or even better than the results from previous studies that developed models to predict pLR during walking only [@Fortune2014; @Veras2020].
@@
 In our protocol, three accelerometer placements (ankle, lower back, and hip) and two vectors (resultant and vertical) have been tested.
-Overall, the prediction models performance was reasonably consistent among the different placement and vector combinations, however, some conclusions can be drawn.
-First, the hip accelerometer placement, that has traditionally been widely employed in research for the direct measure of physical activity [@Migueles2017], yielded a slightly better accuracy compared the other placements tested.
+Overall, the prediction models performance was reasonably consistent among the different placement and vector combinations, allowing some conclusions can be drawn.
+First, the hip accelerometer placement which, traditionally, has been widely employed in research for the direct measurement of physical activity [@Migueles2017], yielded a slightly better accuracy compared to the other placements tested.
 This could be useful, as previous data collected from hip-worn accelerometers, even for other purposes, can be adequately used to obtain mechanical loading predictions.
-Also, studies that aim to extract mechanical loading and other variables from accelerometry data, that use the hip as the default placement, can comfortably do so.
+Also, studies that aim to extract mechanical loading together with other variables derived from accelerometry data and that use the hip as the default accelerometer placement, can reliably use this placement.
 Second, as the magnitude of the resultant vector of both pGRF and pLR, in the activities that were tested, is derived mostly from the vertical vector, and there were no considerable differences in performance between models using both vectors, they can be used interchangeably to predict mechanical loading.
-Besides that, while models using data from the vertical vector allow the use of uniaxial accelerometers, data derived from the resultant vector have the important advantage of not depending on the correct accelerometer axes orientation, which cannot be guaranteed in free-living data collection [@Welk2005].
-
-The development of accelerometry-based mechanical loading prediction models was done with the direct purpose of allowing to assess the loading beared by the skeletal system during daily physical activity.
-This would allow some clinical applications.
-For instance, these models, when used in combination with other prediction equations developed for lower impact activities [@Veras2020], can be used to fully analyse the mechanical loading spectrum that people with different characteristics are submitted to daily and relate to bone health parameters.
-This would allow a more complete understanding on how mechanical loading influences bone health, and also allow to identify which loading volume and magnitude range, for example, are more efficient in inducing bone health benefits.
-Also, by the development and distribution of an R package that automates the steps necessary to apply the models that were proposed was made to enhance its applicability [@Veras2021].
-This can prove to be an important tool, as the application of these models depend on non-trivial data processing skills, such as applying digital filters to reduce noise from the signal, detection of the peaks in this signal and the computation of derivatives to obtain the acceleration rate, necessary for the prediction of pLR.
+Besides that, while models using data from the vertical vector allow the use of uniaxial accelerometers, data derived from the resultant vector have the important advantage of not depending on the correct accelerometer axes orientation, which cannot be guaranteed during data collection in free-living conditions [@Welk2005].
 
@@
 Second, no external sample for model validation was recruited.
-However, the leave-one-out cross-validation method that was used is the recommended in these situations [@Staudenmayer2012].
+However, the leave-one-out cross-validation method that was used is the recommended strategy in these situations [@Staudenmayer2012].
 Third, to apply these models in accelerometer-recorded free-living physical activity data, there is a need to identify, in the acceleration signal, the periods in which the activity being performed is jumping.
@@
 
-Mechanical loading prediction equations based on accelerometer data have already been proposed for several populations [@Neugebauer2012; @Neugebauer2018; @Veras2020] and using an accelerometer in different body placements [@Fortune2014; @Veras2020].
-What these past few studies have in common is that they validated their models using locomotion activities, such as walking only or walking and running.
-The current study, by setting means to predict mechanical loading in several types and heights of jumps, broadens the application of such models, getting a step closer to the ability to accurately identify daily-living mechanical loading due to physical activity.
-This ability would allow the monitoring of loading in different contexts and to more fully understand the relationship between mechanical stimulation and its consequent effect on bone health.
+In conclusion, accelerometry-based mechanical loading prediction models, especially for pGRF, have proved to be a valid method for determining the mechanical stimulation induced by jumps of various types and heights.
+These prediction models were validated using data from both the resultant and vertical vectors and with accelerometers placed at the ankle, lower back and hip.
+Results from this study will enable the continuous assessment of mechanical loading in clinical settings, providing means to objectively determine the osteogenic potential of daily physical activity and to better monitor and prescribe exercise aimed to improve bone health.
+Compared to pGRF, prediction of pLR showed lower accuracy.
+
+# Acknowledgements {.unnumbered}
+
+This study was funded by the Foundation for Science and Technology (FCT) [grant number PTDC/DTP-DES/0968/2014] and by the European Regional Development Fund (ERDF) through the Operational Competitiveness Programme (COMPETE) [grant number POCI-01-0145-FEDER-016707].
+The study was developed in the Research Centre in Physical Activity, Health and Leisure (CIAFEL) funded by ERDF through the COMPETE and by the FCT [grant number UID/DTP/00617/2020], and Laboratory for Integrative and Translational Research in Population Health (ITR), funded by the FCT [grant number LA/P/0064/2020].
+Lucas  Veras and  Giorjines Boppre are supported by the FCT [grant numbers UI/BD/150673/2020 and SFRH/BD/146976/2019].
+The authors thank the participants who took part in this research and all who have collaborated in the research project.
 
@@
 
-**Figure 3:** A comparison of the relationship between peak acceleration and peak ground reaction force (Panel A) and peak acceleration rate and peak loading rate (Panel B) in three ways: i) as measured by the force plates (black round points); ii) predicted by the equations developed in the current study (blue triangle points); and iii) predicted by the equations developed for walking by Veras et al (red square points).
-Abbreviations: pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.
-
 **Supplemental Figure S1:** Distribution of the vertical vector of mechanical loading (peak ground reaction force and peak loading rate) and acceleration (peak acceleration and peak acceleration rate) variables per jump type and height combination.

### 2022-10-31 figures/fig1.R
Add spaces between boxplots of different jump types

@@ -18,6 +18,34 @@ mechanical_load_data <- mechanical_load_data |>
     )
   )
 
+# Add new jump values to separate the boxplot by jump types
+mechanical_load_data <- mechanical_load_data |>
+  # filter(acc_placement == "Lower back" & vector == "resultant") |>
+  add_row(
+    acc_placement = rep(c("Ankle", "Lower back", "Hip"), 4),
+    vector = c(rep("resultant", 6), rep("vertical", 6)),
+    jump = rep(c(rep("1", 3), rep("2", 3)), 2),
+    .before = 1
+  ) |>
+  mutate(
+    jump = fct_relevel(
+      as.factor(jump),
+      levels = c(
+        "Drop jumps 5cm", "Drop jumps 10cm", "Drop jumps 15cm",
+        "Drop jumps 20cm", "Drop jumps 25cm", "Drop jumps 30cm",
+        "Drop jumps 35cm", "Drop jumps 40cm", "1",
+        "Box jumps 5cm", "Box jumps 15cm", "Box jumps 30cm", "2",
+        "Continuous jumps 5cm", "Continuous jumps 15cm"
+      )
+    ),
+    acc_placement = fct_relevel(
+      as.factor(acc_placement),
+      levels = c("Hip", "Lower back", "Ankle")
+    )
+  )
+
+levels(mechanical_load_data$acc_placement)
+
 # GRF boxplot -------------------------------------------------------------
 
 boxplot_GRF_res <- mechanical_load_data |>
@@ -43,6 +71,15 @@ boxplot_ACC_res <- mechanical_load_data |>
   ggplot(aes(x = jump, y = pACC_g)) +
   geom_boxplot(aes(fill = acc_placement), outlier.size = 0.8) +
   scale_fill_manual(values = c("gray90", "gray70", "gray50")) +
+  scale_x_discrete(
+    labels = c(
+      "Drop jumps 5cm", "Drop jumps 10cm", "Drop jumps 15cm",
+      "Drop jumps 20cm", "Drop jumps 25cm", "Drop jumps 30cm",
+      "Drop jumps 35cm", "Drop jumps 40cm", " ",
+      "Box jumps 5cm", "Box jumps 15cm", "Box jumps 30cm", " ",
+      "Continuous jumps 5cm", "Continuous jumps 15cm"
+    )
+  ) +
   scale_y_continuous(
     limits = c(0, 16),
[…]

### 2022-10-31 figures/figS1.R
Rename files

@@ -1,121 +1,0 @@
-# Load packages and functions ---------------------------------------------
-
-library(here)
-library(tidyverse)
-library(ragg)
-library(patchwork)
-
-# Load data ---------------------------------------------------------------
-
-load(here("data", "mechanical_load_data.rda"))
-mechanical_load_data <- mechanical_load_data |>
-  mutate(
-    acc_placement = recode_factor(
-      acc_placement,
-      hip = "Hip",
-      lower_back = "Lower back",
-      ankle = "Ankle"
-    )
-  )
-
-# GRF boxplot -------------------------------------------------------------
-
-boxplot_GRF_ver <- mechanical_load_data |>
-  filter(acc_placement == "Lower back" & vector == "vertical") |>
-  ggplot(aes(x = jump, y = pGRF_BW)) +
-  geom_boxplot() +
-  scale_y_continuous(
-    limits = c(0, 7),
-    expand = c(0, 0),
-    breaks = seq(0, 7, 1)
-  ) +
-  theme_light() +
-  theme(
-    axis.text.y = element_text(size = 13),
-    axis.title.y = element_text(size = 16)
-  ) +
-  labs(x = "", y = "pGRF (BW)")
-
-# ACC boxplots ------------------------------------------------------------
-
-boxplot_ACC_ver <- mechanical_load_data |>
-  filter(vector == "vertical") |>
-  ggplot(aes(x = jump, y = pACC_g)) +
-  geom_boxplot(aes(fill = acc_placement), outlier.size = 0.8) +
-  scale_fill_manual(values = c("gray90", "gray70", "gray50")) +
-  scale_y_continuous(
-    limits = c(0, 14),
-    expand = c(0, 0),
-    breaks = seq(0, 14, 2)
-  ) +
-  theme_light() +
-  theme(
-    axis.text.x = element_text(size = 13, angle = 45, hjust = 1),
-    axis.text.y = element_text(size = 13),
-    axis.title.y = element_text(size = 16),
-    legend.title = element_blank()
-  ) +
-  labs(x = "", y = quote("pACC" ~ (italic(g))))
-
-# LR boxplot --------------------------------------------------------------
-
-boxplot_LR_ver <- mechanical_load_data |>
-  filter(acc_placement == "Lower back" & vector == "vertical") |>
-  ggplot(aes(x = jump, y = pLR_BWs)) +
[…]

### 2022-10-31 figures/figS3.R

@@ -1,242 +1,121 @@
-# Load packages -----------------------------------------------------------
+# Load packages and functions ---------------------------------------------
 
 library(here)
 library(tidyverse)
-library(lvmisc)
-library(ggsci)
-library(patchwork)
 library(ragg)
+library(patchwork)
 
 # Load data ---------------------------------------------------------------
 
-load(here("output", "loocv_data.rda"))
+load(here("data", "mechanical_load_data.rda"))
+mechanical_load_data <- mechanical_load_data |>
+  mutate(
+    acc_placement = recode_factor(
+      acc_placement,
+      hip = "Hip",
+      lower_back = "Lower back",
+      ankle = "Ankle"
+    )
+  )
 
-# Ankle resultant pLR Bland-Altman plot -----------------------------------
+# GRF boxplot -------------------------------------------------------------
 
-BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
-  plot_bland_altman(color = BMI_cat, alpha = 0.5) +
-  scale_color_nejm() +
+boxplot_GRF_ver <- mechanical_load_data |>
+  filter(acc_placement == "Lower back" & vector == "vertical") |>
+  ggplot(aes(x = jump, y = pGRF_BW)) +
+  geom_boxplot() +
   scale_y_continuous(
-    labels = scales::label_number(),
-    limits = c(-120000, 120000),
+    limits = c(0, 7),
     expand = c(0, 0),
-    breaks = seq(-120000, 120000, 40000)
-  ) +
-  scale_x_continuous(
-    labels = scales::label_number(),
-    limits = c(0, 260000),
-    expand = c(0, 0),
-    breaks = seq(50000, 250000, 50000)
+    breaks = seq(0, 7, 1)
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
-    legend.title = element_blank(),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
     axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    axis.title.y = element_text(size = 16)
   ) +
-  guides(alpha = "none") +
-  labs(
-    title = "Resultant vector - Ankle placement",
[…]

### 2022-10-31 figures/figS5.R

@@ -1,0 +1,242 @@
+# Load packages -----------------------------------------------------------
+
+library(here)
+library(tidyverse)
+library(lvmisc)
+library(ggsci)
+library(patchwork)
+library(ragg)
+
+# Load data ---------------------------------------------------------------
+
+load(here("output", "loocv_data.rda"))
+
+# Ankle resultant pLR Bland-Altman plot -----------------------------------
+
+BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
+  plot_bland_altman(color = BMI_cat, alpha = 0.5) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    labels = scales::label_number(),
+    limits = c(-120000, 120000),
+    expand = c(0, 0),
+    breaks = seq(-120000, 120000, 40000)
+  ) +
+  scale_x_continuous(
+    labels = scales::label_number(),
+    limits = c(0, 260000),
+    expand = c(0, 0),
+    breaks = seq(50000, 250000, 50000)
+  ) +
+  theme_light() +
+  theme(
+    plot.title = element_text(size = 15, hjust = 0.5),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 12),
+    axis.title.y = element_text(size = 13),
+    axis.title.x = element_text(size = 13),
+    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 13)
+  ) +
+  guides(alpha = "none") +
+  labs(
+    title = "Resultant vector - Ankle placement",
+    x = quote("Mean of Actual and Predicted pLR" ~ (N %.% s^-1)),
+    y = quote("Actual - Predicted pLR" ~ (N %.% s^-1))
+  )
+
+# Lower back resultant pLR Bland-Altman plot ------------------------------
+
+BA_LR_res_back <- cv_res_LR_models$lower_back %>%
+  plot_bland_altman(color = BMI_cat, alpha = 0.5) +
+  scale_color_nejm() +
+  scale_y_continuous(
+    labels = scales::label_number(),
+    limits = c(-120000, 120000),
+    expand = c(0, 0),
+    breaks = seq(-120000, 120000, 40000)
+  ) +
+  scale_x_continuous(
+    labels = scales::label_number(),
+    limits = c(0, 270000),
+    expand = c(0, 0),
+    breaks = seq(50000, 250000, 50000)
+  ) +
+  theme_light() +
[…]

### 2022-10-31 figures/figS3.R
Update figure names

@@ -98,7 +98,7 @@ boxplot_AR_ver <- mechanical_load_data |>
 
 # Combine and save plots --------------------------------------------------
 
-figS1 <- boxplot_GRF_ver + theme(axis.text.x = element_blank()) +
+figS3 <- boxplot_GRF_ver + theme(axis.text.x = element_blank()) +
   boxplot_LR_ver + theme(axis.text.x = element_blank()) +
   boxplot_ACC_ver +
   boxplot_AR_ver +
@@ -110,12 +110,12 @@ figS1 <- boxplot_GRF_ver + theme(axis.text.x = element_blank()) +
   )
 
 agg_png(
-  here("figures", "figS1.png"),
+  here("figures", "figS3.png"),
   width = 90,
   height = 80,
   units = "cm",
   res = 100,
   scaling = 2
 )
-plot(figS1)
+plot(figS3)
 dev.off()

### 2022-10-31 figures/figS4.R

@@ -205,7 +205,7 @@ BA_GRF_ver_hip <- cv_ver_GRF_models$hip |>
 
 # Combine and save plots --------------------------------------------------
 
-figS2 <- BA_GRF_res_ankle +
+figS4 <- BA_GRF_res_ankle +
   BA_GRF_res_back +
   BA_GRF_res_hip +
   BA_GRF_ver_ankle +
@@ -219,12 +219,12 @@ figS2 <- BA_GRF_res_ankle +
   )
 
 agg_png(
-  here("figures", "figS2.png"),
+  here("figures", "figS4.png"),
   width = 120,
   height = 50,
   units = "cm",
   res = 100,
   scaling = 2
 )
-plot(figS2)
+plot(figS4)
 dev.off()

### 2022-10-31 figures/figS5.R

@@ -217,7 +217,7 @@ BA_LR_ver_hip <- cv_ver_LR_models$hip %>%
 
 # Combine and save plots --------------------------------------------------
 
-figS3 <- BA_LR_res_ankle +
+figS5 <- BA_LR_res_ankle +
   BA_LR_res_back +
   BA_LR_res_hip +
   BA_LR_ver_ankle +
@@ -231,12 +231,12 @@ figS3 <- BA_LR_res_ankle +
   )
 
 agg_png(
-  here("figures", "figS3.png"),
+  here("figures", "figS5.png"),
   width = 120,
   height = 50,
   units = "cm",
   res = 100,
   scaling = 2
 )
-plot(figS3)
+plot(figS5)
 dev.off()

### 2022-10-31 figures/fig1.R
Delete unnecessary code

@@ -20,7 +20,6 @@ mechanical_load_data <- mechanical_load_data |>
 
 # Add new jump values to separate the boxplot by jump types
 mechanical_load_data <- mechanical_load_data |>
-  # filter(acc_placement == "Lower back" & vector == "resultant") |>
   add_row(
     acc_placement = rep(c("Ankle", "Lower back", "Hip"), 4),
     vector = c(rep("resultant", 6), rep("vertical", 6)),
@@ -43,8 +42,6 @@ mechanical_load_data <- mechanical_load_data |>
       levels = c("Hip", "Lower back", "Ankle")
     )
   )
-
-levels(mechanical_load_data$acc_placement)
 
 # GRF boxplot -------------------------------------------------------------
 

### 2022-10-31 figures/figS3.R
Add spaces between boxplots of different jump types

@@ -18,6 +18,32 @@ mechanical_load_data <- mechanical_load_data |>
     )
   )
 
+# Add new jump values to separate the boxplot by jump types
+mechanical_load_data <- mechanical_load_data |>
+  # filter(acc_placement == "Lower back" & vector == "resultant") |>
+  add_row(
+    acc_placement = rep(c("Ankle", "Lower back", "Hip"), 4),
+    vector = c(rep("resultant", 6), rep("vertical", 6)),
+    jump = rep(c(rep("1", 3), rep("2", 3)), 2),
+    .before = 1
+  ) |>
+  mutate(
+    jump = fct_relevel(
+      as.factor(jump),
+      levels = c(
+        "Drop jumps 5cm", "Drop jumps 10cm", "Drop jumps 15cm",
+        "Drop jumps 20cm", "Drop jumps 25cm", "Drop jumps 30cm",
+        "Drop jumps 35cm", "Drop jumps 40cm", "1",
+        "Box jumps 5cm", "Box jumps 15cm", "Box jumps 30cm", "2",
+        "Continuous jumps 5cm", "Continuous jumps 15cm"
+      )
+    ),
+    acc_placement = fct_relevel(
+      as.factor(acc_placement),
+      levels = c("Hip", "Lower back", "Ankle")
+    )
+  )
+
 # GRF boxplot -------------------------------------------------------------
 
 boxplot_GRF_ver <- mechanical_load_data |>
@@ -43,6 +69,15 @@ boxplot_ACC_ver <- mechanical_load_data |>
   ggplot(aes(x = jump, y = pACC_g)) +
   geom_boxplot(aes(fill = acc_placement), outlier.size = 0.8) +
   scale_fill_manual(values = c("gray90", "gray70", "gray50")) +
+  scale_x_discrete(
+    labels = c(
+      "Drop jumps 5cm", "Drop jumps 10cm", "Drop jumps 15cm",
+      "Drop jumps 20cm", "Drop jumps 25cm", "Drop jumps 30cm",
+      "Drop jumps 35cm", "Drop jumps 40cm", " ",
+      "Box jumps 5cm", "Box jumps 15cm", "Box jumps 30cm", " ",
+      "Continuous jumps 5cm", "Continuous jumps 15cm"
+    )
+  ) +
   scale_y_continuous(
     limits = c(0, 14),
     expand = c(0, 0),
[…]

### 2022-11-01 figures/figS3.R
Adjust text size

@@ -57,7 +57,7 @@ boxplot_GRF_ver <- mechanical_load_data |>
   ) +
   theme_light() +
   theme(
-    axis.text.y = element_text(size = 13),
+    axis.text.y = element_text(size = 16),
     axis.title.y = element_text(size = 16)
   ) +
   labs(x = "", y = "pGRF (BW)")
@@ -85,10 +85,11 @@ boxplot_ACC_ver <- mechanical_load_data |>
   ) +
   theme_light() +
   theme(
-    axis.text.x = element_text(size = 13, angle = 45, hjust = 1),
-    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
+    axis.text.y = element_text(size = 16),
     axis.title.y = element_text(size = 16),
-    legend.title = element_blank()
+    legend.title = element_blank(),
+    legend.text = element_text(size = 16)
   ) +
   labs(x = "", y = quote("pACC" ~ (italic(g))))
 
@@ -105,7 +106,7 @@ boxplot_LR_ver <- mechanical_load_data |>
   ) +
   theme_light() +
   theme(
-    axis.text.y = element_text(size = 13),
+    axis.text.y = element_text(size = 16),
     axis.title.y = element_text(size = 16)
   ) +
   labs(x = "", y = quote("pLR" ~ (BW %.% s^-1)))
@@ -133,10 +134,11 @@ boxplot_AR_ver <- mechanical_load_data |>
   ) +
   theme_light() +
   theme(
-    axis.text.x = element_text(size = 13, angle = 45, hjust = 1),
-    axis.text.y = element_text(size = 13),
+    axis.text.x = element_text(size = 16, angle = 45, hjust = 1),
+    axis.text.y = element_text(size = 16),
     axis.title.y = element_text(size = 16),
     legend.title = element_blank(),
+    legend.text = element_text(size = 16)
   ) +
   labs(x = "", y = quote("pAR" ~ (italic(g) %.% s^-1)))
 

### 2022-11-01 figures/figS4.R

@@ -28,13 +28,13 @@ BA_GRF_res_ankle <- cv_res_GRF_models$ankle |>
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
+    plot.title = element_text(size = 18, hjust = 0.5),
     legend.title = element_blank(),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    legend.text = element_text(size = 18),
+    axis.title.y = element_text(size = 16),
+    axis.title.x = element_text(size = 16),
+    axis.text.y = element_text(size = 16),
+    axis.text.x = element_text(size = 16)
   ) +
   guides(alpha = "none") +
   labs(
@@ -60,13 +60,13 @@ BA_GRF_res_back <- cv_res_GRF_models$lower_back |>
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
+    plot.title = element_text(size = 18, hjust = 0.5),
     legend.title = element_blank(),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    legend.text = element_text(size = 18),
+    axis.title.y = element_text(size = 16),
+    axis.title.x = element_text(size = 16),
+    axis.text.y = element_text(size = 16),
+    axis.text.x = element_text(size = 16)
   ) +
   guides(alpha = "none") +
   labs(
@@ -92,13 +92,13 @@ BA_GRF_res_hip <- cv_res_GRF_models$hip |>
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
+    plot.title = element_text(size = 18, hjust = 0.5),
     legend.title = element_blank(),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    legend.text = element_text(size = 18),
+    axis.title.y = element_text(size = 16),
+    axis.title.x = element_text(size = 16),
[…]

### 2022-11-01 figures/figS5.R

@@ -30,13 +30,13 @@ BA_LR_res_ankle <- cv_res_LR_models$ankle %>%
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
+    plot.title = element_text(size = 18, hjust = 0.5),
     legend.title = element_blank(),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    legend.text = element_text(size = 18),
+    axis.title.y = element_text(size = 16),
+    axis.title.x = element_text(size = 16),
+    axis.text.y = element_text(size = 16),
+    axis.text.x = element_text(size = 16)
   ) +
   guides(alpha = "none") +
   labs(
@@ -64,13 +64,13 @@ BA_LR_res_back <- cv_res_LR_models$lower_back %>%
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
+    plot.title = element_text(size = 18, hjust = 0.5),
     legend.title = element_blank(),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    legend.text = element_text(size = 18),
+    axis.title.y = element_text(size = 16),
+    axis.title.x = element_text(size = 16),
+    axis.text.y = element_text(size = 16),
+    axis.text.x = element_text(size = 16)
   ) +
   guides(alpha = "none") +
   labs(
@@ -98,13 +98,13 @@ BA_LR_res_hip <- cv_res_LR_models$hip %>%
   ) +
   theme_light() +
   theme(
-    plot.title = element_text(size = 15, hjust = 0.5),
+    plot.title = element_text(size = 18, hjust = 0.5),
     legend.title = element_blank(),
-    legend.text = element_text(size = 12),
-    axis.title.y = element_text(size = 13),
-    axis.title.x = element_text(size = 13),
-    axis.text.y = element_text(size = 13),
-    axis.text.x = element_text(size = 13)
+    legend.text = element_text(size = 18),
+    axis.title.y = element_text(size = 16),
+    axis.title.x = element_text(size = 16),
[…]

### 2022-11-01 manuscript/manuscript.Rmd
Add supplemental figures

@@
 
-# Figure captions {.unnumbered}
+# Supplemental Material {.unnumbered}
 
-**Figure 1:** Distribution of the resultant vector of mechanical loading (peak ground reaction force and peak loading rate) and acceleration (peak acceleration and peak acceleration rate) variables per jump type and height combination.
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=0.7\textwidth]{figS1.png}
+	\end{measuredfigure}
+\end{figure}
+\footnotesize \textbf{Supplemental Figure S1.} An example of a participant performing each of the jump types assayed.
+Panels A and B show the initial (take-off phase) and final (landing phase) phases of the drop jumps, respectively, while Panels C and D show the initial and final phases of the box jumps and Panels E and F show the initial and final phase of the continuous jumps.
+
+\newpage
+
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=0.7\textwidth]{figS2.png}
+	\end{measuredfigure}
+\end{figure}
+\footnotesize \textbf{Supplemental Figure S2.} Accelerometers positioning on the body.
+Details of the accelerometer positioning on the ankle, lower back and hip are depicted in Panels A, B and C, respectively.
+Panel D shows all accelerometer placements simultaneously.
+
+\newpage
+
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{figS3.png}
+	\end{measuredfigure}
+\end{figure}
+\footnotesize \textbf{Supplemental Figure S3.} Distribution of the vertical vector of mechanical loading (peak ground reaction force and peak loading rate) and acceleration (peak acceleration and peak acceleration rate) variables per jump type and height combination.
 Abbreviations: BW, body weight; pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.
 
-**Figure 2:** Bland-Altman plots showing agreement between actual and predicted peak ground reaction force (Panel A) and loading rate (Panel B) for the resultant vector from hip-worn accelerometers.
-Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias \pm 1.96 standard deviation).
-Abbreviations: pGRF, peak ground reaction force; pLR, peak loading rate.
+\newpage
 
-**Supplemental Figure S1:** Distribution of the vertical vector of mechanical loading (peak ground reaction force and peak loading rate) and acceleration (peak acceleration and peak acceleration rate) variables per jump type and height combination.
-Abbreviations: BW, body weight; pACC, peak acceleration; pAR, peak acceleration rate; pGRF, peak ground reaction force; pLR, peak loading rate.
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{figS4.png}
+	\end{measuredfigure}
+\end{figure}
+\footnotesize \textbf{Supplemental Figure S4.} Bland-Altman plots showing agreement between actual and predicted peak ground reaction force for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias ± 1.96 standard deviation).
+Abbreviations: pGRF, peak ground reaction force.
 
-**Supplemental Figure S2:** Bland-Altman plots showing agreement between actual and predicted peak ground reaction force for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
-Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias ±1.96 standard deviation).
-Abbreviations: pGRF, peak ground reaction force.
+\newpage
+
+\begin{figure}[ht]
+	\captionsetup{singlelinecheck = false, format = hang, justification = raggedright, font = footnotesize, labelsep = space}
+	\centering
+	\begin{measuredfigure}
+		\includegraphics[width=\textwidth]{figS5.png}
+	\end{measuredfigure}
+\end{figure}
+\footnotesize \textbf{Supplemental Figure S5.} Bland-Altman plots showing agreement between actual and predicted peak loading rate for accelerometers worn at ankle (Panels A and D), lower back (Panels B and E) and hip (Panels C and F), for both the resultant vector (Panels A to C) and its vertical component (Panels D to F).
+Continuous thick lines show bias (average of the differences between actual and predicted values) while dashed lines show the limits of agreement (bias ± 1.96 standard deviation).
+Abbreviations: pLR, peak loading rate.

### 2022-11-02 figures/fig3.R
Delete files

@@ -1,153 +1,0 @@
-# Load packages -----------------------------------------------------------
-
-library(here)
-library(tidyverse)
-library(lvmisc)
-library(ggsci)
-library(patchwork)
-library(ragg)
-
-# Load and prepare data ---------------------------------------------------
-
-load(here("output", "veras2020_predictions.rda"))
-load(here("output", "loocv_data.rda"))
-
-plot_data_GRF <- cv_res_GRF_models$hip |>
-  left_join(
-    select(
-      hip_res_GRF_df,
-      id, trial, filename, subj, acc_placement, vector,
-      jump_type, jump_height, pGRF_N, .predicted_Veras2020
-    )
-  ) |>
-  select(pACC_g, pGRF_N, .predicted, .predicted_Veras2020) |>
-  pivot_longer(
-    cols = c(pGRF_N, .predicted, .predicted_Veras2020),
-    names_to = "measure",
-    values_to = "pGRF"
-  ) |>
-  mutate(
-    measure = recode_factor(
-      as.factor(measure),
-      pGRF_N = "Measured by the force plates",
-      .predicted = "Predicted by equation for jumping",
-      .predicted_Veras2020 = "Predicted by equation for walking"
-    )
-  )
-
-plot_data_LR <- cv_res_LR_models$hip |>
-  left_join(
-    select(
-      hip_res_LR_df,
-      id, trial, filename, subj, acc_placement, vector,
-      jump_type, jump_height, pLR_Ns, .predicted_Veras2020
-    )
-  ) |>
-  select(pAR_gs, pLR_Ns, .predicted, .predicted_Veras2020) |>
-  pivot_longer(
-    cols = c(pLR_Ns, .predicted, .predicted_Veras2020),
-    names_to = "measure",
-    values_to = "pLR"
-  ) |>
-  mutate(
-    measure = recode_factor(
-      as.factor(measure),
-      pLR_Ns = "Measured by the force plates",
-      .predicted = "Predicted by equation for jumping",
-      .predicted_Veras2020 = "Predicted by equation for walking"
-    )
-  )
-
-# Plot GRF predictions ----------------------------------------------------
-
-predicted_GRF_plot <- ggplot(plot_data_GRF) +
-  geom_point(
-    aes(x = pACC_g, y = pGRF, color = measure, shape = measure),
-    alpha = 0.3
-  ) +
-  geom_hline(yintercept = 0) +
[…]

### 2022-11-02 tables/tab2.R

@@ -1,64 +1,0 @@
-# Load packages and functions ---------------------------------------------
-
-library(here)
-library(tidyverse)
-library(lvmisc)
-library(broman)
-library(kableExtra)
-source(here("code", "funs.R"))
-
-# Load and prepare data ---------------------------------------------------
-
-load(here("output", "veras2020_accuracy.rda"))
-
-# Build table -------------------------------------------------------------
-
-tab2_df <- list(
-  lower_back_res_GRF_accuracy,
-  hip_res_GRF_accuracy,
-  lower_back_ver_GRF_accuracy,
-  hip_ver_GRF_accuracy,
-  lower_back_res_LR_accuracy,
-  hip_res_LR_accuracy,
-  lower_back_ver_LR_accuracy,
-  hip_ver_LR_accuracy
-) |>
-  map_dfr(~ mutate(.x, MAPE = as.numeric(MAPE) * 100)) |>
-  mutate(
-    across(where(is.numeric), myround, 1),
-    MAPE = paste0(MAPE, "\\%"),
-    Vector = rep(c("Resultant", "", "Vertical", ""), 2),
-    `Accelerometer placement` = rep(c("Lower back", "Hip"), 4),
-    .before = MAE
-  )
-
-tab2 <- knitr::kable(
-  tab2_df, booktabs = TRUE, escape = FALSE, linesep = "",
-  label = "none",
-  caption = "Regression equations, $R^2$ and accuracy indices",
-  col.names = c(
-    "Vector",
-    linebreak("Accelerometer\nplacement"),
-    "MAE",
-    "MAPE",
-    "RMSE"
-  ),
-  align = c(rep("l", 2), rep("r", 3))
-) %>%
-  footnote(
-    general = "Abbreviations: MAE, mean absolute error; MAPE, mean absolute percent error; RMSE, root mean square error",
-    general_title = "",
-    threeparttable = TRUE
-  ) %>%
-  # kable_styling(latex_options = "scale_down") %>%
-  pack_rows(
-    "Peak ground reaction force prediction", 1, 4,
-    latex_gap_space = "0.5em",
-    bold = FALSE
-  ) %>%
-  pack_rows(
-    "Peak loading rate prediction", 5, 8,
-    latex_gap_space = "0.5em",
-    bold = FALSE
-  ) %>%
-  landscape()

### 2022-11-02 tables/tab3.R

@@ -1,73 +1,0 @@
-# Load packages and functions ---------------------------------------------
-
-library(here)
-library(tidyverse)
-library(kableExtra)
-source(here("code", "funs.R"))
-
-# Load and prepare data ---------------------------------------------------
-
-load(here("output", "jump_type_accuracy.rda"))
-
-# Build table -------------------------------------------------------------
-
-# GRF table
-res_GRF_tab2_df <- build_accuracy_table(
-  drop_jumps_res_GRF_accuracy, "resultant", "drop"
-) |>
-  left_join(
-    build_accuracy_table(box_jumps_res_GRF_accuracy, "resultant", "box")
-  ) |>
-  left_join(
-    build_accuracy_table(
-      continuous_jumps_res_GRF_accuracy, "resultant", "continuous"
-    )
-  )
-ver_GRF_tab2_df <- build_accuracy_table(
-  drop_jumps_ver_GRF_accuracy, "vertical", "drop"
-) |>
-  left_join(
-    build_accuracy_table(box_jumps_ver_GRF_accuracy, "vertical", "box")
-  ) |>
-  left_join(
-    build_accuracy_table(
-      continuous_jumps_ver_GRF_accuracy, "vertical", "continuous"
-    )
-  )
-GRF_tab2_df <- rbind(res_GRF_tab2_df, ver_GRF_tab2_df)
-# LR table
-res_LR_tab2_df <- build_accuracy_table(
-  drop_jumps_res_LR_accuracy, "resultant", "drop"
-) |>
-  left_join(
-    build_accuracy_table(box_jumps_res_LR_accuracy, "resultant", "box")
-  ) |>
-  left_join(
-    build_accuracy_table(
-      continuous_jumps_res_LR_accuracy, "resultant", "continuous"
-    )
-  )
-ver_LR_tab2_df <- build_accuracy_table(
-  drop_jumps_ver_LR_accuracy, "vertical", "drop"
-) |>
-  left_join(
-    build_accuracy_table(box_jumps_ver_LR_accuracy, "vertical", "box")
-  ) |>
-  left_join(
-    build_accuracy_table(
-      continuous_jumps_ver_LR_accuracy, "vertical", "continuous"
-    )
-  )
-LR_tab2_df <- rbind(res_LR_tab2_df, ver_LR_tab2_df)
-# Join both tables
-tab2_df <- rbind(GRF_tab2_df, LR_tab2_df)
-# Remove vector name duplicates
[…]

### 2022-11-02 code/05_compare_equations.R
Delete file

@@ -1,112 +1,0 @@
-# Load packages and functions ---------------------------------------------
-
-library(here)
-library(tidyverse)
-library(lvmisc)
-source(here("code", "funs.R"))
-
-# Load and prepare data ---------------------------------------------------
-
-load(here("data", "mechanical_load_data.rda"))
-
-lower_back_res_data <- mechanical_load_data |>
-  filter(acc_placement == "lower_back" & vector == "resultant")
-hip_res_data <- mechanical_load_data |>
-  filter(acc_placement == "hip" & vector == "resultant")
-
-lower_back_ver_data <- mechanical_load_data |>
-  filter(acc_placement == "lower_back" & vector == "vertical")
-hip_ver_data <- mechanical_load_data |>
-  filter(acc_placement == "hip" & vector == "vertical")
-
-# Predict using Veras et al. 2020 -----------------------------------------
-
-# GRF
-lower_back_res_GRF_df <- predict_Veras2020(
-  lower_back_res_data,
-  outcome = "GRF", acc_placement = "lower_back", vector = "resultant"
-)
-hip_res_GRF_df <- predict_Veras2020(
-  hip_res_data,
-  outcome = "GRF", acc_placement = "hip", vector = "resultant"
-)
-
-lower_back_ver_GRF_df <- predict_Veras2020(
-  lower_back_ver_data,
-  outcome = "GRF", acc_placement = "lower_back", vector = "vertical"
-)
-hip_ver_GRF_df <- predict_Veras2020(
-  hip_ver_data,
-  outcome = "GRF", acc_placement = "hip", vector = "vertical"
-)
-
-# LR
-lower_back_res_LR_df <- predict_Veras2020(
-  lower_back_res_data,
-  outcome = "LR", acc_placement = "lower_back", vector = "resultant"
-)
-hip_res_LR_df <- predict_Veras2020(
-  hip_res_data,
-  outcome = "LR", acc_placement = "hip", vector = "resultant"
-)
-
-lower_back_ver_LR_df <- predict_Veras2020(
-  lower_back_ver_data,
-  outcome = "LR", acc_placement = "lower_back", vector = "vertical"
-)
-hip_ver_LR_df <- predict_Veras2020(
-  hip_ver_data,
[…]
